---
title: "Report"
output:
  html_document: default
  pdf_document: default
  always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include=FALSE}
library(tidyverse)
library(tigris)
library(rgdal)
library(ggplot2)
library(readxl)
library(lubridate)
library(leaflet)
theme_set(theme_classic())
library(lemon)
library(janitor)
library(writexl)
knit_print.data.frame <- lemon_print
```

## Import data

```{r}
setwd('~/Desktop/OSR2019')
sessions <- read.csv("DATA/B4Sprograms/clean_summer_session_days.csv") #17,243 obs of 18 vars
programs <- read.csv("DATA/B4Sprograms/clean_summer_programs.csv") #4,094 obs of 14 vars
```

## Analysis
### 1. What is the distribution in the number of programs available across the summer?

On what days there are summer program sessions? 
(note: analysis unit: program session day)

```{r}
# just in case
sessions$calendar_date <- as.POSIXct(sessions$calendar_date)

g1 <- ggplot(sessions, aes(calendar_date)) +
  geom_bar() + 
  labs(title="Distribution of out-of-school summer programs", 
       subtitle = "Session calendar days",
       caption="Source: Blueprint4Summer",
       x="Program session calendar date")

g1
```


A table on hours in each week could be more helpful. 
Resources https://cran.r-project.org/web/packages/aweek/aweek.pdf
The following fuctions get us the week of the year as decimal number (01–53) as defined in ISO 8601.
The ISO week-numbering year starts at the first day (Monday) of week 01 and ends at the Sunday before the new ISO year (hence without overlap or gap).
 One thing is this does not make sense straightforwardly. One way is to make an excile and merge it into this. Probably the fast method. 

```{r render=lemon_print}
sessions$week <- strftime(sessions$calendar_date, format = "%V")
# total hours in each week
hours_in_week <- sessions %>% 
  group_by(week) %>%
  summarise(hours = sum(duration)) %>%
  mutate(p = hours/sum(hours))
hours_in_week$p <-round(hours_in_week$p, digits = 2)
colnames(hours_in_week) <- c("Week", "Total hours", "Total hours percentage")

# read in time-week relational data
time_week <- read.csv("DATA/time.csv")
# merge
hours_in_week <- merge(hours_in_week, time_week, by = "Week")

hours_in_week <- hours_in_week %>% 
  arrange_at("Total hours", desc)

hours_in_week

# write this table to excel so we can hand fix the percentages that are smaller than 0.01
# write_xlsx(hours_in_week, "REPORT/TABLES/Table1.xlsx")
```

### 2. What is difference in access to sessions with different program length (in hours)?

#### Analysis unit: session day. 

```{r}
# summary statistics & top 3 freq
sessions$duration <- as.numeric(sessions$duration)
summary(sessions$duration)

# graph to check the distribution
g2 <- ggplot(sessions, aes(duration)) +
  geom_bar() + 
  labs(title="How long are program sessions per day?", 
       subtitle = "Summer programs, session day",
       caption="Source: Blueprint4Summer",
       x="Session day duration(hours)")
g2

# more detailed numbers with high-low order on frequency
duration_freq <- data.frame(tabyl(sessions$duration))
colnames(duration_freq) <- c("Hours", "Counts", "Percent")
duration_freq$Hours <- format(round(duration_freq$Hours, 2), nsmall = 2)
duration_freq$Percent <- format(round(duration_freq$Percent, 2), nsmall = 2)
duration_freq <- duration_freq %>% 
  arrange_at("Counts", desc)
duration_freq
# take top three 
duration_freq3 <- duration_freq[1:3,]

duration_freq3
```

#### According to the duration of the program session days, we define three categories of program length as the following:

1. half day: 4 hours and less than 4 hours
2. partial day: more than 4 hours to 7 hours (does not include 7 hours)
3. longer than one day: 7 hours and more

```{r}
sessions$day_length[sessions$duration <= 1] <- "One hour and less"
sessions$day_length[sessions$duration <= 4 & sessions$duration > 1] <- "One to four hours"
sessions$day_length[sessions$duration > 4 & sessions$duration < 7] <- "Four to seven hours"
sessions$day_length[sessions$duration >= 7] <- "Seven hours and more"
# set up factor levels 
sessions$day_length <- factor(sessions$day_length, levels = c("One hour and less", "One to four hours", "Four to seven hours", "Seven hours and more"))

day_freq_table <- data.frame(tabyl(sessions$day_length))
colnames(day_freq_table)[1] <- "Duration"
colnames(day_freq_table)[2] <- "Count"
# day_freq_table # uncomment this if you want to see details on exact numbers

g3 <- ggplot(day_freq_table, aes(x=reorder(Duration, percent), y=percent)) + 
  geom_bar(stat = "identity", width=0.8, fill="#b2182b") + 
  coord_flip() +
  labs(title="Percentage of Program Session Days with Different Duration", 
       caption="Source: Blueprint4Summer",
       x="", y="")
g3

day_freq_table 
```

#### Does program session day length vary by program type?

First create program categories for both level data
```{r}
#academic
programs$Academic <- FALSE
programs$Academic[
  which(grepl("academic|stem",programs$session_categories))
] <- TRUE

#art
programs$Art <- FALSE
programs$Art[
  which(grepl("art|cooking|dance|drama|music",programs$session_categories))
] <- TRUE

#has sports
programs$Sports <- FALSE
programs$Sports[
  which(grepl("sports",programs$session_categories))
] <- TRUE

#has nature
programs$Nature <- FALSE
programs$Nature[which(
  grepl("nature",programs$session_categories)
)] <- TRUE

#special needs
programs$Special_needs_offerings <- FALSE
programs$Special_needs_offerings[
  grepl("specialNeeds",programs$session_categories,ignore.case = TRUE)
] <- TRUE

#for session days
#academic
sessions$Academic <- FALSE
sessions$Academic[
  which(grepl("academic|stem",sessions$session_categories))
] <- TRUE

#art
sessions$Art <- FALSE
sessions$Art[
  which(grepl("art|cooking|dance|drama|music",sessions$session_categories))
] <- TRUE

#has sports
sessions$Sports <- FALSE
sessions$Sports[
  which(grepl("sports",sessions$session_categories))
] <- TRUE

#has nature
sessions$Nature <- FALSE
sessions$Nature[which(
  grepl("nature",sessions$session_categories)
)] <- TRUE

#special needs
sessions$Special_needs_offerings <- FALSE
sessions$Special_needs_offerings[
  grepl("specialNeeds",sessions$session_categories,ignore.case = TRUE)
] <- TRUE
```

Make a table that contains mean, median, min, max of duration by program type
Note: we intentionally double count if there are overlap in categories. 
```{r}
academic <- sessions %>% filter(Academic == T)
art <- sessions %>% filter(Art == T)
sports <- sessions %>% filter(Sports == T)
nature <- sessions %>% filter(Nature == T)
special_needs <- sessions %>% filter(Special_needs_offerings == T)

summary(special_needs$duration)
```

Make panels of day_length graph. (cannot do percentage because of double-counting)
```{r}
library(dplyr)
temp <- sessions %>% 
  gather(category, status, Academic:Special_needs_offerings) %>% 
  filter(status == T)
temp$category <- as.factor(temp$category)
temp$category <- factor(temp$category, levels = c("Academic", "Art", "Nature", "Sports", "Special_needs_offerings"))

temp2 <- temp %>% group_by(day_length, category) %>% summarise(count = n())
  
ggplot(temp2, aes( x = day_length, y = count)) + 
      geom_bar( stat = "identity" ) + 
      facet_wrap( ~ category ) + 
      coord_flip() +
      labs(title="Session Day Length by Program Type", 
      caption="Source: Blueprint4Summer",
       x="", y="")
```


### 3. What is the correlation between time and cost, specifically looking at cost/per hour?

```{r}
# cost per hour
programs$cost_hour <- programs$session_cost/programs$total.duration
programs$cost_hour <- as.numeric(programs$cost_hour)
programs$cost_hour <- round(programs$cost_hour, digits = 0) # round up to whole number
```

First we do an age break down.
3-5(only this age range, exclude);6-9;10-13;14-18
```{r}
head(programs)
programs$session_min_age <- as.numeric(programs$session_min_age)
programs$Age_range[programs$session_max_age <= 5] <- "3-5"
programs$Age_range[programs$session_max_age <= 9 & programs$session_min_age > 5] <- "6-9"
programs$Age_range[programs$session_max_age <= 13 & programs$session_max_age >= 10] <- "10-13"
programs$Age_range[programs$session_max_age <= 18 & programs$session_max_age >= 14] <- "14-18"
programs$Age_range[programs$session_max_age > 18] <- NA
programs$Age_range <- factor(programs$Age_range, levels = c("3-5", "6-9", "10-13", "14-18"))

library(dplyr)

costbyage <- programs %>% 
  group_by(Age_range) %>% 
  summarize(mean = mean(cost_hour),
            median = median(cost_hour),
            min = min(cost_hour),
            max = max(cost_hour)) #not sure why it does not throw out the second category

d1 <- programs %>% filter(Age_range == "3-5") 
summary(d1$cost_hour) # just for double check

d2 <- programs %>% filter(Age_range == "6-9") # caculate for the second category
summary(d2$cost_hour)
```

Then we drop all programs that only serve kids under 5 years old. 
```{r}
programs_age <- programs %>% filter(Age_range != "3-5" & !is.na(Age_range)) #3964

str(programs_age)
```

```{r}
freq_cost_hour <- data.frame(tabyl(programs_age$cost_hour)) 
colnames(freq_cost_hour)[1] <- "Cost_per_hour" 
freq_cost_hour <- freq_cost_hour %>% 
  arrange_at("n", desc)

summary(programs_age$cost_hour)
summary(programs_age$session_cost)

freq_cost_hour # 653 free program sessions

ggplot(programs_age, aes(cost_hour)) +
  geom_bar() + 
  labs(title="Distribution of program session cost", 
       subtitle = "Summer programs",
       caption="Source: Blueprint4Summer",
       x="Cost per hour")
freq_cost_hour
```

```{r}
g5 <- ggplot(programs_age, aes(total.duration, cost_hour)) +
         geom_point(size = 0.7, col = "Dark Grey", alpha = 0.4) + 
         labs(title="Correlation between time and cost", 
              subtitle = "Summer programs in Denver",
              caption="Source: Blueprint4Summer",
              x="Program Length (Hours)",
              y="Cost Per Hour ($)")

g5

# remove outliers (total duration > 100 hours)
temp <- programs_age %>% filter(total.duration < 100)
g9 <- ggplot(temp, aes(total.duration, cost_hour)) +
         geom_point(size = 0.7, col = "Dark Grey", alpha = 0.4) + 
         labs(title="Correlation between time and cost", 
              subtitle = "Summer programs in Denver",
              caption="Source: Blueprint4Summer",
              x="Program Length (Hours)",
              y="Cost Per Hour ($)")

g9
 
summary(programs_age$total.duration)
```

```{r}
cor(programs_age$cost_hour, programs_age$total.duration, use="complete.obs")
```

See if cost per hour varies by program type 
```{r}
str(programs_age)

temp <- programs_age %>% 
  gather(category, status, Academic:Special_needs_offerings) %>% 
  filter(status == T)
temp$category <- as.factor(temp$category)
temp$category <- factor(temp$category, levels = c("Academic", "Art", "Nature", "Sports", "Special_needs_offerings"))

d1 <- temp %>% filter(category == "Academic") 
summary(d1$cost_hour)

d2 <- temp %>% filter(category == "Art") 
summary(d2$cost_hour)
                                                 
d3 <- temp %>% filter(category == "Sports") 
summary(d3$cost_hour)

d4 <- temp %>% filter(category == "Nature") 
summary(d4$cost_hour)

d5 <- temp %>% filter(category == "Special_needs_offerings") 
summary(d5$cost_hour)
```


### 4. How many hours are going into programming across the city rather than just number of programs?

```{r}
# read in the relational data file
programs_nbhd <- read.csv("DATA/B4Sprograms/clean_summer_programs_nbhd.csv")
# filter out the NAs in nbhd name
programs_nbhd <- filter(programs_nbhd, !is.na(NBHD_Name)) #1980 rows 
time <- programs_nbhd %>% group_by(NBHD_Name) %>% summarise(sum(total.duration))
colnames(time) <- c("NBHD_NA", "Time")

# import nbhd shape file 
nbhd <- readOGR("DATA/nbhd_dem_shape", "nbhd_dem_shapes")
# merge time into nbhd, prepare for plotting
nbhd <- merge(nbhd, time, by = "NBHD_NA")

# color 
qpal <- colorQuantile("Blues", nbhd@data$Time, n = 5)

leaflet(nbhd) %>%
  addTiles() %>%
  setView(lng = -104.901531, lat = 39.722043, zoom = 11) %>% 
  addPolygons(stroke = FALSE, smoothFactor = 0.3,
              opacity = 1.0, fillOpacity = 0.5, 
              color = ~qpal(Time)) %>%
  addLegend("bottomright", 
              pal = qpal, 
              values = nbhd@data$Time,
              title = htmltools::HTML("Hours"),
              opacity = 1) 


summary(nbhd@data$Time)
```

### 5. Can we observe the variation in maximum capacity of programs by nbhd? 

Analysis unit: program session. 
```{r}
sum(is.na(programs_nbhd$session_size))
  # 1082 rows of missing data
  # filter them out
programs_temp <- programs_nbhd %>% 
  filter(session_size != " ")
  # this leaves us with 898 obs
summary(programs_temp$session_size) 

freq_size <- data.frame(tabyl(programs_temp$session_size)) 
colnames(freq_size)[1] <- "Session_size" 
freq_size <- freq_size %>% 
  arrange_at("n", desc)
freq_size 

nbhd_capacity <- programs_temp %>%
  group_by(NBHD_Name) %>%
  summarise(capacity = sum(session_size))

# create a new var with categories 
programs_temp$session_size2[programs_temp$session_size <= 20] <- "20 and fewer" 
programs_temp$session_size2[programs_temp$session_size > 20 & programs_temp$session_size < 100] <- "21 to 99" 
programs_temp$session_size2[programs_temp$session_size >= 100] <- "100 and more" 
tabyl(programs_temp$session_size2)

 
# now merge the numbers into nbhd shape file 
capacity <- programs_temp %>% group_by(NBHD_Name) %>% summarise(sum(session_size))
colnames(capacity) <- c("NBHD_NA", "Slots")

nbhd <- merge(nbhd, capacity, by = "NBHD_NA")

# Mapping
# color 
qpal <- colorQuantile("Greens", nbhd@data$Slots, n = 5)

leaflet(nbhd) %>%
  addTiles() %>%
  setView(lng = -104.901531, lat = 39.722043, zoom = 11) %>% 
  addPolygons(stroke = FALSE, smoothFactor = 0.3,
              opacity = 1.0, fillOpacity = 0.5, 
              color = ~qpal(Slots)) %>%
  addLegend("bottomright", 
              pal = qpal, 
              values = nbhd@data$Slots,
              title = htmltools::HTML("Slots"),
              opacity = 1) 
```


