---
title: "Google Analytics Functions"
author: "Andrew Taylor"
date: "7/6/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Functions
The Google Analytics API markdown shows our pull from the API, as well as the application of these functions to create a functional database. This markdown lists all functions created for that analysis.  

###Parsing Search Path Function
This function turns our html string into something we can split systematically
```{r}
ParsePath <- function(htmlString) {
  htmlString <-(gsub("<.*?>", "", htmlString))
  htmlString <- (gsub("/search","",htmlString))
  htmlString <- (gsub("&",",",htmlString))
  htmlString <- (gsub("?","",htmlString))
  htmlString <- (gsub("/?","",htmlString))
  htmlString <- (gsub("selected","",htmlString))
  return(htmlString)
}
```

###Creating recognition functions
These functions read the full parsed html string, determine the feature, and then return the split version of the url. Could we have combined these with the cleaning function? Probably. 
```{r}
ReturnKeyWords <- function(x){
  check <- grepl("keywords",x)
    if(check==TRUE){
    stringcheck <- str_split_fixed(x,",",Inf)
    for (i in 1:(length(stringcheck))){
      check.i <- grepl("keywords",stringcheck[i])
      if (check.i==TRUE){
        return(stringcheck[i])}
    }}}

Returnscholarships <- function(x){
  check <- grepl("sessionAttributes",x)
    if(check==TRUE){
    stringcheck <- str_split_fixed(x,",",Inf)
    for (i in 1:(length(stringcheck))){
      check.i <- grepl("scholarships",stringcheck[i])
      if (check.i==TRUE){
        return(stringcheck[i])}
    }}}

ReturnspecialNeeds <- function(x){
  check <- grepl("sessionAttributes",x)
    if(check==TRUE){
    stringcheck <- str_split_fixed(x,",",Inf)
    for (i in 1:(length(stringcheck))){
      check.i <- grepl("specialNeeds",stringcheck[i])
      if (check.i==TRUE){
        return(stringcheck[i])}
    }}}

ReturnGifted <- function(x){
  check <- grepl("sessionAttributes",x)
    if(check==TRUE){
    stringcheck <- str_split_fixed(x,",",Inf)
    for (i in 1:(length(stringcheck))){
      check.i <- grepl("giftedStudent",stringcheck[i])
      if (check.i==TRUE){
        return(stringcheck[i])}
    }}}

ReturnBeforeCare <- function(x){
  check <- grepl("sessionAttributes",x)
    if(check==TRUE){
    stringcheck <- str_split_fixed(x,",",Inf)
    for (i in 1:(length(stringcheck))){
      check.i <- grepl("offersBeforeAfterCare",stringcheck[i])
      if (check.i==TRUE){
        return(stringcheck[i])}
    }}}


ReturnTime <- function(x){
  check <- grepl("sessionTimes",x)
  if(check==TRUE){
    stringcheck <- str_split_fixed(x,",",Inf)
    for (i in 1:(length(stringcheck))){
      check.i <- grepl("sessionTimes",stringcheck[i])
      if (check.i==TRUE){
        return(stringcheck[i])}
    }}}

ReturnMaxCost <- function(x){
  check <- grepl("maximumCost",x)
  if(check==TRUE){
    stringcheck <- str_split_fixed(x,",",Inf)
    for (i in 1:(length(stringcheck))){
      check.i <- grepl("maximumCost",stringcheck[i])
      if (check.i==TRUE){
        return(stringcheck[i])}
    }}}

ReturnMinCost <- function(x){
  check <- grepl("minimumCost",x)
  if(check==TRUE){
    stringcheck <- str_split_fixed(x,",",Inf)
    for (i in 1:(length(stringcheck))){
      check.i <- grepl("minimumCost",stringcheck[i])
      if (check.i==TRUE){
        return(stringcheck[i])}
    }}}

ReturnDistance <- function(x){
  check <- grepl("distance",x)
  if(check==TRUE){
    stringcheck <- str_split_fixed(x,",",Inf)
    for (i in 1:(length(stringcheck))){
      check.i <- grepl("distance",stringcheck[i])
      if (check.i==TRUE){
        return(stringcheck[i])}
    }}}

ReturnCategory <- function(x){
  check <- grepl("Categories",x)
  if(check==TRUE){
    stringcheck <- str_split_fixed(x,",",Inf)
    for (i in 1:(length(stringcheck))){
      check.i <- grepl("Categories",stringcheck[i])
      if (check.i==TRUE){
        return(stringcheck[i])}
    }}}

ReturnGender <- function(x){
  check <- grepl("gender",x)
  if(check==TRUE){
    stringcheck <- str_split_fixed(x,",",Inf)
    for (i in 1:(length(stringcheck))){
      check.i <- grepl("gender",stringcheck[i])
      if (check.i==TRUE){
        return(stringcheck[i])}
    }}}

ReturnMaxAge <- function(x){
  check <- grepl("maximumAge",x)
  if(check==TRUE){
    stringcheck <- str_split_fixed(x,",",Inf)
    for (i in 1:(length(stringcheck))){
      check.i <- grepl("maximumAge",stringcheck[i])
      if (check.i==TRUE){
        return(stringcheck[i])}
    }}}

ReturnMinAge <- function(x){
  check <- grepl("minimumAge",x)
  if(check==TRUE){
    stringcheck <- str_split_fixed(x,",",Inf)
    for (i in 1:(length(stringcheck))){
      check.i <- grepl("minimumAge",stringcheck[i])
      if (check.i==TRUE){
        return(stringcheck[i])}
    }}}

ReturnSort <-function(x){
  check <- grepl("sort",x)
  if(check==TRUE){
    stringcheck <- str_split_fixed(x,",",Inf)
    for (i in 1:(length(stringcheck))){
      check.i <- grepl("sort",stringcheck[i])
      if (check.i==TRUE){
        return(stringcheck[i])}
    }}}

ReturnLocation <- function(x){
  check <- grepl("currentLocation",x)
  if(check==TRUE){
    stringcheck <- str_split_fixed(x,",",Inf)
    for (i in 1:(length(stringcheck))){
      check.i <- grepl("currentLocation",stringcheck[i])
      if (check.i==TRUE){
        return(stringcheck[i])}
    }}}
```


###Cleaning Functions
Now that we've returned the split version of the html string, we need to clean it up so that only the relevant information is included. 
```{r}
CleanHash<- function(x){
  stringcheck <- str_split_fixed(x,"=",Inf)
  ifelse(length(stringcheck)>1,return(stringcheck[2]),"")
}
```
