---
title: "Access_Index_Functions"
author: "Andrew Taylor"
date: "7/23/2018"
output: html_document
---

##Intro
This markdown defines functions for use in calculating an access index per user specifications on filtering reschool program data. Additionally, this markdown provides leaflet code for plotting the results of the index by census block group. This markdown reads census_centroids as a shapefile from my machine, as we haven't pushed this to the RDS yet.
**As of July 25th** we have decided not to push multiple access index files to the the RDS, and instead to rely on the script here to run specific indicies.

```{r}
library(rgeos)
library(rgdal)
library(raster)
library(tidyverse)
library(RPostgreSQL)
```

###RDS Pull
```{r}
mypath <- dirname(rstudioapi::getActiveDocumentContext()$path)
setwd(mypath)

# loads the PostgreSQL driver
drv <- dbDriver("PostgreSQL")

# make a file called "cred.txt" of the form
#
# user: "YOUR_AWS_USERNAME"
# password: "YOUR_AWS_PASSWORD"
#
# in the directory above the osr_dssg2018 folder

source(file.path(dirname(dirname(mypath)),"cred.txt"))

# creates a connection to the postgres database
# note that "con" will be used later in each connection to the database
con <- dbConnect(drv, dbname = "dssg2018uw",
                 host = "localhost", port =9000,
                 user = user, password = password)

#get reschool_programs
reschool_programs <- dbGetQuery(con, "select * from clean.reschool_summer_programs")
block_distance <- dbGetQuery(con, "select * from clean.block_distance")

#disconnect
dbDisconnect(con) 
dbUnloadDriver(drv)
```

###Establishing Cost Thresholds and Umbrella Program Types
```{r}
reschool_programs$academic <- reschool_programs$has_academic==TRUE | reschool_programs$has_stem==TRUE

reschool_programs$art <- reschool_programs$has_arts==TRUE | reschool_programs$has_cooking==TRUE | reschool_programs$has_dance==TRUE | reschool_programs$has_drama==TRUE | reschool_programs$has_music==TRUE

#cost thresolds, to be updated w/ time data from selam
reschool_programs$free <- reschool_programs$session_cost==0
reschool_programs$lowcost <- reschool_programs$session_cost<=50
reschool_programs$highcost <- reschool_programs$session_cost>50

#establishing "1" per row for use in aggergation
reschool_programs$n <- 1 
```


###Defining for Flexible Access Index Calculation
Calculating the Access Index requires a few discrete steps. Currently, these are established with the functions defined below. 
```{r}
#1. Setting type and cost parameteres

type <- c("academic","art") #needs to be a vector of whatever get's passed to shiny
cost <- c("free") #these are examples, if set to NULL for either, the set of functions will return the full index

#2. First we subset to ReSchool Programs Dataset per user specificiations

Make_Subset <- function(dataframe,type=NULL,cost=NULL){
  dataframe$n <- 1
  if(is.null(type)){
    dataframe <- dataframe
    }
  if(length(type)==1){
    dataframe <- dataframe[dataframe[,type],]
  }
  if(length(type)>1){
    dataframe <- dataframe[apply(FUN=any,X=dataframe[,c(type)],MARGIN=1),]
  }
  if(is.null(cost)){
    return(dataframe)
  }
  else{
     dataframe <- dataframe <- dataframe[dataframe[,cost],]
  }
}

#3. Then we aggregate sessions the subset per the specifications

Aggregate_Subset <- function(dataframe){
  return(aggregate(n ~ session_address_1+lat+long, data=dataframe,FUN=sum)) #note that "n" in this case ==1 per row, which n this case corresponds to one unique program session
}

#4. Here we merge the aggregated sessiosn into the block_distance data set to incorporate travel times 

Merge_Set <- function(dataframe){
  relevant_columns <- c("Id2","session_address_1","driving_morning","walking_morning","transit_morning","kilometers","block_lat","block_long")
  block_distance_new <- block_distance[,relevant_columns] #here we're cutting duplicate program_addresses, and other antiquated info (previous aggregations, PCT vhcl ownership) that was use in previous calculations
  temp <- merge(block_distance_new,dataframe,all.y=TRUE)
  temp$n[is.na(temp$n)==TRUE] <- 0
  return(temp)
}

#5 Here we define the gravity threshold decay function. Note that as of July 25th we've decided not use this equation, and instead use the exponential equation per the results of the sensitivity analysis. 

decay_fxn <- function(dataframe,mode){
  if (mode=="transit") {
    scale = 5
    column="transit_morning"
  }
  if (mode=="drive"){
    scale = 5
    column="driving_morning"
  }
  mode=mode
  dataframe[,"n"]*(ifelse(dataframe[,column]<=scale,1,(1/(dataframe[,column]/scale))^2))
}

#6 Here we define the expoential decay function

decay_fxn_exp <- function(dataframe,mode){
  if (mode=="transit") {
    scale = 5
    column="transit_morning"
  }
  if (mode=="drive"){
    scale = 5
    column="driving_morning"
  }
  mode=mode
  dataframe[,"n"]*exp(-dataframe[,column]/scale)
}

#7 here we define a function to normalize results on a 0-1 scale. Note that after we calculate an total acess index in this script, max will be == the max AI of any one blockgroup.

normalize <- function(vec,max) {
  return(vec/max)
}
```

###Get total access index
This is critical to establishing a range to normalize all subset values relative to the overall value. The idea being, this way the max access = access index score at all prorams at all costs with the quickest (driving) travel times. Access will then decrease if the user instead view access by transit or a subset of programs. 
```{r}
library(scales)
#User specifies query (type)

#subset dataframe by user selection
user_set <- reschool_programs

#aggregate dataframe by user selection
agg_set <- Aggregate_Subset(user_set)

#merge n of user selected programs to block distances dataframe
merged_set <- Merge_Set(agg_set)

#calculate access index per location 

# if we don't have transit info, instead use the walking distance
merged_set$transit_morning[is.na(merged_set$transit_morning)] <- merged_set$walking_morning[is.na(merged_set$transit_morning)]
merged_set$transit_morning[merged_set$transit_morning==0] <- merged_set$walking_morning[merged_set$transit_morning==0]

merged_set$CarAccessIndex <- decay_fxn_exp(merged_set,"drive") 
merged_set$TransitAccessIndex <- decay_fxn_exp(merged_set,"transit") 

#get average of AccessIndex col, return to summary dataframe by blockgroup ID
summary_index_total <- aggregate(CarAccessIndex ~ Id2,data=merged_set,FUN=sum)
max_access <- max(summary_index_total$CarAccessIndex)
summary_index_total$CarAccessIndex <- normalize(summary_index_total$CarAccessIndex,max_access)
```

###User Script for Driving
This is the exact script to use to query, for an example type.
```{r}
#User specifies query (type)
#type <- c("art") #"type" can include categories or cost thresholds. Pending cost per day calculations from Selam
#cost <- c("free")
type <- c("academic")
cost <- NULL

#subset dataframe by user selection
user_set <- Make_Subset(reschool_programs,type,cost)

#aggregate dataframe by user selection
agg_set <- Aggregate_Subset(user_set)

#merge n of user selected programs to block distances dataframe
merged_set <- Merge_Set(agg_set)

#calculate access index per location 
merged_set$AccessIndex <- decay_fxn_exp(merged_set,"drive") #why not apply? who knows

#get average of AccessIndex col, return to summary dataframe by blockgroup ID
summary_index_user <- aggregate(AccessIndex ~ Id2,data=merged_set,FUN=sum) #critical that we use sum, not mean, BECAUSE, for some specifications some program lengths will drop

#normalize
summary_index_user$AccessIndex <- normalize(summary_index_user$AccessIndex,max_access) 
```

###User Script for Transit
```{r}
library(scales)
#User specifies query (type)
#type <- c("has_sports","art")
type <- c("academic")
cost <- NULL

#subset dataframe by user selection
user_set <- Make_Subset(reschool_programs,type)

#aggregate dataframe by user selection
agg_set <- Aggregate_Subset(user_set)

#merge n of user selected programs to block distances dataframe
merged_set <- Merge_Set(agg_set)

#replace na and 0 transit times with wallking times
merged_set$transit_morning[is.na(merged_set$transit_morning)] <- merged_set$walking_morning[is.na(merged_set$transit_morning)] #there shouldn't be NAs or 0s
merged_set$transit_morning[merged_set$transit_morning==0] <- merged_set$walking_morning[merged_set$transit_morning==0] #but if there is, as sometimes occurs with the API calls, here we replace non-existant transit times with walking times, under the assumption the purpose of this index is to view access for individuals without a car

#calculate access index per location, #get average of AccessIndex col, return to summary dataframe by blockgroup ID
system.time(merged_set$AccessIndex <- decay_fxn_exp(merged_set,"transit"))
system.time(summary_index_user_transit <- aggregate(AccessIndex ~ Id2,data=merged_set,FUN=sum))

#normalize
summary_index_user_transit$AccessIndex <- normalize(summary_index_user_transit$AccessIndex,max_access)
```

###Get Census Blocks Data for Plotting
Note that you need all files in this directory, and that it is not in the RDS. 
```{r}
library(rgeos)
library(rgdal)
library(raster)

dirname(mypath))
#census_blocks <- SpatialPointsDataFrame(gCentroid(spdf, byid=TRUE), spdf@data, match.ID = FALSE)
```

###Plot Driving Summary Index
```{r}
library(leaflet)
library(RColorBrewer)

#Ordering
census_blocks$Id2 <- as.numeric(census_blocks$Id2)
merged_set$Id2 <- as.numeric(merged_set$Id2)
summary_index_user$Id2 <- as.numeric(summary_index_user$Id2)

census_blocks <- census_blocks[order(census_blocks$Id2),]
merged_set <- merged_set[order(merged_set$Id2),]
summary_index_user <- summary_index_user[order(summary_index_user$Id2),] 

#specifiying color pallet
colourCount <- length(unique(agg_set$n))
getPalette <- colorRampPalette(brewer.pal(9,"Reds"))

pal_legend <- colorBin(c("#F0F8FF","#A1CAF1","#99BADD","#71A6D2","#6495ED","#318CE7","#4682B4","#1560BD","#0047AB","#0F4D92","#08457E","#003366","#002147"),domain = 0:1,bins=10,pretty = TRUE)

pal <- colorBin(c("#F0F8FF","#A1CAF1","#99BADD","#71A6D2","#6495ED","#318CE7","#4682B4","#1560BD","#0047AB","#0F4D92","#08457E","#003366","#002147"),domain = 0:1,bins=1000,pretty = TRUE)

pal2 <- colorNumeric(
  palette = rev(getPalette(colourCount)),
  domain = range(agg_set$n))

#indicating one popup info
info <- paste("n",agg_set$n)

#creating leaflet map
Ai_driving <- leaflet(census_blocks) %>%
  addTiles() %>%
  addPolygons(color = "RdYlBu",weight =1,smoothFactor = 0.5,opacity = 0.5,fillOpacity = 0.75,
              fillColor = ~pal(summary_index_user$AccessIndex),
              popup=paste(as.numeric(summary_index_user$Id2),summary_index_user$AccessIndex,as.numeric(census_blocks$Id2)),
              highlightOptions = highlightOptions(color = "white", weight = 0.1)) %>%
  
  addLegend(pal = pal_legend, values = ~summary_index_user$AccessIndex, opacity = 1,title = "Driving AI Score") %>%
  
  addCircles(data = agg_set, 
             opacity = 1,weight=7, popup=info,
             color = ~pal2(agg_set$n) ) %>%
  
  setView(lat=39.7,lng=-104.9,zoom=11.5)
```

###Plot Transit Index
```{r}
library(leaflet)
library(RColorBrewer)

#Ordering
census_blocks$Id2 <- as.numeric(census_blocks$Id2)
merged_set$Id2 <- as.numeric(merged_set$Id2)
summary_index_user_transit$Id2 <- as.numeric(summary_index_user_transit$Id2)

census_blocks <- census_blocks[order(census_blocks$Id2),]
merged_set <- merged_set[order(merged_set$Id2),]
summary_index_user_transit <- summary_index_user_transit[order(summary_index_user_transit$Id2),] 

#specifiying color pallet
colourCount <- length(unique(agg_set$n))
getPalette <- colorRampPalette(brewer.pal(9,"Reds"))

pal_legend <- colorBin(c("#F0F8FF","#A1CAF1","#99BADD","#71A6D2","#6495ED","#318CE7","#4682B4","#1560BD","#0047AB","#0F4D92","#08457E","#003366","#002147"),domain = 0:1,bins=10,pretty = TRUE)

pal <- colorBin(c("#F0F8FF","#A1CAF1","#99BADD","#71A6D2","#6495ED","#318CE7","#4682B4","#1560BD","#0047AB","#0F4D92","#08457E","#003366","#002147"),domain = 0:1,bins=1000,pretty = TRUE)

pal2 <- colorNumeric(
  palette = rev(getPalette(colourCount)),
  domain = range(agg_set$n))

#indicating one popup info
info <- paste("n",agg_set$n)

#creating leaflet map
Ai_transit <- leaflet(census_blocks) %>%
  addTiles() %>%
  addPolygons(color = "RdYlBu",weight =1,smoothFactor = 0.5,opacity = 0.5,fillOpacity = 0.9,
              fillColor = ~pal(summary_index_user_transit$AccessIndex),
              popup=paste(as.numeric(summary_index_user_transit$Id2),summary_index_user_transit$AccessIndex,as.numeric(census_blocks$Id2)),
              highlightOptions = highlightOptions(color = "white", weight = 0.1)) %>%
  
  addLegend(pal = pal_legend, values = ~summary_index_user_transit$AccessIndex, opacity = 1,title = "Transit AI Score") %>%
  
  addCircles(data = agg_set, 
             opacity = 1,weight=7, popup=info,
             color = ~pal2(agg_set$n) ) %>%
  
  setView(lat=39.7,lng=-104.9,zoom=11.5)
```

###Diff between transit and driving
Here we check the difference between them. This is a validation test, as per the specification we're using, where both transit and driving use the same decay function with the same scale, transit should always be lower than driving, particularly in places with lots of quick access to multiple programs. 
```{r}
library(leaflet)
library(RColorBrewer)

#Ordering
census_blocks$Id2 <- as.numeric(census_blocks$Id2)
merged_set$Id2 <- as.numeric(merged_set$Id2)
summary_index_user_transit$Id2 <- as.numeric(summary_index_user_transit$Id2)

census_blocks <- census_blocks[order(census_blocks$Id2),]
merged_set <- merged_set[order(merged_set$Id2),]
summary_index_user <- summary_index_user[order(summary_index_user$Id2),] 

summary_index_user$dif <- summary_index_user$AccessIndex - summary_index_user_transit$AccessIndex

#specifiying color pallet
colourCount <- length(unique(agg_set$n))
getPalette <- colorRampPalette(brewer.pal(9,"Reds"))

#pal <- colorNumeric(
  #palette = "RdYlBu",
  #domain =  summary_index_total$dif)

pal2 <- colorNumeric(
  palette = rev(getPalette(colourCount)),
  domain = range(agg_set$n))

#indicating one popup info
info <- paste("n",agg_set$n)

#creating leaflet map
Ai_leaflet_dif <- leaflet(census_blocks) %>%
  addTiles() %>%
  addPolygons(color = "RdYlBu",weight =1,smoothFactor = 0.5,opacity = 0.5,fillOpacity = 0.9,
              fillColor = ~pal(summary_index_user$dif),
              popup=paste(as.numeric(summary_index_user$Id2),summary_index_user$dif,as.numeric(census_blocks$Id2)),
              highlightOptions = highlightOptions(color = "white", weight = 0.1)) %>%
  
  addLegend(pal = pal, values = ~summary_index_user$dif, opacity = 1,title = "Dif in Access Index Score") %>%
  
  addCircles(data = agg_set, 
             opacity = 1,weight=7, popup=info,
             color = ~pal2(agg_set$n) ) %>%
  
  setView(lat=39.7,lng=-104.9,zoom=11.5)
```


