---
title: "Access_Index_SA"
author: "Andrew Taylor"
date: "7/24/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Intro
This markdown performs the sensitivity analysis on the access index equation, comparing the changes in blockgroup access index scores/ranks (as they compare to other blockgroups), for the 1/r^2 function with indifference threshold of 5-15 minutes, as well as for the exp function with indifference threshold of 5-15 minutes, and the gravity model, and **THE SOFT GRAVITY** model, which is the choice we end up using. We test with scale 5-15 minutes though there is no change in rank at any stage. This code also produces visualizations to describe these results. 

**Note** the markdown makes use of the Access Index Functions defined in the Access_Index_Functions markdown, as well as "census_blocks" which is currently running off our machines individualls. You'll need to grab those functions and the centroid file to run this markdown. 

###RDS Pull
```{r}
library(rgeos)
library(rgdal)
library(raster)
library(tidyverse)
library(RPostgreSQL)
# loads the PostgreSQL driver
drv <- dbDriver("PostgreSQL")

# creates a connection to the postgres database
# note that "con" will be used later in each connection to the database
con <- dbConnect(drv, dbname = "dssg2018uw",
                 host = "localhost", port = 9000,
                 user = user, password = password) #local source credentials

#get reschool_programs
reschool_programs <- dbGetQuery(con, "select * from clean.reschool_summer_programs")
block_distance <- dbGetQuery(con, "select * from clean.block_distance")

#disconnect
dbDisconnect(con) 
dbUnloadDriver(drv)
```

##Testing the effect of scale
###Sensitivity Analysis Driving Test
Because we are to some extent selecting an arbitrary indifference threshold of five minutes, it is critical that we assess how sensitive our conclusions are to threshold. So here we test different indifference thresholds and see how the order of top block groups changes. Basically, we scale up the threshold from 5 five minutes to 15 minutes, checking changes in the order of the block groups at each minute threshold. 
```{r}
#run overall access index driving
library(scales)
#User specifies query (type), NULL = all
type <- NULL
cost <- NULL

#subset dataframe by user selection
user_set <- Make_Subset(reschool_programs,type,cost)

#aggregate dataframe by user selection
agg_set <- Aggregate_Subset(user_set)

#merge n of user selected programs to block distances dataframe
merged_set <- Merge_Set(agg_set)

#calculate access index per location 
merged_set$AccessIndex <- decay_fxn(merged_set,"drive") #why not apply? who knows

#get average of AccessIndex col, return to summary dataframe by blockgroup ID
summary_index_user <- aggregate(AccessIndex ~ Id2,data=merged_set,FUN=sum) #critical that we use sum, not mean, BECAUSE, for some specifications some program lengths will drop

#make base dataframe
summary_index_user <- summary_index_user[order(summary_index_user$AccessIndex),]
rownames(summary_index_user)  <- NULL
summary_index_user$base_rank <- rownames(summary_index_user)

for (i in 1:9){
#generate overall
type <- NULL
cost <- NULL

#subset dataframe by user selection
user_set <- Make_Subset(reschool_programs,type,cost)

#aggregate dataframe by user selection
agg_set <- Aggregate_Subset(user_set)

#merge n of user selected programs to block distances dataframe
merged_set <- Merge_Set(agg_set)

#alter decay function
decay_fxn_sensitive <- function(dataframe,mode){
  if (mode=="transit") {
    scale = 5+i
    column="transit_morning"
  }
  if (mode=="drive"){
    scale = 5+i
    column="driving_morning"
  }
  mode=mode
  dataframe[,"n"]*(ifelse(dataframe[,column]<=scale,1,(1/(dataframe[,column]/scale))^2))
}

#apply decay function
merged_set$AccessIndex <- decay_fxn_sensitive(merged_set,"drive")

#get average of AccessIndex col, return to summary dataframe by blockgroup ID
summary_index_mover <- aggregate(AccessIndex ~ Id2,data=merged_set,FUN=sum)

#order
summary_index_mover <- summary_index_mover[order(summary_index_mover$AccessIndex),]

#prep for merge
rownames(summary_index_mover) <- NULL
summary_index_mover$rank.i <- rownames(summary_index_mover)
colnames(summary_index_mover)[colnames(summary_index_mover)=="rank.i"] <- paste("rank",i)

#print summary of AI for validation on if this is working
print(summary(summary_index_mover$AccessIndex))
summary_index_mover$AccessIndex <- NULL

#merge new data to dataframe
summary_index_user <- merge(summary_index_user,summary_index_mover,by="Id2",all.y=TRUE)
print(paste(i,"iteration..."))
}

#store as a new data frame
sensitivity_frame <- summary_index_user 

#get the total changes in rank from threshold of five minutes to 16 minutes
sensitivity_frame$changes <- abs(as.numeric(sensitivity_frame$base_rank) - as.numeric(sensitivity_frame$`rank 9`))
```

###Plot driving changes in rank
```{r}
library(reshape2)
library(ggplot2)
library(gridExtra)
#Prep & order frame
sensitivity_frame <- summary_index_user 
sensitivity_frame$base_rank <- as.numeric(sensitivity_frame$base_rank)
sensitivity_frame <- sensitivity_frame[order(sensitivity_frame$base_rank),]

#subsetframe
top_25 <- sensitivity_frame[1:25,]
sa_25_50 <- sensitivity_frame[1:50,]
sa_50_75 <- sensitivity_frame[1:75,]
sa_75_100 <- sensitivity_frame[1:100,]
sa_tail <- sensitivity_frame[456:481,]
sa_tailstart <- sensitivity_frame[406:481,]


#Top 25 Plot
df <- melt(top_25,id=c("Id2","AccessIndex"),measure.vars=c("base_rank","rank 1","rank 2","rank 3","rank 4","rank 5","rank 6","rank 7","rank 8","rank 9"))
df$value <- as.numeric(df$value)
df$variable <- as.factor(df$variable)

driving_Top_25_Plot <- ggplot(df, aes(variable,value,group = Id2, color = as.factor(Id2))) + geom_line() + theme(legend.position="none") + labs(title="top 25") + ylim(1,490)

#Top 50 Plot
df <- melt(sa_25_50,id=c("Id2","AccessIndex"),measure.vars=c("base_rank","rank 1","rank 2","rank 3","rank 4","rank 5","rank 6","rank 7","rank 8","rank 9"))
df$value <- as.numeric(df$value)
df$variable <- as.factor(df$variable)

driving_Top_50_Plot <- ggplot(df, aes(variable,value,group = Id2, color = as.factor(Id2))) + geom_line() + theme(legend.position="none") + labs(title="top 50") + ylim(1,490)

#Top 75 Plot
df <- melt(sa_50_75,id=c("Id2","AccessIndex"),measure.vars=c("base_rank","rank 1","rank 2","rank 3","rank 4","rank 5","rank 6","rank 7","rank 8","rank 9"))
df$value <- as.numeric(df$value)
df$variable <- as.factor(df$variable)

driving_Top_75_Plot <- ggplot(df, aes(variable,value,group = Id2, color = as.factor(Id2))) + geom_line() + theme(legend.position="none") + labs(title="top 75") + ylim(1,490)

#Top 100 Plot
df <- melt(sa_75_100,id=c("Id2","AccessIndex"),measure.vars=c("base_rank","rank 1","rank 2","rank 3","rank 4","rank 5","rank 6","rank 7","rank 8","rank 9"))
df$value <- as.numeric(df$value)
df$variable <- as.factor(df$variable)

driving_Top_100_Plot <- ggplot(df, aes(variable,value,group = Id2, color = as.factor(Id2))) + geom_line() + theme(legend.position="none") + labs(title="top 100") + ylim(1,490)

#Bottom 50
df <- melt(sa_tail,id=c("Id2","AccessIndex"),measure.vars=c("base_rank","rank 1","rank 2","rank 3","rank 4","rank 5","rank 6","rank 7","rank 8","rank 9"))
df$value <- as.numeric(df$value)
df$variable <- as.factor(df$variable)

driving_Tail_Plot <- ggplot(df, aes(variable,value,group = Id2, color = as.factor(Id2))) + geom_line() + theme(legend.position="none") + labs(title="Bottom 50") + ylim(1,490)

#Bottom 25
df <- melt(sa_tailstart,id=c("Id2","AccessIndex"),measure.vars=c("base_rank","rank 1","rank 2","rank 3","rank 4","rank 5","rank 6","rank 7","rank 8","rank 9"))
df$value <- as.numeric(df$value)
df$variable <- as.factor(df$variable)

driving_Bottom25_Plot <- ggplot(df, aes(variable,value,group = Id2, color = as.factor(Id2))) + geom_line() + theme(legend.position="none") + labs(title="Bottom 100") + ylim(1,490)

#grid.arrange(driving_Top_25_Plot,driving_Top_50_Plot,driving_Top_75_Plot,driving_Top_100_Plot,driving_Bottom25_Plot,driving_Tail_Plot,ncol=3)
```

###Redo with transit
```{r}
#run overall access index driving
library(scales)
#User specifies query (type)
#type <- c("art") #"type" can include categories or cost thresholds. Pending cost per day calculations from Selam
#cost <- c("free")
type <- NULL
cost <- NULL

#subset dataframe by user selection
user_set <- Make_Subset(reschool_programs,type,cost)

#aggregate dataframe by user selection
agg_set <- Aggregate_Subset(user_set)

#merge n of user selected programs to block distances dataframe
merged_set <- Merge_Set(agg_set)

#calculate access index per location 
merged_set$AccessIndex <- decay_fxn(merged_set,"transit") #why not apply? who knows

#get average of AccessIndex col, return to summary dataframe by blockgroup ID
summary_index_user <- aggregate(AccessIndex ~ Id2,data=merged_set,FUN=sum) #critical that we use sum, not mean, BECAUSE, for some specifications some program lengths will drop

#make base dataframe
summary_index_user <- summary_index_user[order(summary_index_user$AccessIndex),]
rownames(summary_index_user)  <- NULL
summary_index_user$base_rank <- rownames(summary_index_user)

for (i in 1:9){
#generate overall
type <- NULL
cost <- NULL

#subset dataframe by user selection
user_set <- Make_Subset(reschool_programs,type,cost)

#aggregate dataframe by user selection
agg_set <- Aggregate_Subset(user_set)

#merge n of user selected programs to block distances dataframe
merged_set <- Merge_Set(agg_set)

#alter decay function
decay_fxn_sensitive <- function(dataframe,mode){
  if (mode=="transit") {
    scale = 5+i
    column="transit_morning"
  }
  if (mode=="drive"){
    scale = 5+i
    column="driving_morning"
  }
  mode=mode
  dataframe[,"n"]*(ifelse(dataframe[,column]<=scale,1,(1/(dataframe[,column]/scale))^2))
}

#apply decay function
merged_set$AccessIndex <- decay_fxn_sensitive(merged_set,"transit")

#get average of AccessIndex col, return to summary dataframe by blockgroup ID
summary_index_mover <- aggregate(AccessIndex ~ Id2,data=merged_set,FUN=sum)

#order
summary_index_mover <- summary_index_mover[order(summary_index_mover$AccessIndex),]

#prep for merge
rownames(summary_index_mover) <- NULL
summary_index_mover$rank.i <- rownames(summary_index_mover)
colnames(summary_index_mover)[colnames(summary_index_mover)=="rank.i"] <- paste("rank",i)

#print summary of AI for validation on if this is working
print(summary(summary_index_mover$AccessIndex))
summary_index_mover$AccessIndex <- NULL

#merge new data to dataframe
summary_index_user <- merge(summary_index_user,summary_index_mover,by="Id2",all.y=TRUE)
print(paste(i,"iteration..."))
}

#store as a new data frame
sensitivity_frame_transit <- summary_index_user 

#get the total changes in rank from threshold of five minutes to 16 minutes
sensitivity_frame_transit$changes <- abs(as.numeric(sensitivity_frame_transit$base_rank) - as.numeric(sensitivity_frame_transit$`rank 9`))
```

###Plot transit changes in rank
```{r}
library(reshape2)
library(ggplot2)
library(gridExtra)
#Prep & order frame
sensitivity_frame_transit <- sensitivity_frame_transit 
sensitivity_frame_transit$base_rank <- as.numeric(sensitivity_frame_transit$base_rank)
sensitivity_frame_transit <- sensitivity_frame_transit[order(sensitivity_frame_transit$base_rank),]

#subsetframe
top_25 <- sensitivity_frame_transit[1:25,]
sa_25_50 <- sensitivity_frame_transit[1:50,]
sa_50_75 <- sensitivity_frame_transit[1:75,]
sa_75_100 <- sensitivity_frame_transit[1:100,]
sa_tail <- sensitivity_frame_transit[456:480,]
sa_tailstart <- sensitivity_frame_transit[406:480,]


#Top 25 Plot
df <- melt(top_25,id=c("Id2","AccessIndex"),measure.vars=c("base_rank","rank 1","rank 2","rank 3","rank 4","rank 5","rank 6","rank 7","rank 8","rank 9"))
df$value <- as.numeric(df$value)
df$variable <- as.factor(df$variable)

Top_25_Plot <- ggplot(df, aes(variable,value,group = Id2, color = as.factor(Id2))) + geom_line() + theme(legend.position="none") + labs(title="top 25") + ylim(1,490)

#Top 50 Plot
df <- melt(sa_25_50,id=c("Id2","AccessIndex"),measure.vars=c("base_rank","rank 1","rank 2","rank 3","rank 4","rank 5","rank 6","rank 7","rank 8","rank 9"))
df$value <- as.numeric(df$value)
df$variable <- as.factor(df$variable)

Top_50_Plot <- ggplot(df, aes(variable,value,group = Id2, color = as.factor(Id2))) + geom_line() + theme(legend.position="none") + labs(title="top 50") + ylim(1,490)

#Top 75 Plot
df <- melt(sa_50_75,id=c("Id2","AccessIndex"),measure.vars=c("base_rank","rank 1","rank 2","rank 3","rank 4","rank 5","rank 6","rank 7","rank 8","rank 9"))
df$value <- as.numeric(df$value)
df$variable <- as.factor(df$variable)

Top_75_Plot <- ggplot(df, aes(variable,value,group = Id2, color = as.factor(Id2))) + geom_line() + theme(legend.position="none") + labs(title="top 75") + ylim(1,490)

#Top 100 Plot
df <- melt(sa_75_100,id=c("Id2","AccessIndex"),measure.vars=c("base_rank","rank 1","rank 2","rank 3","rank 4","rank 5","rank 6","rank 7","rank 8","rank 9"))
df$value <- as.numeric(df$value)
df$variable <- as.factor(df$variable)

Top_100_Plot <- ggplot(df, aes(variable,value,group = Id2, color = as.factor(Id2))) + geom_line() + theme(legend.position="none") + labs(title="top 100") + ylim(1,490)

#Bottom 50
df <- melt(sa_tail,id=c("Id2","AccessIndex"),measure.vars=c("base_rank","rank 1","rank 2","rank 3","rank 4","rank 5","rank 6","rank 7","rank 8","rank 9"))
df$value <- as.numeric(df$value)
df$variable <- as.factor(df$variable)

Tail_Plot <- ggplot(df, aes(variable,value,group = Id2, color = as.factor(Id2))) + geom_line() + theme(legend.position="none") + labs(title="Bottom 50") + ylim(1,490)

#Bottom 25
df <- melt(sa_tailstart,id=c("Id2","AccessIndex"),measure.vars=c("base_rank","rank 1","rank 2","rank 3","rank 4","rank 5","rank 6","rank 7","rank 8","rank 9"))
df$value <- as.numeric(df$value)
df$variable <- as.factor(df$variable)

Bottom25_Plot <- ggplot(df, aes(variable,value,group = Id2, color = as.factor(Id2))) + geom_line() + theme(legend.position="none") + labs(title="Bottom 100") + ylim(1,490)

#grid.arrange(Top_25_Plot,Top_50_Plot,Top_75_Plot,Top_100_Plot,Bottom25_Plot,Tail_Plot,ncol=3)
```

###Driving Exp Function
```{r}
#run overall access index driving
library(scales)
#User specifies query (type)
#type <- c("art") #"type" can include categories or cost thresholds. Pending cost per day calculations from Selam
#cost <- c("free")
type <- NULL
cost <- NULL

#subset dataframe by user selection
user_set <- Make_Subset(reschool_programs,type,cost)

#aggregate dataframe by user selection
agg_set <- Aggregate_Subset(user_set)

#merge n of user selected programs to block distances dataframe
merged_set <- Merge_Set(agg_set)

#calculate access index per location 
decay_fxn_exp_sensitive <- function(dataframe,mode){
  if (mode=="transit") {
    scale = 5+i
    column="transit_morning"
  }
  if (mode=="drive"){
    scale = 5+i
    column="driving_morning"
  }
  mode=mode
  dataframe[,"n"]*exp(-dataframe[,column]/scale)
}
i <- 0
merged_set$AccessIndex <- decay_fxn_exp_sensitive(merged_set,"drive") #why not apply? who knows

#get average of AccessIndex col, return to summary dataframe by blockgroup ID
summary_index_user <- aggregate(AccessIndex ~ Id2,data=merged_set,FUN=sum) #critical that we use sum, not mean, BECAUSE, for some specifications some program lengths will drop

#make base dataframe
summary_index_user <- summary_index_user[order(summary_index_user$AccessIndex),]
rownames(summary_index_user)  <- NULL
summary_index_user$base_rank <- rownames(summary_index_user)

for (i in 1:9){
#generate overall
type <- NULL
cost <- NULL

#subset dataframe by user selection
user_set <- Make_Subset(reschool_programs,type,cost)

#aggregate dataframe by user selection
agg_set <- Aggregate_Subset(user_set)

#merge n of user selected programs to block distances dataframe
merged_set <- Merge_Set(agg_set)

#alter decay function
decay_fxn_exp_sensitive <- function(dataframe,mode){
  if (mode=="transit") {
    scale = 5+i
    column="transit_morning"
  }
  if (mode=="drive"){
    scale = 5+i
    column="driving_morning"
  }
  mode=mode
  dataframe[,"n"]*exp(-dataframe[,column]/scale)
}

#apply decay function
merged_set$AccessIndex <- decay_fxn_exp_sensitive(merged_set,"drive")

#get average of AccessIndex col, return to summary dataframe by blockgroup ID
summary_index_mover <- aggregate(AccessIndex ~ Id2,data=merged_set,FUN=sum)

#order
summary_index_mover <- summary_index_mover[order(summary_index_mover$AccessIndex),]

#prep for merge
rownames(summary_index_mover) <- NULL
summary_index_mover$rank.i <- rownames(summary_index_mover)
colnames(summary_index_mover)[colnames(summary_index_mover)=="rank.i"] <- paste("rank",i)

#print summary of AI for validation on if this is working
print(summary(summary_index_mover$AccessIndex))
summary_index_mover$AccessIndex <- NULL

#merge new data to dataframe
summary_index_user <- merge(summary_index_user,summary_index_mover,by="Id2",all.y=TRUE)
print(paste(i,"iteration..."))
}

#store as a new data frame
sensitivity_frame_exp <- summary_index_user 

#get the total changes in rank from threshold of five minutes to 16 minutes
sensitivity_frame_exp$changes <- abs(as.numeric(sensitivity_frame_exp$base_rank) - as.numeric(sensitivity_frame_exp$`rank 9`))
```

###Plot Exp 
```{r}
library(reshape2)
library(ggplot2)
library(gridExtra)
#Prep & order frame
sensitivity_frame_exp <- sensitivity_frame_exp 
sensitivity_frame_exp$base_rank <- as.numeric(sensitivity_frame_exp$base_rank)
sensitivity_frame_exp <- sensitivity_frame_exp[order(sensitivity_frame_exp$base_rank),]

#subsetframe
top_25 <- sensitivity_frame_exp[1:25,]
sa_25_50 <- sensitivity_frame_exp[1:50,]
sa_50_75 <- sensitivity_frame_exp[1:75,]
sa_75_100 <- sensitivity_frame_exp[1:100,]
sa_tail <- sensitivity_frame_exp[456:480,]
sa_tailstart <- sensitivity_frame_exp[406:480,]


#Top 25 Plot
df <- melt(top_25,id=c("Id2","AccessIndex"),measure.vars=c("base_rank","rank 1","rank 2","rank 3","rank 4","rank 5","rank 6","rank 7","rank 8","rank 9"))
df$value <- as.numeric(df$value)
df$variable <- as.factor(df$variable)

driving_exp_Top_25_Plot <- ggplot(df, aes(variable,value,group = Id2, color = as.factor(Id2))) + geom_line() + theme(legend.position="none") + labs(title="top 25") + ylim(1,490)

#Top 50 Plot
df <- melt(sa_25_50,id=c("Id2","AccessIndex"),measure.vars=c("base_rank","rank 1","rank 2","rank 3","rank 4","rank 5","rank 6","rank 7","rank 8","rank 9"))
df$value <- as.numeric(df$value)
df$variable <- as.factor(df$variable)

driving_exp_Top_50_Plot <- ggplot(df, aes(variable,value,group = Id2, color = as.factor(Id2))) + geom_line() + theme(legend.position="none") + labs(title="top 50") + ylim(1,490)

#Top 75 Plot
df <- melt(sa_50_75,id=c("Id2","AccessIndex"),measure.vars=c("base_rank","rank 1","rank 2","rank 3","rank 4","rank 5","rank 6","rank 7","rank 8","rank 9"))
df$value <- as.numeric(df$value)
df$variable <- as.factor(df$variable)

driving_exp_Top_75_Plot <- ggplot(df, aes(variable,value,group = Id2, color = as.factor(Id2))) + geom_line() + theme(legend.position="none") + labs(title="top 75") + ylim(1,490)

#Top 100 Plot
df <- melt(sa_75_100,id=c("Id2","AccessIndex"),measure.vars=c("base_rank","rank 1","rank 2","rank 3","rank 4","rank 5","rank 6","rank 7","rank 8","rank 9"))
df$value <- as.numeric(df$value)
df$variable <- as.factor(df$variable)

driving_exp_Top_100_Plot <- ggplot(df, aes(variable,value,group = Id2, color = as.factor(Id2))) + geom_line() + theme(legend.position="none") + labs(title="top 100") + ylim(1,490)

#Bottom 50
df <- melt(sa_tail,id=c("Id2","AccessIndex"),measure.vars=c("base_rank","rank 1","rank 2","rank 3","rank 4","rank 5","rank 6","rank 7","rank 8","rank 9"))
df$value <- as.numeric(df$value)
df$variable <- as.factor(df$variable)

driving_exp_Tail_Plot <- ggplot(df, aes(variable,value,group = Id2, color = as.factor(Id2))) + geom_line() + theme(legend.position="none") + labs(title="Bottom 50") + ylim(1,490)

#Bottom 25
df <- melt(sa_tailstart,id=c("Id2","AccessIndex"),measure.vars=c("base_rank","rank 1","rank 2","rank 3","rank 4","rank 5","rank 6","rank 7","rank 8","rank 9"))
df$value <- as.numeric(df$value)
df$variable <- as.factor(df$variable)

driving_exp_Bottom25_Plot <- ggplot(df, aes(variable,value,group = Id2, color = as.factor(Id2))) + geom_line() + theme(legend.position="none") + labs(title="Bottom 100") + ylim(1,490)

grid.arrange(driving_exp_Top_25_Plot,driving_exp_Top_50_Plot,driving_exp_Top_75_Plot,driving_exp_Top_100_Plot,driving_exp_Bottom25_Plot,driving_exp_Tail_Plot,ncol=3)
```

###Compare Driving Exp and Base
Here we plot the difference between the exp and base models, ideally testing sensitivity between models. We find that at almost every iteration, differences are clustered primarily in the center ranks, with very limited changes along the tails, which are the areas we are primarily concerned with for an endogenous index. **GIVEN** that exp model is less sensitive to the threshold, and there seems to be marginal sensitivity between models, seems like we should go with the exp model over the threshold model, since the results will be less biased by a selection of an arbitrary cutoff.   

```{r}
compare_exp_base <- merge(sensitivity_frame,sensitivity_frame_exp,by="Id2",all.y=TRUE)
compare_exp_base$changes <- NULL
compare_exp_base$AccessIndex.x <- NULL

compare_exp_base$base_rank.x <- as.numeric(compare_exp_base$base_rank.x)
compare_exp_base$base_rank.y <- as.numeric(compare_exp_base$base_rank.y)

compare_exp_base$`rank 1.x` <- as.numeric(compare_exp_base$`rank 1.x`)
compare_exp_base$`rank 1.y` <- as.numeric(compare_exp_base$`rank 1.y`)

compare_exp_base$`rank 2.x` <- as.numeric(compare_exp_base$`rank 2.x`)
compare_exp_base$`rank 2.y` <- as.numeric(compare_exp_base$`rank 2.y`)

compare_exp_base$`rank 3.x` <- as.numeric(compare_exp_base$`rank 3.x`)
compare_exp_base$`rank 3.y` <- as.numeric(compare_exp_base$`rank 3.y`)

compare_exp_base$`rank 4.x` <- as.numeric(compare_exp_base$`rank 4.x`)
compare_exp_base$`rank 4.y` <- as.numeric(compare_exp_base$`rank 4.y`)

compare_exp_base$`rank 5.x` <- as.numeric(compare_exp_base$`rank 5.x`)
compare_exp_base$`rank 5.y` <- as.numeric(compare_exp_base$`rank 5.y`)

compare_exp_base$`rank 6.x` <- as.numeric(compare_exp_base$`rank 6.x`)
compare_exp_base$`rank 6.y` <- as.numeric(compare_exp_base$`rank 6.y`)

compare_exp_base$`rank 7.x` <- as.numeric(compare_exp_base$`rank 7.x`)
compare_exp_base$`rank 7.y` <- as.numeric(compare_exp_base$`rank 7.y`)

compare_exp_base$`rank 8.x` <- as.numeric(compare_exp_base$`rank 8.x`)
compare_exp_base$`rank 8.y` <- as.numeric(compare_exp_base$`rank 8.y`)

compare_exp_base$`rank 9.x` <- as.numeric(compare_exp_base$`rank 9.x`)
compare_exp_base$`rank 9.y` <- as.numeric(compare_exp_base$`rank 9.y`)

compare_exp_base$`rank 1.x` <- as.numeric(compare_exp_base$`rank 1.x`)
compare_exp_base$`rank 1.y` <- as.numeric(compare_exp_base$`rank 1.y`)

comp_one <- ggplot(compare_exp_base, aes(x=compare_exp_base$base_rank.x, y=compare_exp_base$base_rank.y)) + geom_point()

comp_two <-ggplot(compare_exp_base, aes(x=compare_exp_base$`rank 1.x`, y=compare_exp_base$`rank 1.y`)) + geom_point()

comp_three <-ggplot(compare_exp_base, aes(x=compare_exp_base$`rank 2.x`, y=compare_exp_base$`rank 2.y`)) + geom_point()

comp_four <-ggplot(compare_exp_base, aes(x=compare_exp_base$`rank 3.x`, y=compare_exp_base$`rank 3.y`)) + geom_point()

comp_five <-ggplot(compare_exp_base, aes(x=compare_exp_base$`rank 4.x`, y=compare_exp_base$`rank 4.y`)) + geom_point()

comp_six <-ggplot(compare_exp_base, aes(x=compare_exp_base$`rank 5.x`, y=compare_exp_base$`rank 5.y`)) + geom_point()

comp_seven <-ggplot(compare_exp_base, aes(x=compare_exp_base$`rank 6.x`, y=compare_exp_base$`rank 6.y`)) + geom_point()

comp_eight <-ggplot(compare_exp_base, aes(x=compare_exp_base$`rank 7.x`, y=compare_exp_base$`rank 7.y`)) + geom_point()

comp_nine <-ggplot(compare_exp_base, aes(x=compare_exp_base$`rank 8.x`, y=compare_exp_base$`rank 8.y`)) + geom_point()

comp_ten <-ggplot(compare_exp_base, aes(x=compare_exp_base$`rank 9.x`, y=compare_exp_base$`rank 9.y`)) + geom_point()

#grid.arrange(comp_one,comp_two,comp_three,comp_four,comp_five,comp_six,comp_seven,comp_eight,comp_nine,comp_ten,ncol=4)
```

###Transit Exp Model
```{r}
#run overall access index driving
library(scales)
#User specifies query (type)
type <- NULL
cost <- NULL

#subset dataframe by user selection
user_set <- Make_Subset(reschool_programs,type,cost)

#aggregate dataframe by user selection
agg_set <- Aggregate_Subset(user_set)

#merge n of user selected programs to block distances dataframe
merged_set <- Merge_Set(agg_set)

#calculate access index per location 
i <- 0

merged_set$AccessIndex <- decay_fxn_exp_sensitive(merged_set,"transit")

#get average of AccessIndex col, return to summary dataframe by blockgroup ID
summary_index_user <- aggregate(AccessIndex ~ Id2,data=merged_set,FUN=sum) #critical that we use sum, not mean, BECAUSE, for some specifications some program lengths will drop

#make base dataframe
summary_index_user <- summary_index_user[order(summary_index_user$AccessIndex),]
rownames(summary_index_user)  <- NULL
summary_index_user$base_rank <- rownames(summary_index_user)

for (i in 1:9){
#generate overall
type <- NULL
cost <- NULL

#subset dataframe by user selection
user_set <- Make_Subset(reschool_programs,type,cost)

#aggregate dataframe by user selection
agg_set <- Aggregate_Subset(user_set)

#merge n of user selected programs to block distances dataframe
merged_set <- Merge_Set(agg_set)

#alter decay function
decay_fxn_exp_sensitive <- function(dataframe,mode){
  if (mode=="transit") {
    scale = 5+i
    column="transit_morning"
  }
  if (mode=="drive"){
    scale = 5+i
    column="driving_morning"
  }
  mode=mode
  dataframe[,"n"]*exp(-dataframe[,column]/scale)
}

#apply decay function
merged_set$AccessIndex <- decay_fxn_exp_sensitive(merged_set,"transit")

#get average of AccessIndex col, return to summary dataframe by blockgroup ID
summary_index_mover <- aggregate(AccessIndex ~ Id2,data=merged_set,FUN=sum)

#order
summary_index_mover <- summary_index_mover[order(summary_index_mover$AccessIndex),]

#prep for merge
rownames(summary_index_mover) <- NULL
summary_index_mover$rank.i <- rownames(summary_index_mover)
colnames(summary_index_mover)[colnames(summary_index_mover)=="rank.i"] <- paste("rank",i)

#print summary of AI for validation on if this is working
print(summary(summary_index_mover$AccessIndex))
summary_index_mover$AccessIndex <- NULL

#merge new data to dataframe
summary_index_user <- merge(summary_index_user,summary_index_mover,by="Id2",all.y=TRUE)
print(paste(i,"iteration..."))
}

#store as a new data frame
sensitivity_frame_exp_transit <- summary_index_user

#get the total changes in rank from threshold of five minutes to 16 minutes
sensitivity_frame_exp_transit$changes <- abs(as.numeric(sensitivity_frame_exp_transit$base_rank) - as.numeric(sensitivity_frame_exp_transit$`rank 9`))
```

###Plot Transit Exp 
```{r}
library(reshape2)
library(ggplot2)
library(gridExtra)
#Prep & order frame
sensitivity_frame_exp_transit <- sensitivity_frame_exp_transit 
sensitivity_frame_exp_transit$base_rank <- as.numeric(sensitivity_frame_exp_transit$base_rank)
sensitivity_frame_exp_transit <- sensitivity_frame_exp_transit[order(sensitivity_frame_exp_transit$base_rank),]

#subsetframe
top_25 <- sensitivity_frame_exp_transit[1:25,]
sa_25_50 <- sensitivity_frame_exp_transit[1:50,]
sa_50_75 <- sensitivity_frame_exp_transit[1:75,]
sa_75_100 <- sensitivity_frame_exp_transit[1:100,]
sa_tail <- sensitivity_frame_exp_transit[456:480,]
sa_tailstart <- sensitivity_frame_exp_transit[406:480,]


#Top 25 Plot
df <- melt(top_25,id=c("Id2","AccessIndex"),measure.vars=c("base_rank","rank 1","rank 2","rank 3","rank 4","rank 5","rank 6","rank 7","rank 8","rank 9"))
df$value <- as.numeric(df$value)
df$variable <- as.factor(df$variable)

transit_exp_Top_25_Plot <- ggplot(df, aes(variable,value,group = Id2, color = as.factor(Id2))) + geom_line() + theme(legend.position="none") + labs(title="top 25") + ylim(1,490)

#Top 50 Plot
df <- melt(sa_25_50,id=c("Id2","AccessIndex"),measure.vars=c("base_rank","rank 1","rank 2","rank 3","rank 4","rank 5","rank 6","rank 7","rank 8","rank 9"))
df$value <- as.numeric(df$value)
df$variable <- as.factor(df$variable)

transit_exp_Top_50_Plot <- ggplot(df, aes(variable,value,group = Id2, color = as.factor(Id2))) + geom_line() + theme(legend.position="none") + labs(title="top 50") + ylim(1,490)

#Top 75 Plot
df <- melt(sa_50_75,id=c("Id2","AccessIndex"),measure.vars=c("base_rank","rank 1","rank 2","rank 3","rank 4","rank 5","rank 6","rank 7","rank 8","rank 9"))
df$value <- as.numeric(df$value)
df$variable <- as.factor(df$variable)

transit_exp_Top_75_Plot <- ggplot(df, aes(variable,value,group = Id2, color = as.factor(Id2))) + geom_line() + theme(legend.position="none") + labs(title="top 75") + ylim(1,490)

#Top 100 Plot
df <- melt(sa_75_100,id=c("Id2","AccessIndex"),measure.vars=c("base_rank","rank 1","rank 2","rank 3","rank 4","rank 5","rank 6","rank 7","rank 8","rank 9"))
df$value <- as.numeric(df$value)
df$variable <- as.factor(df$variable)

transit_exp_Top_100_Plot <- ggplot(df, aes(variable,value,group = Id2, color = as.factor(Id2))) + geom_line() + theme(legend.position="none") + labs(title="top 100") + ylim(1,490)

#Bottom 50
df <- melt(sa_tail,id=c("Id2","AccessIndex"),measure.vars=c("base_rank","rank 1","rank 2","rank 3","rank 4","rank 5","rank 6","rank 7","rank 8","rank 9"))
df$value <- as.numeric(df$value)
df$variable <- as.factor(df$variable)

transit_exp_Tail_Plot <- ggplot(df, aes(variable,value,group = Id2, color = as.factor(Id2))) + geom_line() + theme(legend.position="none") + labs(title="Bottom 50") + ylim(1,490)

#Bottom 25
df <- melt(sa_tailstart,id=c("Id2","AccessIndex"),measure.vars=c("base_rank","rank 1","rank 2","rank 3","rank 4","rank 5","rank 6","rank 7","rank 8","rank 9"))
df$value <- as.numeric(df$value)
df$variable <- as.factor(df$variable)

transit_exp_Bottom25_Plot <- ggplot(df, aes(variable,value,group = Id2, color = as.factor(Id2))) + geom_line() + theme(legend.position="none") + labs(title="Bottom 100") + ylim(1,490)

grid.arrange(transit_exp_Top_25_Plot,transit_exp_Top_50_Plot,transit_exp_Top_75_Plot,transit_exp_Top_100_Plot,transit_exp_Bottom25_Plot,transit_exp_Tail_Plot,ncol=3)
```

###Compare Transit Exp and Transit Gravity Threshold
```{r}
compare_exp_base_transit <- merge(sensitivity_frame_transit,sensitivity_frame_exp_transit,by="Id2",all.y=TRUE)
compare_exp_base_transit$changes <- NULL
compare_exp_base_transit$AccessIndex.x <- NULL

compare_exp_base_transit$base_rank.x <- as.numeric(compare_exp_base_transit$base_rank.x)
compare_exp_base_transit$base_rank.y <- as.numeric(compare_exp_base_transit$base_rank.y)

compare_exp_base_transit$`rank 1.x` <- as.numeric(compare_exp_base_transit$`rank 1.x`)
compare_exp_base_transit$`rank 1.y` <- as.numeric(compare_exp_base_transit$`rank 1.y`)

compare_exp_base_transit$`rank 2.x` <- as.numeric(compare_exp_base_transit$`rank 2.x`)
compare_exp_base_transit$`rank 2.y` <- as.numeric(compare_exp_base_transit$`rank 2.y`)

compare_exp_base_transit$`rank 3.x` <- as.numeric(compare_exp_base_transit$`rank 3.x`)
compare_exp_base_transit$`rank 3.y` <- as.numeric(compare_exp_base_transit$`rank 3.y`)

compare_exp_base_transit$`rank 4.x` <- as.numeric(compare_exp_base_transit$`rank 4.x`)
compare_exp_base_transit$`rank 4.y` <- as.numeric(compare_exp_base_transit$`rank 4.y`)

compare_exp_base_transit$`rank 5.x` <- as.numeric(compare_exp_base_transit$`rank 5.x`)
compare_exp_base_transit$`rank 5.y` <- as.numeric(compare_exp_base_transit$`rank 5.y`)

compare_exp_base_transit$`rank 6.x` <- as.numeric(compare_exp_base_transit$`rank 6.x`)
compare_exp_base_transit$`rank 6.y` <- as.numeric(compare_exp_base_transit$`rank 6.y`)

compare_exp_base_transit$`rank 7.x` <- as.numeric(compare_exp_base_transit$`rank 7.x`)
compare_exp_base_transit$`rank 7.y` <- as.numeric(compare_exp_base_transit$`rank 7.y`)

compare_exp_base_transit$`rank 8.x` <- as.numeric(compare_exp_base_transit$`rank 8.x`)
compare_exp_base_transit$`rank 8.y` <- as.numeric(compare_exp_base_transit$`rank 8.y`)

compare_exp_base_transit$`rank 9.x` <- as.numeric(compare_exp_base_transit$`rank 9.x`)
compare_exp_base_transit$`rank 9.y` <- as.numeric(compare_exp_base_transit$`rank 9.y`)

compare_exp_base_transit$`rank 1.x` <- as.numeric(compare_exp_base_transit$`rank 1.x`)
compare_exp_base_transit$`rank 1.y` <- as.numeric(compare_exp_base_transit$`rank 1.y`)

comp_transit_one <- ggplot(compare_exp_base_transit, aes(x=compare_exp_base_transit$base_rank.x, y=compare_exp_base_transit$base_rank.y)) + geom_point()

comp_transit_two <-ggplot(compare_exp_base_transit, aes(x=compare_exp_base_transit$`rank 1.x`, y=compare_exp_base_transit$`rank 1.y`)) + geom_point()

comp_transit_three <-ggplot(compare_exp_base_transit, aes(x=compare_exp_base_transit$`rank 2.x`, y=compare_exp_base_transit$`rank 2.y`)) + geom_point()

comp_transit_four <-ggplot(compare_exp_base_transit, aes(x=compare_exp_base_transit$`rank 3.x`, y=compare_exp_base_transit$`rank 3.y`)) + geom_point()

comp_transit_five <-ggplot(compare_exp_base_transit, aes(x=compare_exp_base_transit$`rank 4.x`, y=compare_exp_base_transit$`rank 4.y`)) + geom_point()

comp_transit_six <-ggplot(compare_exp_base_transit, aes(x=compare_exp_base_transit$`rank 5.x`, y=compare_exp_base_transit$`rank 5.y`)) + geom_point()

comp_transit_seven <-ggplot(compare_exp_base_transit, aes(x=compare_exp_base_transit$`rank 6.x`, y=compare_exp_base_transit$`rank 6.y`)) + geom_point()

comp_transit_eight <-ggplot(compare_exp_base_transit, aes(x=compare_exp_base_transit$`rank 7.x`, y=compare_exp_base_transit$`rank 7.y`)) + geom_point()

comp_transit_nine <-ggplot(compare_exp_base_transit, aes(x=compare_exp_base_transit$`rank 8.x`, y=compare_exp_base_transit$`rank 8.y`)) + geom_point()

comp_transit_ten <-ggplot(compare_exp_base_transit, aes(x=compare_exp_base_transit$`rank 9.x`, y=compare_exp_base_transit$`rank 9.y`)) + geom_point()

#grid.arrange(comp_transit_one,comp_transit_two,comp_transit_three,comp_transit_four,comp_transit_five,comp_transit_six,comp_transit_seven,comp_transit_eight,comp_transit_nine,comp_transit_ten,ncol=4)
```

###Gravity Model
At the end of the day, scale in this case in a constant. However, it's fucked for dealing with close proximity tons.
```{r}
#run overall access index driving
library(scales)
#User specifies query (type)
#type <- c("art") #"type" can include categories or cost thresholds. Pending cost per day calculations from Selam
#cost <- c("free")
type <- NULL
cost <- NULL

#subset dataframe by user selection
user_set <- Make_Subset(reschool_programs,type,cost)

#aggregate dataframe by user selection
agg_set <- Aggregate_Subset(user_set)

#merge n of user selected programs to block distances dataframe
merged_set <- Merge_Set(agg_set)

#calculate access index per location 
i <- 0
merged_set$AccessIndex <- decay_fxn_sensitive_gravity(merged_set,"drive") #why not apply? who knows

#get average of AccessIndex col, return to summary dataframe by blockgroup ID
summary_index_user <- aggregate(AccessIndex ~ Id2,data=merged_set,FUN=sum) #critical that we use sum, not mean, BECAUSE, for some specifications some program lengths will drop

#make base dataframe
summary_index_user <- summary_index_user[order(summary_index_user$AccessIndex),]
rownames(summary_index_user)  <- NULL
summary_index_user$base_rank <- rownames(summary_index_user)

for (i in 1:9){
#generate overall
type <- NULL
cost <- NULL

#subset dataframe by user selection
user_set <- Make_Subset(reschool_programs,type,cost)

#aggregate dataframe by user selection
agg_set <- Aggregate_Subset(user_set)

#merge n of user selected programs to block distances dataframe
merged_set <- Merge_Set(agg_set)

#alter decay function
decay_fxn_sensitive_gravity <- function(dataframe,mode){
  if (mode=="transit") {
    scale = 5+i
    column="transit_morning"
  }
  if (mode=="drive"){
    scale = 5+i
    column="driving_morning"
  }
  mode=mode
  dataframe[,"n"]*((1/(dataframe[,column]/scale))^2)
}

#apply decay function
merged_set$AccessIndex <- decay_fxn_sensitive_gravity(merged_set,"drive")

#get average of AccessIndex col, return to summary dataframe by blockgroup ID
summary_index_mover <- aggregate(AccessIndex ~ Id2,data=merged_set,FUN=sum)

#order
summary_index_mover <- summary_index_mover[order(summary_index_mover$AccessIndex),]

#prep for merge
rownames(summary_index_mover) <- NULL
summary_index_mover$rank.i <- rownames(summary_index_mover)
colnames(summary_index_mover)[colnames(summary_index_mover)=="rank.i"] <- paste("rank",i)

#print summary of AI for validation on if this is working
print(summary(summary_index_mover$AccessIndex))
summary_index_mover$AccessIndex <- NULL

#merge new data to dataframe
summary_index_user <- merge(summary_index_user,summary_index_mover,by="Id2",all.y=TRUE)
print(paste(i,"iteration..."))
}

#store as a new data frame
sensitivity_frame_gravity <- summary_index_user 

#get the total changes in rank from threshold of five minutes to 16 minutes
sensitivity_frame_gravity$changes <- abs(as.numeric(sensitivity_frame_gravity$base_rank) - as.numeric(sensitivity_frame_gravity$`rank 9`))
```

###Plot Gravity
```{r}
library(reshape2)
library(ggplot2)
library(gridExtra)
#Prep & order frame
sensitivity_frame_gravity <- sensitivity_frame_gravity 
sensitivity_frame_gravity$base_rank <- as.numeric(sensitivity_frame_gravity$base_rank)
sensitivity_frame_gravity <- sensitivity_frame_gravity[order(sensitivity_frame_gravity$base_rank),]

#subsetframe
top_25 <- sensitivity_frame_gravity[1:25,]
sa_25_50 <- sensitivity_frame_gravity[1:50,]
sa_50_75 <- sensitivity_frame_gravity[1:75,]
sa_75_100 <- sensitivity_frame_gravity[1:100,]
sa_tail <- sensitivity_frame_gravity[456:480,]
sa_tailstart <- sensitivity_frame_gravity[406:480,]


#Top 25 Plot
df <- melt(top_25,id=c("Id2","AccessIndex"),measure.vars=c("base_rank","rank 1","rank 2","rank 3","rank 4","rank 5","rank 6","rank 7","rank 8","rank 9"))
df$value <- as.numeric(df$value)
df$variable <- as.factor(df$variable)

driving_gravity_Top_25_Plot <- ggplot(df, aes(variable,value,group = Id2, color = as.factor(Id2))) + geom_line() + theme(legend.position="none") + labs(title="top 25") + ylim(1,490)

#Top 50 Plot
df <- melt(sa_25_50,id=c("Id2","AccessIndex"),measure.vars=c("base_rank","rank 1","rank 2","rank 3","rank 4","rank 5","rank 6","rank 7","rank 8","rank 9"))
df$value <- as.numeric(df$value)
df$variable <- as.factor(df$variable)

driving_gravity_Top_50_Plot <- ggplot(df, aes(variable,value,group = Id2, color = as.factor(Id2))) + geom_line() + theme(legend.position="none") + labs(title="top 50") + ylim(1,490)

#Top 75 Plot
df <- melt(sa_50_75,id=c("Id2","AccessIndex"),measure.vars=c("base_rank","rank 1","rank 2","rank 3","rank 4","rank 5","rank 6","rank 7","rank 8","rank 9"))
df$value <- as.numeric(df$value)
df$variable <- as.factor(df$variable)

driving_gravity_Top_75_Plot <- ggplot(df, aes(variable,value,group = Id2, color = as.factor(Id2))) + geom_line() + theme(legend.position="none") + labs(title="top 75") + ylim(1,490)

#Top 100 Plot
df <- melt(sa_75_100,id=c("Id2","AccessIndex"),measure.vars=c("base_rank","rank 1","rank 2","rank 3","rank 4","rank 5","rank 6","rank 7","rank 8","rank 9"))
df$value <- as.numeric(df$value)
df$variable <- as.factor(df$variable)

driving_gravity_Top_100_Plot <- ggplot(df, aes(variable,value,group = Id2, color = as.factor(Id2))) + geom_line() + theme(legend.position="none") + labs(title="top 100") + ylim(1,490)

#Bottom 50
df <- melt(sa_tail,id=c("Id2","AccessIndex"),measure.vars=c("base_rank","rank 1","rank 2","rank 3","rank 4","rank 5","rank 6","rank 7","rank 8","rank 9"))
df$value <- as.numeric(df$value)
df$variable <- as.factor(df$variable)

driving_gravity_Tail_Plot <- ggplot(df, aes(variable,value,group = Id2, color = as.factor(Id2))) + geom_line() + theme(legend.position="none") + labs(title="Bottom 50") + ylim(1,490)

#Bottom 25
df <- melt(sa_tailstart,id=c("Id2","AccessIndex"),measure.vars=c("base_rank","rank 1","rank 2","rank 3","rank 4","rank 5","rank 6","rank 7","rank 8","rank 9"))
df$value <- as.numeric(df$value)
df$variable <- as.factor(df$variable)

driving_gravity_Bottom25_Plot <- ggplot(df, aes(variable,value,group = Id2, color = as.factor(Id2))) + geom_line() + theme(legend.position="none") + labs(title="Bottom 100") + ylim(1,490)

grid.arrange(driving_gravity_Top_25_Plot,driving_gravity_Top_50_Plot,driving_gravity_Top_75_Plot,driving_gravity_Top_100_Plot,driving_gravity_Bottom25_Plot,driving_gravity_Tail_Plot,ncol=3)
```

###Plot Gravity Blocks
```{r}
library(leaflet)
library(RColorBrewer)
test <- sensitivity_frame_gravity
test$AccessIndex <- normalize(test$AccessIndex,max = max(test$AccessIndex))
test$Id2 <- as.numeric(test$Id2)
test <- test[order(test$Id2),]

#Ordering
census_blocks$Id2 <- as.numeric(census_blocks$Id2)
merged_set$Id2 <- as.numeric(merged_set$Id2)
summary_index_user$Id2 <- as.numeric(summary_index_user$Id2)

census_blocks <- census_blocks[order(census_blocks$Id2),]
merged_set <- merged_set[order(merged_set$Id2),]
summary_index_user <- summary_index_user[order(summary_index_user$Id2),] 

#specifiying color pallet
colourCount <- length(unique(agg_set$n))
getPalette <- colorRampPalette(brewer.pal(9,"Reds"))

pal <- colorNumeric(
  palette = "Blues",
  domain =  0:1)

pal2 <- colorNumeric(
  palette = rev(getPalette(colourCount)),
  domain = range(agg_set$n))

#indicating one popup info
info <- paste("n",agg_set$n)

#creating leaflet map
Ai_driving_gravity <- leaflet(census_blocks) %>%
  addTiles() %>%
  addPolygons(color = "Blues",weight =1,smoothFactor = 0.5,opacity = 0.5,fillOpacity = 0.75,
              fillColor = ~pal(test$AccessIndex),
              popup=paste(as.numeric(summary_index_user$Id2),as.numeric(test$AccessIndex),as.numeric(census_blocks$Id2)),
              highlightOptions = highlightOptions(color = "white", weight = 0.1)) %>%
  
  addLegend(pal = pal, values = ~test$AccessIndex, opacity = 1,title = "Driving Gravity AI Score") %>%
  
  addCircles(data = agg_set, 
             opacity = 1,weight=7, popup=info,
             color = ~pal2(agg_set$n) ) %>%
  
  setView(lat=39.7,lng=-104.9,zoom=11.5)
```

###Plot Exp Blocks
```{r}
library(leaflet)
library(RColorBrewer)
test <- sensitivity_frame_exp
test$AccessIndex <- normalize(test$AccessIndex,max = max(test$AccessIndex))
test$Id2 <- as.numeric(test$Id2)
test <- test[order(test$Id2),]

#Ordering
census_blocks$Id2 <- as.numeric(census_blocks$Id2)
merged_set$Id2 <- as.numeric(merged_set$Id2)
summary_index_user$Id2 <- as.numeric(summary_index_user$Id2)

census_blocks <- census_blocks[order(census_blocks$Id2),]
merged_set <- merged_set[order(merged_set$Id2),]
summary_index_user <- summary_index_user[order(summary_index_user$Id2),] 

#specifiying color pallet
colourCount <- length(unique(agg_set$n))
getPalette <- colorRampPalette(brewer.pal(9,"Reds"))

pal <- colorNumeric(
  palette = "Blues",
  domain =  0:1)

pal2 <- colorNumeric(
  palette = rev(getPalette(colourCount)),
  domain = range(agg_set$n))

#indicating one popup info
info <- paste("n",agg_set$n)

#creating leaflet map
Ai_driving_exp <- leaflet(census_blocks) %>%
  addTiles() %>%
  addPolygons(color = "Blues",weight =1,smoothFactor = 0.5,opacity = 0.5,fillOpacity = 0.75,
              fillColor = ~pal(test$AccessIndex),
              popup=paste(as.numeric(summary_index_user$Id2),as.numeric(test$AccessIndex),as.numeric(census_blocks$Id2)),
              highlightOptions = highlightOptions(color = "white", weight = 0.1)) %>%
  
  addLegend(pal = pal, values = ~test$AccessIndex, opacity = 1,title = "Driving Exp AI Score") %>%
  
  addCircles(data = agg_set, 
             opacity = 1,weight=7, popup=info,
             color = ~pal2(agg_set$n) ) %>%
  
  setView(lat=39.7,lng=-104.9,zoom=11.5)
```

###Plot Driving Base
```{r}
library(leaflet)
library(RColorBrewer)
type <- NULL
cost <- NULL

#subset dataframe by user selection
user_set <- Make_Subset(reschool_programs,type,cost)

#aggregate dataframe by user selection
agg_set <- Aggregate_Subset(user_set)

#merge n of user selected programs to block distances dataframe
merged_set <- Merge_Set(agg_set)

#calculate access index per location 
merged_set$AccessIndex <- decay_fxn(merged_set,"drive") #why not apply? who knows

#get average of AccessIndex col, return to summary dataframe by blockgroup ID
summary_index_user <- aggregate(AccessIndex ~ Id2,data=merged_set,FUN=sum) #critical that

#normalize
max_access <- max(summary_index_user$AccessIndex)
summary_index_user$AccessIndex <- normalize(summary_index_user$AccessIndex,max_access)

#Ordering
census_blocks$Id2 <- as.numeric(census_blocks$Id2)
merged_set$Id2 <- as.numeric(merged_set$Id2)
summary_index_user$Id2 <- as.numeric(summary_index_user$Id2)

census_blocks <- census_blocks[order(census_blocks$Id2),]
merged_set <- merged_set[order(merged_set$Id2),]
summary_index_user <- summary_index_user[order(summary_index_user$Id2),] 

#specifiying color pallet
colourCount <- length(unique(agg_set$n))
getPalette <- colorRampPalette(brewer.pal(9,"Reds"))

pal <- colorNumeric(
  palette = "Blues",
  domain =  0:1)

pal2 <- colorNumeric(
  palette = rev(getPalette(colourCount)),
  domain = range(agg_set$n))

#indicating one popup info
info <- paste("n",agg_set$n)

#creating leaflet map
Ai_driving <- leaflet(census_blocks) %>%
  addTiles() %>%
  addPolygons(color = "Blues",weight =1,smoothFactor = 0.5,opacity = 0.5,fillOpacity = 0.75,
              fillColor = ~pal(summary_index_user$AccessIndex),
              popup=paste(as.numeric(summary_index_user$Id2),summary_index_user$AccessIndex,as.numeric(census_blocks$Id2)),
              highlightOptions = highlightOptions(color = "white", weight = 0.1)) %>%
  
  addLegend(pal = pal, values = ~summary_index_user$AccessIndex, opacity = 1,title = "Driving AI Score") %>%
  
  addCircles(data = agg_set, 
             opacity = 1,weight=7, popup=info,
             color = ~pal2(agg_set$n) ) %>%
  
  setView(lat=39.7,lng=-104.9,zoom=11.5)
```

##Exp vs Base Blocks
The main difference between these is that block groups close to addresses with many programs are more heavily weighted. Hence, we have a trade off between the base model which is more sensitive to the arbitrary threshold chosen, vs. the exp model which is more sensitivte to addresses with many sessions. Because the latter is founded on the data and conceptually valid, while the former introduces arbitrary indifference assumption (n of minutes) when we have already aggregated starting points to an arbitrary spot (centroids), the exp model seems more valid to me. 

There is a large problem here, which is that the transit times are more volatile in response to the exponential model, while the driving times are far more volatile in response to the gravity threshold model with respect to the choice of scale. As the "scale" is an assumption we make with respect to decay, selecting a model who's results vary depending on our assumption risk reporting results which are biased by our choice of our assumption over the actual travel times in the data.

To recifty this problem, we also test a "soft gravity" model which adds an additional minute to each travel time as approx average offset from the distance of depature address to block group centroid.


###"Soft" Gravity Model Driving
```{r}
#run overall access index driving
library(scales)
#User specifies query (type), NULL = all
type <- NULL
cost <- NULL
i <- 0 

#subset dataframe by user selection
user_set <- Make_Subset(reschool_programs,type,cost)

#aggregate dataframe by user selection
agg_set <- Aggregate_Subset(user_set)

#merge n of user selected programs to block distances dataframe
merged_set <- Merge_Set(agg_set)

#soft gravity function
decay_fxn_softGravity <- function(dataframe,mode){
  if (mode=="transit") {
    scale = 5+i
    column="transit_morning"
  }
  if (mode=="drive"){
    scale = 5+i
    column="driving_morning"
  }
  mode=mode
  dataframe[,"n"]*(1/(1+(dataframe[,column]/scale))^2)
}

#calculate access index per location 
merged_set$AccessIndex <- decay_fxn_softGravity(merged_set,"drive") #why not apply? who knows

#get average of AccessIndex col, return to summary dataframe by blockgroup ID
summary_index_user <- aggregate(AccessIndex ~ Id2,data=merged_set,FUN=sum) #critical that we use sum, not mean, BECAUSE, for some specifications some program lengths will drop

#make base dataframe
summary_index_user <- summary_index_user[order(summary_index_user$AccessIndex),]
rownames(summary_index_user)  <- NULL
summary_index_user$base_rank <- rownames(summary_index_user)

for (i in 1:9){
#generate overall
type <- NULL
cost <- NULL

#subset dataframe by user selection
user_set <- Make_Subset(reschool_programs,type,cost)

#aggregate dataframe by user selection
agg_set <- Aggregate_Subset(user_set)

#merge n of user selected programs to block distances dataframe
merged_set <- Merge_Set(agg_set)

#apply decay function
merged_set$AccessIndex <- decay_fxn_softGravity(merged_set,"drive")

#get average of AccessIndex col, return to summary dataframe by blockgroup ID
summary_index_mover <- aggregate(AccessIndex ~ Id2,data=merged_set,FUN=sum)

#order
summary_index_mover <- summary_index_mover[order(summary_index_mover$AccessIndex),]

#prep for merge
rownames(summary_index_mover) <- NULL
summary_index_mover$rank.i <- rownames(summary_index_mover)
colnames(summary_index_mover)[colnames(summary_index_mover)=="rank.i"] <- paste("rank",i)

#print summary of AI for validation on if this is working
print(summary(summary_index_mover$AccessIndex))
summary_index_mover$AccessIndex <- NULL

#merge new data to dataframe
summary_index_user <- merge(summary_index_user,summary_index_mover,by="Id2",all.y=TRUE)
print(paste(i,"iteration..."))
}

#store as a new data frame
sensitivity_frame_softgravity <- summary_index_user 
```

###"Soft" Gravity Model Transit
```{r}
#run overall access index driving
library(scales)
#User specifies query (type), NULL = all
type <- NULL
cost <- NULL

#subset dataframe by user selection
user_set <- Make_Subset(reschool_programs,type,cost)

#aggregate dataframe by user selection
agg_set <- Aggregate_Subset(user_set)

#merge n of user selected programs to block distances dataframe
merged_set <- Merge_Set(agg_set)

#calculate access index per location 
merged_set$AccessIndex <- decay_fxn_softGravity(merged_set,"transit") #why not apply? who knows

#get average of AccessIndex col, return to summary dataframe by blockgroup ID
summary_index_user <- aggregate(AccessIndex ~ Id2,data=merged_set,FUN=sum) #critical that we use sum, not mean, BECAUSE, for some specifications some program lengths will drop

#make base dataframe
summary_index_user <- summary_index_user[order(summary_index_user$AccessIndex),]
rownames(summary_index_user)  <- NULL
summary_index_user$base_rank <- rownames(summary_index_user)

for (i in 1:9){
#generate overall
type <- NULL
cost <- NULL

#subset dataframe by user selection
user_set <- Make_Subset(reschool_programs,type,cost)

#aggregate dataframe by user selection
agg_set <- Aggregate_Subset(user_set)

#merge n of user selected programs to block distances dataframe
merged_set <- Merge_Set(agg_set)

#apply decay function
merged_set$AccessIndex <- decay_fxn_softGravity(merged_set,"transit")

#get average of AccessIndex col, return to summary dataframe by blockgroup ID
summary_index_mover <- aggregate(AccessIndex ~ Id2,data=merged_set,FUN=sum)

#order
summary_index_mover <- summary_index_mover[order(summary_index_mover$AccessIndex),]

#prep for merge
rownames(summary_index_mover) <- NULL
summary_index_mover$rank.i <- rownames(summary_index_mover)
colnames(summary_index_mover)[colnames(summary_index_mover)=="rank.i"] <- paste("rank",i)

#print summary of AI for validation on if this is working
print(summary(summary_index_mover$AccessIndex))
summary_index_mover$AccessIndex <- NULL

#merge new data to dataframe
summary_index_user <- merge(summary_index_user,summary_index_mover,by="Id2",all.y=TRUE)
print(paste(i,"iteration..."))
}

#store as a new data frame
sensitivity_frame_softgravity_transit <- summary_index_user 
```

###Plot Soft Gravity Driving
```{r}
library(reshape2)
library(ggplot2)
library(gridExtra)
#Prep & order frame
sensitivity_frame_softgravity <- sensitivity_frame_softgravity 
sensitivity_frame_softgravity$base_rank <- as.numeric(sensitivity_frame_softgravity$base_rank)
sensitivity_frame_softgravity <- sensitivity_frame_softgravity[order(sensitivity_frame_softgravity$base_rank),]

#subsetframe
Top_SG__25 <- sensitivity_frame_softgravity[1:25,]
sa_25_50 <- sensitivity_frame_softgravity[1:50,]
sa_50_75 <- sensitivity_frame_softgravity[1:75,]
sa_75_100 <- sensitivity_frame_softgravity[1:100,]
sa_tail <- sensitivity_frame_softgravity[456:480,]
sa_tailstart <- sensitivity_frame_softgravity[406:480,]


#Top 25 Plot
df <- melt(Top_SG__25,id=c("Id2","AccessIndex"),measure.vars=c("base_rank","rank 1","rank 2","rank 3","rank 4","rank 5","rank 6","rank 7","rank 8","rank 9"))
df$value <- as.numeric(df$value)
df$variable <- as.factor(df$variable)

Top_SG__25_Plot <- ggplot(df, aes(variable,value,group = Id2, color = as.factor(Id2))) + geom_line() + theme(legend.position="none") + labs(title="top 25") + ylim(1,490)

#Top 50 Plot
df <- melt(sa_25_50,id=c("Id2","AccessIndex"),measure.vars=c("base_rank","rank 1","rank 2","rank 3","rank 4","rank 5","rank 6","rank 7","rank 8","rank 9"))
df$value <- as.numeric(df$value)
df$variable <- as.factor(df$variable)

Top_SG__50_Plot <- ggplot(df, aes(variable,value,group = Id2, color = as.factor(Id2))) + geom_line() + theme(legend.position="none") + labs(title="top 50") + ylim(1,490)

#Top 75 Plot
df <- melt(sa_50_75,id=c("Id2","AccessIndex"),measure.vars=c("base_rank","rank 1","rank 2","rank 3","rank 4","rank 5","rank 6","rank 7","rank 8","rank 9"))
df$value <- as.numeric(df$value)
df$variable <- as.factor(df$variable)

Top_SG__75_Plot <- ggplot(df, aes(variable,value,group = Id2, color = as.factor(Id2))) + geom_line() + theme(legend.position="none") + labs(title="top 75") + ylim(1,490)

#Top 100 Plot
df <- melt(sa_75_100,id=c("Id2","AccessIndex"),measure.vars=c("base_rank","rank 1","rank 2","rank 3","rank 4","rank 5","rank 6","rank 7","rank 8","rank 9"))
df$value <- as.numeric(df$value)
df$variable <- as.factor(df$variable)

Top_SG__100_Plot <- ggplot(df, aes(variable,value,group = Id2, color = as.factor(Id2))) + geom_line() + theme(legend.position="none") + labs(title="top 100") + ylim(1,490)

#Bottom 50
df <- melt(sa_tail,id=c("Id2","AccessIndex"),measure.vars=c("base_rank","rank 1","rank 2","rank 3","rank 4","rank 5","rank 6","rank 7","rank 8","rank 9"))
df$value <- as.numeric(df$value)
df$variable <- as.factor(df$variable)

SG_Tail_Plot <- ggplot(df, aes(variable,value,group = Id2, color = as.factor(Id2))) + geom_line() + theme(legend.position="none") + labs(title="Bottom 50") + ylim(1,490)

#Bottom 25
df <- melt(sa_tailstart,id=c("Id2","AccessIndex"),measure.vars=c("base_rank","rank 1","rank 2","rank 3","rank 4","rank 5","rank 6","rank 7","rank 8","rank 9"))
df$value <- as.numeric(df$value)
df$variable <- as.factor(df$variable)

SG_Bottom25_Plot <- ggplot(df, aes(variable,value,group = Id2, color = as.factor(Id2))) + geom_line() + theme(legend.position="none") + labs(title="Bottom 100") + ylim(1,490)

#grid.arrange(Top_SG__25_Plot,Top_SG__50_Plot,Top_SG__75_Plot,Top_SG__100_Plot,SG_Bottom25_Plot,SG_Tail_Plot,ncol=3)
```

###Plot Soft Gravity Transit
```{r}
library(reshape2)
library(ggplot2)
library(gridExtra)
#Prep & order frame
sensitivity_frame_softgravity_transit <- sensitivity_frame_softgravity_transit 
sensitivity_frame_softgravity_transit$base_rank <- as.numeric(sensitivity_frame_softgravity_transit$base_rank)
sensitivity_frame_softgravity_transit <- sensitivity_frame_softgravity_transit[order(sensitivity_frame_softgravity_transit$base_rank),]

#subsetframe
Top_SGtransit_25 <- sensitivity_frame_softgravity_transit[1:25,]
sa_25_50 <- sensitivity_frame_softgravity_transit[1:50,]
sa_50_75 <- sensitivity_frame_softgravity_transit[1:75,]
sa_75_100 <- sensitivity_frame_softgravity_transit[1:100,]
sa_tail <- sensitivity_frame_softgravity_transit[456:480,]
sa_tailstart <- sensitivity_frame_softgravity_transit[406:480,]

#Top 25 Plot
df <- melt(Top_SGtransit_25,id=c("Id2","AccessIndex"),measure.vars=c("base_rank","rank 1","rank 2","rank 3","rank 4","rank 5","rank 6","rank 7","rank 8","rank 9"))
df$value <- as.numeric(df$value)
df$variable <- as.factor(df$variable)

Top_SGtransit_25_Plot <- ggplot(df, aes(variable,value,group = Id2, color = as.factor(Id2))) + geom_line() + theme(legend.position="none") + labs(title="top 25") + ylim(1,490)

#Top 50 Plot
df <- melt(sa_25_50,id=c("Id2","AccessIndex"),measure.vars=c("base_rank","rank 1","rank 2","rank 3","rank 4","rank 5","rank 6","rank 7","rank 8","rank 9"))
df$value <- as.numeric(df$value)
df$variable <- as.factor(df$variable)

Top_SGtransit_50_Plot <- ggplot(df, aes(variable,value,group = Id2, color = as.factor(Id2))) + geom_line() + theme(legend.position="none") + labs(title="top 50") + ylim(1,490)

#Top 75 Plot
df <- melt(sa_50_75,id=c("Id2","AccessIndex"),measure.vars=c("base_rank","rank 1","rank 2","rank 3","rank 4","rank 5","rank 6","rank 7","rank 8","rank 9"))
df$value <- as.numeric(df$value)
df$variable <- as.factor(df$variable)

Top_SGtransit_75_Plot <- ggplot(df, aes(variable,value,group = Id2, color = as.factor(Id2))) + geom_line() + theme(legend.position="none") + labs(title="top 75") + ylim(1,490)

#Top 100 Plot
df <- melt(sa_75_100,id=c("Id2","AccessIndex"),measure.vars=c("base_rank","rank 1","rank 2","rank 3","rank 4","rank 5","rank 6","rank 7","rank 8","rank 9"))
df$value <- as.numeric(df$value)
df$variable <- as.factor(df$variable)

Top_SGtransit_100_Plot <- ggplot(df, aes(variable,value,group = Id2, color = as.factor(Id2))) + geom_line() + theme(legend.position="none") + labs(title="top 100") + ylim(1,490)

#Bottom 50
df <- melt(sa_tail,id=c("Id2","AccessIndex"),measure.vars=c("base_rank","rank 1","rank 2","rank 3","rank 4","rank 5","rank 6","rank 7","rank 8","rank 9"))
df$value <- as.numeric(df$value)
df$variable <- as.factor(df$variable)

SGTransit_Tail_Plot <- ggplot(df, aes(variable,value,group = Id2, color = as.factor(Id2))) + geom_line() + theme(legend.position="none") + labs(title="Bottom 50") + ylim(1,490)

#Bottom 25
df <- melt(sa_tailstart,id=c("Id2","AccessIndex"),measure.vars=c("base_rank","rank 1","rank 2","rank 3","rank 4","rank 5","rank 6","rank 7","rank 8","rank 9"))
df$value <- as.numeric(df$value)
df$variable <- as.factor(df$variable)

SGTransit_Bottom25_Plot <- ggplot(df, aes(variable,value,group = Id2, color = as.factor(Id2))) + geom_line() + theme(legend.position="none") + labs(title="Bottom 100") + ylim(1,490)

#grid.arrange(Top_SGtransit_25_Plot,Top_SGtransit_50_Plot,Top_SGtransit_75_Plot,Top_SGtransit_100_Plot,SGTransit_Bottom25_Plot,SGTransit_Tail_Plot,ncol=3)
```

##Conclusion
While the gravity model is the least sensitive to the choice in scale, the model is extremely sensitive to the proximity of nearby addresses, producing that are both questionable with respect to the realistic assumption that there is some in difference between a travel of time in the case of very short travel times (for example, one minute vs. two minutes) as well as bad for interpreabillity as it heavily weights blockgroups with centroids located near programs, which is hardly an accurate representation and the location of all homes in a centroid.

By contrast, the exp model and the gravity threshold model yield conflicting results. Transit times are extremely sensitive to the exp model with a scale of 5-15 minutes, while driving times are extremely sensitive to the gravity threshold model with respect to 5-15 mintues. In either case, the somewhat arbitrary choice of scale could signficantly alter conclusions, questioning the construct validity of our results.

In contrast, the use of the "soft gravity model" is by far the least sensitive to choice of scale, and provides results which are useful for interpretation. 