---
title: "Access_Index_Functions"
author: "Andrew Taylor"
date: "7/23/2018"
output: html_document
---
##Intro
This markdown defines functions for use in calculating an access index per user specifications on filtering reschool program data. Additionally, this markdown provides leaflet code for plotting the results of the index by census block group. This markdown reads census_centroids as a shapefile from my machine, as we haven't pushed this to the RDS yet. ***AS OF JULY 27th*** We have deicded to pre-compute different combinations to return an average (rather than aggregate score) for the shiny interface. 

```{r}
library(rgeos)
library(rgdal)
library(raster)
library(tidyverse)
library(RPostgreSQL)
```

###RDS Pull
```{r}
mypath <- dirname(rstudioapi::getActiveDocumentContext()$path)
setwd(mypath)

# loads the PostgreSQL driver
drv <- dbDriver("PostgreSQL")

# make a file called "cred.txt" of the form
#
# user: "YOUR_AWS_USERNAME"
# password: "YOUR_AWS_PASSWORD"
#
# in the directory above the osr_dssg2018 folder

source(file.path(dirname(dirname(mypath)),"cred.txt"))

# creates a connection to the postgres database
# note that "con" will be used later in each connection to the database
con <- dbConnect(drv, dbname = "dssg2018uw",
                 host = "localhost", port =9000,
                 user = user, password = password)

#get reschool_programs
reschool_programs <- dbGetQuery(con, "select * from clean.reschool_summer_programs")
block_distance <- dbGetQuery(con, "select * from clean.block_distance")

#disconnect
dbDisconnect(con) 
dbUnloadDriver(drv)
```

###Establishing Cost Thresholds and Umbrella Program Types
```{r}
reschool_programs$academic <- reschool_programs$has_academic==TRUE | reschool_programs$has_stem==TRUE

reschool_programs$art <- reschool_programs$has_arts==TRUE | reschool_programs$has_cooking==TRUE | reschool_programs$has_dance==TRUE | reschool_programs$has_drama==TRUE | reschool_programs$has_music==TRUE

#cost thresolds, to be updated w/ time data from selam
reschool_programs$free <- reschool_programs$session_cost==0
reschool_programs$lowcost <- reschool_programs$session_cost<=50
reschool_programs$anycost <- TRUE

#establishing "1" per row for use in aggergation
reschool_programs$n <- 1 
```


###Defining for Flexible Access Index Calculation
Calculating the Access Index requires a few discrete steps. Currently, these are established with the functions defined below. 
```{r}
#1. Setting type and cost parameteres

type <- c("academic","art") #needs to be a vector of whatever get's passed to shiny
cost <- c("free") #these are examples, if set to NULL for either, the set of functions will return the full index

#2. First we subset to ReSchool Programs Dataset per user specificiations

Make_Subset <- function(dataframe,type=NULL,cost=NULL){
  dataframe$n <- 1
  if(is.null(type)){
    dataframe <- dataframe
    }
  if(length(type)==1){
    dataframe <- dataframe[dataframe[,type],]
  }
  if(length(type)>1){
    dataframe <- dataframe[apply(FUN=any,X=dataframe[,c(type)],MARGIN=1),]
  }
  if(is.null(cost)){
    return(dataframe)
  }
  else{
     dataframe <- dataframe <- dataframe[dataframe[,cost],]
  }
}

#3. Then we aggregate sessions the subset per the specifications

Aggregate_Subset <- function(dataframe){
  return(aggregate(n ~ session_address_1+lat+long, data=dataframe,FUN=sum)) #note that "n" in this case ==1 per row, which n this case corresponds to one unique program session
}

#4. Here we merge the aggregated sessiosn into the block_distance data set to incorporate travel times 

Merge_Set <- function(dataframe){
  relevant_columns <- c("Id2","session_address_1","driving_morning","walking_morning","transit_morning","kilometers","block_lat","block_long")
  block_distance_new <- block_distance[,relevant_columns] #here we're cutting duplicate program_addresses, and other antiquated info (previous aggregations, PCT vhcl ownership) that was use in previous calculations
  temp <- merge(block_distance_new,dataframe,all.y=TRUE)
  temp$n[is.na(temp$n)==TRUE] <- 0
  return(temp)
}

#5 Here we define the gravity threshold decay function. Note that as of July 25th we've decided not use this equation, and instead use the exponential equation per the results of the sensitivity analysis. 

decay_fxn <- function(dataframe,mode){
  if (mode=="transit") {
    scale = 5
    column="transit_morning"
  }
  if (mode=="drive"){
    scale = 5
    column="driving_morning"
  }
  mode=mode
  dataframe[,"n"]*(ifelse(dataframe[,column]<=scale,1,(1/(dataframe[,column]/scale))^2))
}

#6 Here we define the expoential decay function

decay_fxn_exp <- function(dataframe,mode){
  if (mode=="transit") {
    scale = 5
    column="transit_morning"
  }
  if (mode=="drive"){
    scale = 5
    column="driving_morning"
  }
  mode=mode
  dataframe[,"n"]*exp(-dataframe[,column]/scale)
}

decay_fxn_softGravity <- function(dataframe,mode){
  if (mode=="transit") {
    scale = 10
    column="transit_morning"
  }
  if (mode=="drive"){
    scale = 10
    column="driving_morning"
  }
  mode=mode
  dataframe[,"n"]*(1/(1+(dataframe[,column]/scale))^2)
}

#7 here we define a function to normalize results on a 0-1 scale. Note that after we calculate an total acess index in this script, max will be == the max AI of any one blockgroup.

normalize <- function(vec,max) {
  return((vec/max)*100)
}
```

###User Script for Driving
This is the exact script to use to query, for an example type.
```{r}
#User specifies query (type)
#type <- c("art") #"type" can include categories or cost thresholds. Pending cost per day calculations from Selam
#cost <- c("free")
type <- c("academic")
cost <- c("free")

#subset dataframe by user selection
user_set <- Make_Subset(reschool_programs,type,cost)

#aggregate dataframe by user selection
agg_set <- Aggregate_Subset(user_set)

#merge n of user selected programs to block distances dataframe
merged_set <- Merge_Set(agg_set)

#calculate access index per location 
merged_set$AccessIndex <- decay_fxn_softGravity(merged_set,"drive") #why not apply? who knows

#get average of AccessIndex col, return to summary dataframe by blockgroup ID
summary_index_user <- aggregate(AccessIndex ~ Id2,data=merged_set,FUN=sum) #critical that we use sum, not mean, BECAUSE, for some specifications some program lengths will drop

#normalize
max_access <- max(summary_index_user$AccessIndex)
summary_index_user$AccessIndex <- normalize(summary_index_user$AccessIndex,max_access) 
```

###User Script for Transit
```{r}
library(scales)
#User specifies query (type)
#type <- c("has_sports","art")
type <- c("academic")
cost <- c("anycost")

#subset dataframe by user selection
user_set <- Make_Subset(reschool_programs,type,cost)

#aggregate dataframe by user selection
agg_set <- Aggregate_Subset(user_set)

#merge n of user selected programs to block distances dataframe
merged_set <- Merge_Set(agg_set)

#replace na and 0 transit times with wallking times
merged_set$transit_morning[is.na(merged_set$transit_morning)] <- merged_set$walking_morning[is.na(merged_set$transit_morning)] #there shouldn't be NAs or 0s
merged_set$transit_morning[merged_set$transit_morning==0] <- merged_set$walking_morning[merged_set$transit_morning==0] #but if there is, as sometimes occurs with the API calls, here we replace non-existant transit times with walking times, under the assumption the purpose of this index is to view access for individuals without a car

#calculate access index per location, #get average of AccessIndex col, return to summary dataframe by blockgroup ID
system.time(merged_set$AccessIndex <- decay_fxn_softGravity(merged_set,"transit"))
merged_set$CarAccess <- decay_fxn_softGravity(merged_set,"drive")

system.time(summary_index_user_transit_graphic <- aggregate(CarAccess ~ Id2,data=merged_set,FUN=sum))
system.time(summary_index_user_transit <- aggregate(AccessIndex ~ Id2,data=merged_set,FUN=sum))

#normalize
max_access <- max(summary_index_user_transit_graphic$CarAccess)
max_transit_access <- max(summary_index_user_transit$AccessIndex)

summary_index_user_transit$AccessIndexGraphic <- normalize(summary_index_user_transit$AccessIndex,max_transit_access)
summary_index_user_transit$AccessIndex <- normalize(summary_index_user_transit$AccessIndex,max_access)
```

##Generating Average Scores 
For overall, we want a weighted average based on category, not a sum of access to programs with no categories. So here we pre-run the index for each discrete category, then the get return the average for an overall dataframe. This chunk returns both driving and transit overall scores.  

```{r}
#Academic
type <- c("academic")
cost <- NULL
user_set <- Make_Subset(reschool_programs,type,cost)
agg_set <- Aggregate_Subset(user_set)
merged_set <- Merge_Set(agg_set)
merged_set$AccessIndex <- decay_fxn_softGravity(merged_set,"drive")
summary_index_user <- aggregate(AccessIndex ~ Id2,data=merged_set,FUN=sum)
max_access <- max(summary_index_user$AccessIndex)
summary_index_user$AccessIndex <- normalize(summary_index_user$AccessIndex,max_access)

#Save index
summary_academic_driving <- summary_index_user

#Arts
type <- c("art")
cost <- NULL
user_set <- Make_Subset(reschool_programs,type,cost)
agg_set <- Aggregate_Subset(user_set)
merged_set <- Merge_Set(agg_set)
merged_set$AccessIndex <- decay_fxn_softGravity(merged_set,"drive")
summary_index_user <- aggregate(AccessIndex ~ Id2,data=merged_set,FUN=sum)
max_access <- max(summary_index_user$AccessIndex)
summary_index_user$AccessIndex <- normalize(summary_index_user$AccessIndex,max_access)
summary_art_driving <- summary_index_user

#Sports
type <- c("has_sports")
cost <- NULL
user_set <- Make_Subset(reschool_programs,type,cost)
agg_set <- Aggregate_Subset(user_set)
merged_set <- Merge_Set(agg_set)
merged_set$AccessIndex <- decay_fxn_softGravity(merged_set,"drive")
summary_index_user <- aggregate(AccessIndex ~ Id2,data=merged_set,FUN=sum)
max_access <- max(summary_index_user$AccessIndex)
summary_index_user$AccessIndex <- normalize(summary_index_user$AccessIndex,max_access)
summary_sports_driving <- summary_index_user

#Nature
type <- c("has_nature")
cost <- NULL
user_set <- Make_Subset(reschool_programs,type,cost)
agg_set <- Aggregate_Subset(user_set)
merged_set <- Merge_Set(agg_set)
merged_set$AccessIndex <- decay_fxn_softGravity(merged_set,"drive")
summary_index_user <- aggregate(AccessIndex ~ Id2,data=merged_set,FUN=sum)
max_access <- max(summary_index_user$AccessIndex)
summary_index_user$AccessIndex <- normalize(summary_index_user$AccessIndex,max_access)
summary_nature_driving <- summary_index_user

#Weighted Average Index
overall_driving <- summary_nature_driving
overall_driving$AccessIndex <-(summary_nature_driving$AccessIndex+summary_sports_driving$AccessIndex+summary_art_driving$AccessIndex+summary_academic_driving$AccessIndex)/4 #note that we can't use sum because that delivers col sums per each dataframe

#Transit Academic
type <- c("academic")
cost <- NULL
user_set <- Make_Subset(reschool_programs,type,cost)
agg_set <- Aggregate_Subset(user_set)
merged_set <- Merge_Set(agg_set)
merged_set$transit_morning[is.na(merged_set$transit_morning)] <- merged_set$walking_morning[is.na(merged_set$transit_morning)] #there shouldn't be NAs or 0s
merged_set$transit_morning[merged_set$transit_morning==0] <- merged_set$walking_morning[merged_set$transit_morning==0]
merged_set$AccessIndex <- decay_fxn_softGravity(merged_set,"transit")

summary_index_user <- aggregate(AccessIndex ~ Id2,data=merged_set,FUN=sum)
merged_set$CarAccessIndex <- decay_fxn_softGravity(merged_set,"drive")
summary_index_transit <- aggregate(CarAccessIndex ~ Id2, data=merged_set,FUN=sum)
max_access <- max(summary_index_transit$CarAccessIndex)

summary_index_user$AccessIndex <- normalize(summary_index_user$AccessIndex,max_access)

#Save index
summary_academic_transit <- summary_index_user

#TransitArts
type <- c("art")
cost <- NULL
user_set <- Make_Subset(reschool_programs,type,cost)
agg_set <- Aggregate_Subset(user_set)
merged_set <- Merge_Set(agg_set)
merged_set$transit_morning[is.na(merged_set$transit_morning)] <- merged_set$walking_morning[is.na(merged_set$transit_morning)] #there shouldn't be NAs or 0s
merged_set$transit_morning[merged_set$transit_morning==0] <- merged_set$walking_morning[merged_set$transit_morning==0]
merged_set$AccessIndex <- decay_fxn_softGravity(merged_set,"transit")
summary_index_user <- aggregate(AccessIndex ~ Id2,data=merged_set,FUN=sum)

merged_set$CarAccessIndex <- decay_fxn_softGravity(merged_set,"drive")
summary_index_transit <- aggregate(CarAccessIndex ~ Id2, data=merged_set,FUN=sum)
max_access <- max(summary_index_transit$CarAccessIndex)

summary_index_user$AccessIndex <- normalize(summary_index_user$AccessIndex,max_access)
summary_art_transit <- summary_index_user

#TransitSports
type <- c("has_sports")
cost <- NULL
user_set <- Make_Subset(reschool_programs,type,cost)
agg_set <- Aggregate_Subset(user_set)
merged_set <- Merge_Set(agg_set)
merged_set$transit_morning[is.na(merged_set$transit_morning)] <- merged_set$walking_morning[is.na(merged_set$transit_morning)] #there shouldn't be NAs or 0s
merged_set$transit_morning[merged_set$transit_morning==0] <- merged_set$walking_morning[merged_set$transit_morning==0]
merged_set$AccessIndex <- decay_fxn_softGravity(merged_set,"transit")
summary_index_user <- aggregate(AccessIndex ~ Id2,data=merged_set,FUN=sum)

merged_set$CarAccessIndex <- decay_fxn_softGravity(merged_set,"drive")

summary_index_transit <- aggregate(CarAccessIndex ~ Id2, data=merged_set,FUN=sum)
max_access <- max(summary_index_transit$CarAccessIndex)

summary_index_user$AccessIndex <- normalize(summary_index_user$AccessIndex,max_access)
summary_sports_transit <- summary_index_user

#Transit Nature
type <- c("has_nature")
cost <- NULL
user_set <- Make_Subset(reschool_programs,type,cost)
agg_set <- Aggregate_Subset(user_set)
merged_set <- Merge_Set(agg_set)
merged_set$transit_morning[is.na(merged_set$transit_morning)] <- merged_set$walking_morning[is.na(merged_set$transit_morning)] #there shouldn't be NAs or 0s
merged_set$transit_morning[merged_set$transit_morning==0] <- merged_set$walking_morning[merged_set$transit_morning==0]
merged_set$AccessIndex <- decay_fxn_softGravity(merged_set,"transit")
summary_index_user <- aggregate(AccessIndex ~ Id2,data=merged_set,FUN=sum)

merged_set$CarAccessIndex <- decay_fxn_softGravity(merged_set,"drive")
summary_index_transit <- aggregate(CarAccessIndex ~ Id2, data=merged_set,FUN=sum)
max_access <- max(summary_index_transit$CarAccessIndex)

summary_index_user$AccessIndex <- normalize(summary_index_user$AccessIndex,max_access)
summary_nature_transit <- summary_index_user

#Transit Weighted Average Index
overall_transit <- summary_nature_transit
overall_transit$AccessIndex <-(summary_nature_transit$AccessIndex+summary_sports_transit$AccessIndex+summary_art_transit$AccessIndex+summary_academic_transit$AccessIndex)/4

#overall programs
##for use in the final plot of circles
full_agg_set <- aggregate(n ~ session_address_1+lat+long, data=reschool_programs,FUN=sum)
```

###Get Census Blocks Data for Plotting
Note that you need all files in this directory, and that it is not in the RDS. 
```{r}
library(rgeos)
library(rgdal)
library(raster)
#census_blocks <- SpatialPointsDataFrame(gCentroid(spdf, byid=TRUE), spdf@data, match.ID = FALSE)
```

###Validation Plot Driving Summary Index
**NOTE** the specifications for this plot, and the transit plot below are for use in validation (i.e, checking for errors) not for use in presenting results. 
```{r}
library(leaflet)
library(RColorBrewer)

#Ordering
census_blocks$Id2 <- as.numeric(census_blocks$Id2)
merged_set$Id2 <- as.numeric(merged_set$Id2)
summary_index_user$Id2 <- as.numeric(summary_index_user$Id2)

census_blocks <- census_blocks[order(census_blocks$Id2),]
merged_set <- merged_set[order(merged_set$Id2),]
summary_index_user <- summary_index_user[order(summary_index_user$Id2),] 

#specifiying color pallet
colourCount <- length(unique(agg_set$n))
getPalette <- colorRampPalette(brewer.pal(9,"Reds"))

pal <- colorNumeric(palette = "Blues",domain = 0:100)

pal2 <- colorNumeric(
  palette = rev(getPalette(colourCount)),
  domain = range(agg_set$n))

#indicating one popup info
info <- paste("n",agg_set$n)

#creating leaflet map
Ai_driving <- leaflet(census_blocks) %>%
  addTiles() %>%
  addPolygons(color = "RdYlBu",weight =1,smoothFactor = 0.5,opacity = 0.5,fillOpacity = 0.75,
              fillColor = ~pal(summary_index_user$AccessIndex),
              popup=paste(as.numeric(summary_index_user$Id2),summary_index_user$AccessIndex,as.numeric(census_blocks$Id2)),
              highlightOptions = highlightOptions(color = "white", weight = 0.1)) %>%
  
  addLegend(pal = pal, values = ~summary_index_user$AccessIndex, opacity = 1,title = "Driving AI Score") %>%
  
  addCircles(data = agg_set, 
             opacity = 1,weight=7, popup=info,
             color = ~pal2(agg_set$n) ) %>%
  
  setView(lat=39.7,lng=-104.9,zoom=11.5)
```

###Validation Plot Transit Index
```{r}
library(leaflet)
library(RColorBrewer)

#Ordering
census_blocks$Id2 <- as.numeric(census_blocks$Id2)
merged_set$Id2 <- as.numeric(merged_set$Id2)
summary_index_user_transit$Id2 <- as.numeric(summary_index_user_transit$Id2)

census_blocks <- census_blocks[order(census_blocks$Id2),]
merged_set <- merged_set[order(merged_set$Id2),]
summary_index_user_transit <- summary_index_user_transit[order(summary_index_user_transit$Id2),] 

#specifiying color pallet
colourCount <- length(unique(agg_set$n))
getPalette <- colorRampPalette(brewer.pal(9,"Reds"))

pal <- colorNumeric(palette = "Greens",domain = 0:100)

pal2 <- colorNumeric(
  palette = rev(getPalette(colourCount)),
  domain = range(agg_set$n))

#indicating one popup info
info <- paste("n",agg_set$n)

#creating leaflet map
Ai_transit <- leaflet(census_blocks) %>%
  addTiles() %>%
  addPolygons(color = "RdYlBu",weight =1,smoothFactor = 0.5,opacity = 0.5,fillOpacity = 0.9,
              fillColor = ~pal(summary_index_user_transit$AccessIndexGraphic),
              popup=paste(as.numeric(summary_index_user_transit$Id2),summary_index_user_transit$AccessIndex,as.numeric(census_blocks$Id2)),
              highlightOptions = highlightOptions(color = "white", weight = 0.1)) %>%
  
  addLegend(pal = pal, values = ~summary_index_user_transit$AccessIndex, opacity = 1,title = "Transit AI Score") %>%
  
  addCircles(data = agg_set, 
             opacity = 1,weight=7, popup=info,
             color = ~pal2(agg_set$n) ) %>%
  
  setView(lat=39.7,lng=-104.9,zoom=11.5)
```

###Plotting Overal Driving & Transit 
Overall = average of four categories calcualted above. 
```{r}
library(leaflet)
library(RColorBrewer)
#Ordering
census_blocks$Id2 <- as.numeric(census_blocks$Id2)
merged_set$Id2 <- as.numeric(merged_set$Id2)
overall_driving$Id2 <- as.numeric(overall_driving$Id2)

census_blocks <- census_blocks[order(census_blocks$Id2),]
merged_set <- merged_set[order(merged_set$Id2),]
overall_driving <- overall_driving[order(overall_driving$Id2),] 

#specifiying color pallet
colourCount <- length(unique(agg_set$n))
getPalette <- colorRampPalette(brewer.pal(9,"Reds"))

range <- range(overall_driving$AccessIndex)

pal <- colorNumeric(palette = "Blues",domain = range)

pal2 <- colorNumeric(
  palette = rev(getPalette(colourCount)),
  domain = range(full_agg_set$n))

#indicating one popup info
info <- paste("n",agg_set$n)

#creating leaflet map
Ai_driving_overall <- leaflet(census_blocks) %>%
  addTiles() %>%
  addPolygons(color = "RdYlBu",weight =1,smoothFactor = 0.5,opacity = 0.5,fillOpacity = 0.75,
              fillColor = ~pal(overall_driving$AccessIndex),
              popup=paste("AI =",round(overall_driving$AccessIndex,digits=2)),
              highlightOptions = highlightOptions(color = "white", weight = 0.1)) %>%
  addLegend(pal = pal2, values = ~full_agg_set$n, opacity = 1,title = "Sessions") %>%
  addCircles(data = full_agg_set, 
             opacity = 1,weight=7, popup=paste("n of sessions =",agg_set$n),
             color = ~pal2(full_agg_set$n) ) %>%
  setView(lat=39.7,lng=-104.9,zoom=11.5) %>%
  addLegend(pal = pal, values = ~overall_driving$AccessIndex, opacity = 1, title = "Access Index")

#Ordering
census_blocks$Id2 <- as.numeric(census_blocks$Id2)
merged_set$Id2 <- as.numeric(merged_set$Id2)
overall_transit$Id2 <- as.numeric(overall_transit$Id2)

census_blocks <- census_blocks[order(census_blocks$Id2),]
merged_set <- merged_set[order(merged_set$Id2),]
overall_transit <- overall_transit[order(overall_transit$Id2),] 

#specifiying color pallet
colourCount <- length(unique(agg_set$n))
getPalette <- colorRampPalette(brewer.pal(9,"Reds"))

range <- range(overall_transit$AccessIndex)

pal <- colorNumeric(palette = "Greens",domain = range)

pal2 <- colorNumeric(
  palette = rev(getPalette(colourCount)),
  domain = range(full_agg_set$n))

#indicating one popup info
info <- paste("n of sessions =",agg_set$n)

#creating leaflet map
Ai_transit_overall <- leaflet(census_blocks) %>%
  addTiles() %>%
  addPolygons(color = "RdYlBu",weight =1,smoothFactor = 0.5,opacity = 0.5,fillOpacity = 0.9,
              fillColor = ~pal(overall_transit$AccessIndex),
              popup=paste("AI =",round(overall_transit$AccessIndex,digits=2)),
              highlightOptions = highlightOptions(color = "white", weight = 0.1)) %>%
  addLegend(pal = pal2, values = ~full_agg_set$n, opacity = 1,title = "Sessions") %>%
  addCircles(data = full_agg_set, 
             opacity = 1,weight=7, popup=info,
             color = ~pal2(full_agg_set$n) ) %>%
  addLegend(pal = pal, values = ~overall_transit$AccessIndex, opacity = 1, title = "Access Index") %>%
  setView(lat=39.7,lng=-104.9,zoom=11.5)

```


###Plot Driving & Transit Normalized Pallet To Subset
Use these plots for actual visualiztion, use the above for validation.
```{r}
library(leaflet)
library(RColorBrewer)
#Ordering
census_blocks$Id2 <- as.numeric(census_blocks$Id2)
merged_set$Id2 <- as.numeric(merged_set$Id2)
summary_index_user$Id2 <- as.numeric(summary_index_user$Id2)

census_blocks <- census_blocks[order(census_blocks$Id2),]
merged_set <- merged_set[order(merged_set$Id2),]
summary_index_user <- summary_index_user[order(summary_index_user$Id2),] 

#specifiying color pallet
colourCount <- length(unique(agg_set$n))
getPalette <- colorRampPalette(brewer.pal(9,"Reds"))

range <- range(summary_index_user$AccessIndex)

pal <- colorNumeric(palette = "Blues",domain = range)

pal2 <- colorNumeric(
  palette = rev(getPalette(colourCount)),
  domain = range(agg_set$n))

#indicating one popup info
info <- paste("n",agg_set$n)

#creating leaflet map
Ai_driving_normal <- leaflet(census_blocks) %>%
  addTiles() %>%
  addPolygons(color = "RdYlBu",weight =1,smoothFactor = 0.5,opacity = 0.5,fillOpacity = 0.75,
              fillColor = ~pal(summary_index_user$AccessIndex),
              popup=paste("AI =",round(summary_index_user$AccessIndex,digits=2)),
              highlightOptions = highlightOptions(color = "white", weight = 0.1)) %>%
  addLegend(pal = pal2, values = ~agg_set$n, opacity = 1,title = "Sessions") %>%
  addCircles(data = agg_set, 
             opacity = 1,weight=7, popup=paste("n of sessions =",agg_set$n),
             color = ~pal2(agg_set$n) ) %>%
  setView(lat=39.7,lng=-104.9,zoom=11.5) %>%
  addLegend(pal = pal, values = ~summary_index_user$AccessIndex, opacity = 1, title = "Access Index")

#Ordering
census_blocks$Id2 <- as.numeric(census_blocks$Id2)
merged_set$Id2 <- as.numeric(merged_set$Id2)
summary_index_user_transit$Id2 <- as.numeric(summary_index_user_transit$Id2)

census_blocks <- census_blocks[order(census_blocks$Id2),]
merged_set <- merged_set[order(merged_set$Id2),]
summary_index_user_transit <- summary_index_user_transit[order(summary_index_user_transit$Id2),] 

#specifiying color pallet
colourCount <- length(unique(agg_set$n))
getPalette <- colorRampPalette(brewer.pal(9,"Reds"))

range <- range(summary_index_user_transit$AccessIndex)

pal <- colorNumeric(palette = "Greens",domain = range)

pal2 <- colorNumeric(
  palette = rev(getPalette(colourCount)),
  domain = range(agg_set$n))

#indicating one popup info
info <- paste("n of sessions =",agg_set$n)

#creating leaflet map
Ai_transit_normal <- leaflet(census_blocks) %>%
  addTiles() %>%
  addPolygons(color = "RdYlBu",weight =1,smoothFactor = 0.5,opacity = 0.5,fillOpacity = 0.9,
              fillColor = ~pal(summary_index_user_transit$AccessIndex),
              popup=paste("AI =",round(summary_index_user_transit$AccessIndex,digits=2)),
              highlightOptions = highlightOptions(color = "white", weight = 0.1)) %>%
  addLegend(pal = pal2, values = ~agg_set$n, opacity = 1,title = "Sessions") %>%
  addCircles(data = agg_set, 
             opacity = 1,weight=7, popup=info,
             color = ~pal2(agg_set$n) ) %>%
  addLegend(pal = pal, values = ~summary_index_user_transit$AccessIndex, opacity = 1, title = "Access Index") %>%
  setView(lat=39.7,lng=-104.9,zoom=11.5)
```

