---
title: "Pre_Computing_Access_Indecies"
author: "Andrew Taylor"
date: "7/27/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Intro
This markdown has the loops for precomputing combinations of program categories and cost for use in the Rshiny app. **NOTE** you must have the Access Index functions defined in your workspace, and the "reschool_data" and ""acs_demographics" pulled from the RDS to run this notebook.  

###Set up dataframe for storing results
```{r}
Id2 <- c(acs_demographics$id2)
base_frame <- as.data.frame(Id2)
```

###Driving Dataframe
```{r}
#Create driving frame
driving_index <- base_frame

#Creating categories list for loop

categories <- c("academic","art","has_sports","has_nature")

#Creating cost list for loop

costs <- c("free","lowcost","anycost")

#Loop for all categories and costs

for (i in 1:length(categories)){
  type <- categories[i]
  for (x in 1:length(costs)){
    cost <- costs[x]
    user_set <- Make_Subset(reschool_programs,type,cost)
    agg_set <- Aggregate_Subset(user_set)
    merged_set <- Merge_Set(agg_set)
    merged_set$AccessIndex <- decay_fxn_softGravity(merged_set,"drive") 
    summary_index_user <- aggregate(AccessIndex ~ Id2,data=merged_set,FUN=sum) 
    max_access <- max(summary_index_user$AccessIndex)
    #update colnames for merge
    summary_index_user$AccessIndex <- normalize(summary_index_user$AccessIndex,max_access)
    colnames(summary_index_user)[colnames(summary_index_user)=="AccessIndex"] <- paste("AI",type,cost,sep = "_") #flexible name writing for loop
    summary_index_user$Id2 <- as.numeric(summary_index_user$Id2) #matching str to baseframe 
    driving_index <- merge(summary_index_user,driving_index,by=c("Id2"))
  }
  print(paste(i,"working"))
  if(length(driving_index)>11){
    print(colnames(driving_index))
  }
}

#Create an overall average by category with no cost

driving_index$AI_overall <- rowMeans(driving_index[,c("AI_has_nature_anycost",'AI_has_sports_anycost','AI_art_anycost','AI_academic_anycost')])
```

###Transit Dataframe
```{r}
#Create driving frame
transit_index <- base_frame

#Creating categories list for loop

categories <- c("academic","art","has_sports","has_nature")

#Creating cost list for loop

costs <- c("free","lowcost","anycost")

#Loop for all categories and costs

for (i in 1:length(categories)){
  type <- categories[i]
  for (x in 1:length(costs)){
    cost <- costs[x]
    user_set <- Make_Subset(reschool_programs,type,cost)
    agg_set <- Aggregate_Subset(user_set)
    merged_set <- Merge_Set(agg_set)
    
    #replace na and 0 transit times with wallking times
    merged_set$transit_morning[is.na(merged_set$transit_morning)] <- merged_set$walking_morning[is.na(merged_set$transit_morning)] #there shouldn't be NAs or 0s
    merged_set$transit_morning[merged_set$transit_morning==0] <- merged_set$walking_morning[merged_set$transit_morning==0] #but if there is, as sometimes occurs with the API calls,        here we replace non-existant transit times with walking times, under the assumption the purpose of this index is to view access for individuals without a ca
    
    merged_set$AccessIndex <- decay_fxn_softGravity(merged_set,"transit") 
    merged_set$CarAccess <- decay_fxn_softGravity(merged_set,"drive")
    
    summary_index_user_transit_graphic <- aggregate(CarAccess ~ Id2,data=merged_set,FUN=sum)
    summary_index_user_transit <- aggregate(AccessIndex ~ Id2,data=merged_set,FUN=sum)

    #normalize, note here we make two columns, one to compare to driving, one for visual rep
    max_access <- max(summary_index_user_transit_graphic$CarAccess)
    summary_index_user_transit$AccessIndex <- normalize(summary_index_user_transit$AccessIndex,max_access)
    
    #update colnames for merge
    colnames(summary_index_user_transit)[colnames(summary_index_user_transit)=="AccessIndex"] <- paste("AI",type,cost,sep = "_") #flexible name writing for loop
    summary_index_user_transit$Id2 <- as.numeric(summary_index_user_transit$Id2) #matching str to baseframe 
    transit_index <- merge(summary_index_user_transit,transit_index,by=c("Id2"))
  }
  print(paste(i,"working"))
  if(length(transit_index)>11){
    print(colnames(transit_index))
  }
}

#Create an overall average by category with no cost

transit_index$AI_overall <- rowMeans(transit_index[,c("AI_has_nature_anycost",'AI_has_sports_anycost','AI_art_anycost','AI_academic_anycost')])
```

###Push to RDS
Remember you have to log in through the command line before running this chunk. 
```{r}
library("RPostgreSQL")

# loads the PostgreSQL driver
drv <- dbDriver("PostgreSQL")

con <- dbConnect(drv, dbname = "dssg2018uw",
                 host = "localhost", port = 9000,
                 user = user, password = password)

# specify data and schema and table names
data <- driving_index  # must be already loaded in your environment
schemaName <- "clean"
tableName <- "driving_index"

dbWriteTable(con,
             c(schemaName, tableName),
             value = data,
             row.names = FALSE,
             overwrite = TRUE  # overwrite an existing table
)

data <- transit_index  # must be already loaded in your environment
schemaName <- "clean"
tableName <- "transit_index"

dbWriteTable(con,
             c(schemaName, tableName),
             value = data,
             row.names = FALSE,
             overwrite = TRUE)


dbListTables(con) #check both are saved

dbDisconnect(con) 
dbUnloadDriver(drv)
```

