---
title: "Access_Index_Calculator"
author: "Andrew Taylor"
date: "7/19/2018"
output: html_document
---

```{r}
library(RPostgreSQL)
# loads the PostgreSQL driver
drv <- dbDriver("PostgreSQL")

# creates a connection to the postgres database
# note that "con" will be used later in each connection to the database
con <- dbConnect(drv, dbname = "dssg2018uw",
                 host = "localhost", port = 9000,
                 user = user, password = password)
```

First get the block group distances from the RDS.
```{r}
block_distance <- dbGetQuery(con, "select * from clean.block_distance")
```

Read in the shape census file
```{r}
#double note: update the link below to your directory
spdf <- readOGR("../census_block_groups/shape_census.shp") #reads the shapefile, NOTE: you need all shapefiles in the directory present
```

#Intro
This notebook takes the block group distances, adds in additional data needed to calculate the Access Index, and then as of 07/20/18, shows our loops for running the equation on the access index for **overall** programs. **HOWEVER** the plots at the end make use of the aggregated dataframes used to merge into the block_distance data-frame, so use your judgement. Maybe see the block_group_distances notebook to get those aggregations. Or don't, I don't care, we're going to change this in the next day or two. 

Now let's define our access index function. 

```{r}
fxn <- function(vector, scale){
  #2^(-vector/scale)
  ifelse(vector<=scale,1,(1/(vector/scale))^2)
}

driving_commute_denver <- 10 #wow these seem arbitrary huh?
transit_commute_denver <- 20 #HUH?
```

###Access index loop
```{r}
block_list <- unique(block_distance$Id2)
block_mover <- block_distance
block_mover$Access_Index <- 0
access_indexices <- as.data.frame(setNames(replicate(16,character(0), simplify = F), colnames(block_distance))) #make an empty data frame identical to block distances
summary_index <- census_centroids
summary_index$Access_Index <- 0

block_distance$Id2 <- as.numeric(block_distance$Id2)
block_distance <- block_distance[order(block_distance$Id2),]

block_list <- as.numeric(block_list)
block_list <- sort(block_list)

summary_index$blockID <- as.numeric(summary_index$blockID)
summary_index <- summary_index[order(summary_index$blockID),]

for (i in 1:length(block_list)) {
  block.i <- block_list[i]
  block_mover <- subset(block_distance, block_distance$Id2 == block.i)
  summary.i <- summary_index
  for (x in 1:nrow(block_mover)) {
    #Car <- block_mover$PCT_Vhcl_w[x]/100
    Car <- 1
    Tc <- block_mover$driving_morning[x]
    Tp <- block_mover$transit_morning[x]
    n <- block_mover$n[x]
    Ai <- (Car*(fxn(Tc,5)))*n
    block_mover$Access_Index[x] <- Ai
  }
  if( i %% 200 == 0){
      print(paste(i,x,"working..."))}
  #block_mover$Access_Index <- rescale(block_mover$Access_Index,to = c(0, 100))
  access_indexices <- rbind(block_mover,access_indexices)
  summary_index$Access_Index[i] <- mean(block_mover$Access_Index,na.rm=TRUE)})
```

###Plotting the Access Index and number of programs
Note that you need to have the reschool programs data set defined as "reschool_programs" for this to work. 
```{r}
library(leaflet)
library(RColorBrewer)
#ensure ids are one:one match
block_distance$Id2 <- as.character(block_distance$Id2)
summary_index$blockID <- as.character(summary_index$blockID)
block_distance$Id2 <- paste0("0", block_distance$Id2)
summary_index$blockID <- paste0("0",summary_index$blockID)

#Ordering
census_blocks <- census_blocks[order(census_blocks$Id2),]
block_distance <- block_distance[order(block_distance$Id2),]
summary_index <- summary_index[order(summary_index$blockID),] 

colourCount <- length(unique(total_programs$n))
getPalette <- colorRampPalette(brewer.pal(9,"Reds"))

pal <- colorNumeric(
  palette = "Blues",
  domain = summary_index$Access_Index)

pal2 <- colorNumeric(
  palette = rev(getPalette(colourCount)),
  domain = range(total_programs$n))

info <- paste("n",total_programs$n)

Ai_leaflet <- leaflet(census_blocks) %>%
  addTiles() %>%
  addPolygons(color = "Blues",weight =1,smoothFactor = 0.5,opacity = 0.5,fillOpacity = 0.75,
              fillColor = ~pal(summary_index$Access_Index),
              popup=paste(as.numeric(summary_index$blockID),summary_index$Access_Index,as.numeric(census_blocks$Id2)),
              highlightOptions = highlightOptions(color = "white", weight = 0.1)) %>%
  addLegend(pal = pal, values = ~summary_index$Access_Index, opacity = 1,title = "Access Index Score") %>%
  
  addCircles(data = total_programs, 
             opacity = 1,weight=7, popup=info,
             color = ~pal2(total_programs$n) ) %>%
  setView(lat=39.7,lng=-104.9,zoom=11)

#%>%
  #addCircles(data = census_centroids, opacity = 0.4,
             #weight=7,color="purple",popup=census_centroids$blockID) #originally we included centroids for validation, now we don't need them
```
