---
title: "Access_Index_Calculator"
author: "Andrew Taylor"
date: "7/19/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Intro
This notebook takes the block group distances, adds in additional data needed to calculate the Access Index, and then as of 07/20/18, shows our loops for running the equation on the access index for **overall** programs. **NOTE** that the "block_distance" data frame is already aggregated, so there's no need to run the first few code chunks to merge the dataframe. **HOWEVER** the plots at the end make use of the aggregated dataframes used to merge into the block_distance data-frame, so use your judgement. 

###Aggregate number of programs per unique address
```{r}
library(tidyverse)
#total sessions
reschool_programs$n <- 1
total_programs <- aggregate(n ~ session_address_1 + lat + long, data=reschool_programs, sum)

#sessions by category
#re-do for each discrete, I KNOW IT SUCKS, but you got a better idea?
#creating aggregated categories
reschool_programs$academic <- ifelse(reschool_programs$has_academic==TRUE | reschool_programs$has_stem==TRUE,TRUE,FALSE) 
reschool_programs$art <- ifelse(reschool_programs$has_arts==TRUE | reschool_programs$has_cooking==TRUE | 
                                  reschool_programs$has_dance==TRUE | reschool_programs$has_drama==TRUE
                                | reschool_programs$has_music==TRUE,
                                TRUE,FALSE)

#aggregating for agg categories and discrete categories
programs_academic <- aggregate(n ~ session_address_1+academic, data=reschool_programs,sum)
programs_arts <- aggregate(n ~ session_address_1+art, data=reschool_programs,sum)
programs_sports <- aggregate(n ~ session_address_1+has_sports, data=reschool_programs,sum)
programs_nature <- aggregate(n ~ session_address_1+has_nature, data=reschool_programs,sum)

programs_academic <- subset(programs_academic,programs_academic$academic==TRUE)
programs_arts <- subset(programs_arts, programs_arts$art==TRUE)
programs_sports <- subset(programs_sports, programs_sports$has_sports==TRUE)
programs_nature <- subset(programs_nature, programs_nature$has_nature==TRUE)

#trimming and creating unique colnames for merge 
colnames(programs_academic)[colnames(programs_academic)=="n"] <- "n_academic"
programs_academic <- programs_academic %>% select(n_academic, session_address_1)

colnames(programs_arts)[colnames(programs_arts)=="n"] <- "n_arts"
programs_arts <- programs_arts %>% select(n_arts, session_address_1)

colnames(programs_sports)[colnames(programs_sports)=="n"] <- "n_sports"
programs_sports <- programs_sports %>% select(n_sports, session_address_1)

colnames(programs_nature)[colnames(programs_nature)=="n"] <- "n_nature"
programs_nature <- programs_nature %>% select(n_nature, session_address_1)

#Create a car owners data frame to merge
#note we could do this in one step from the full census data.frame, but this is much simpler and less error prone
census_car_owners <- as.data.frame(census_blocks)
colnames(census_car_owners)[colnames(census_car_owners)=="INTPTLA"] <- "lat"  # for consistency across files
colnames(census_car_owners)[colnames(census_car_owners)=="INTPTLO"] <- "long"
census_car_owners <- census_car_owners %>% select(Id2, PCT_Vhcl_w, lat, long)
```

###Merge in additional data
```{r}
#merging total programs
block_distance <- merge(block_distance, total_programs, by="session_address_1")

#programs by category
block_distance <- merge(block_distance,programs_academic,all.x=TRUE)
block_distance <- merge(block_distance,programs_arts,all.x=TRUE)
block_distance <- merge(block_distance,programs_nature,all.x=TRUE)
block_distance <- merge(block_distance,programs_sports,all.x=TRUE)

#nas to 0s
block_distance$n_academic[is.na(block_distance$n_academic)==TRUE] <- 0
block_distance$n_arts[is.na(block_distance$n_arts)==TRUE] <- 0
block_distance$n_nature[is.na(block_distance$n_nature)==TRUE] <- 0
block_distance$n_sports[is.na(block_distance$n_sports)==TRUE] <- 0

#hit a strange bug here, where some census centroid IDs lost there suffix or prefix 0s. Converting them to numeric is a work around, unit tested that they are still unique, and they are. 
#For validation, we now also merge in census centroid lat/lons to this data frame in case the Id2 will break future merges
block_distance$Id2 <- as.numeric(block_distance$Id2) 
census_car_owners$Id2 <- as.numeric(census_car_owners$Id2)
block_distance <- merge(block_distance, census_car_owners, by="Id2")

#update names for new lat/long combos
colnames(block_distance)[colnames(block_distance)=="lat.x"] <- "lat"
colnames(block_distance)[colnames(block_distance)=="long.x"] <- "long"
colnames(block_distance)[colnames(block_distance)=="lat.y"] <- "block_lat"
colnames(block_distance)[colnames(block_distance)=="long.y"] <- "block_long"
block_distance$Id2 <- as.character(block_distance$Id2) #convert Ids back to characters

#final removal of duplictes, double check nothing messed around from merging
block_list <- unique(block_distance$Id2)
```

###Gravity Access index loop
Here we write a loop to calculate an access index score per each unique address, rescale the results per address, store each unique address result in a data frame in the style of block distances, and store the average access index in a summary dataframe **at the block group level**. This loop uses the "gravity" model, where k = 1.5. 
```{r}
library(scales)
block_mover$Access_Index <- 0
access_indexices <- as.data.frame(setNames(replicate(16,character(0), simplify = F), colnames(block_distance))) #make an empty data frame identical to block distances
summary_index <- census_centroids
summary_index$Access_Index <- 0

for (i in 1:length(block_list)){
  block.i <- block_list[i]
  block_mover <- subset(block_distance,block_distance$Id2==block.i)
  for (x in 1:nrow(block_mover)){
    Car <- block_mover$PCT_Vhcl_w[x]/100
    Tc <- block_mover$driving_morning[x]
    Tp <- block_mover$transit_morning[x]
    k <- 1.5
    n <- block_mover$n[x]
    Ai <- ifelse(is.na(Tp)==FALSE,((Car*(1/Tc^k)+(1-Car)*(1/Tp^k))*n),1*((1/Tc^k)*n)) #note here, there are some locations with no transit access, for these we assume only access by transit, hence if transit=NA, the equation is equaly to 100% driving access
    block_mover$Access_Index[x] <- Ai
    if( x %% 100 == 0){
      print(paste(i,x,"working..."))}
    }
  block_mover$Access_Index <- rescale(block_mover$Access_Index,to = c(0, 100))
  access_indexices <- rbind(block_mover,access_indexices)
  summary_index$Access_Index[i] <- mean(block_mover$Access_Index)}

#block n = 80319801001, has no info on % of car ownership. Thinking drop it
```

###Exponential Function 
Defined for use in the exponential function model for decay. 
```{r}
fxn <- function(vector, scale){
  2^(-vector/scale)}

driving_commute_denver <- 10 #wow these seem arbitrary huh?
transit_commute_denver <- 20 #HUH?
```

###Exp Access index loop
Here we use a loop to calculate the access index in the same fashion as above, using a different decay function. 
```{r}
library(scales)
block_list <- unique(block_distance$Id2)
block_mover$Access_Index <- 0
access_indexices <- as.data.frame(setNames(replicate(16,character(0), simplify = F), colnames(block_distance))) #make an empty data frame identical to block distances
summary_index <- census_centroids
summary_index$Access_Index <- 0

for (i in 1:length(block_list)){
  block.i <- block_list[i]
  block_mover <- subset(block_distance,block_distance$Id2==block.i)
  for (x in 1:nrow(block_mover)){
    Car <- block_mover$PCT_Vhcl_w[x]/100
    Tc <- block_mover$driving_morning[x]
    Tp <- block_mover$transit_morning[x]
    n <- block_mover$n[x]
    Ai <- ifelse(is.na(Tp)==FALSE,((Car*(fxn(Tc,driving_commute_denver))+(1-Car)*(fxn(Tp,transit_commute_denver)))*n),1*((fxn(Tc,transit_commute_denver))*n))
    block_mover$Access_Index[x] <- Ai
    if( i %% 200 == 0){
      print(paste(i,x,"working..."))}
    }
  #block_mover$Access_Index <- rescale(block_mover$Access_Index,to = c(0, 100)) #here we comment out rescaling for to test validation
  access_indexices <- rbind(block_mover,access_indexices)
  summary_index$Access_Index[i] <- mean(block_mover$Access_Index)}
```

###Exp Access Index Loop Simplified, return max
This is the same as the above but we no longer multiply by the number of programs at each site, and we return the max Ai value not the average for all destinations. The idea here being we want to distance **just** the role of travel times to locations to validate if we've written our equation correctly. In particular, we are concernced about a few block groups with very unequal access, despite being adjacent, (see the next chunk for plotting). Spot checking these block groups confirms that travel time calculcations appear correct. 

HERES THE F******* PROBLEM
1.) even when we only return max for plotting, we see block groups that seem confusing
2.) this may be due to the threshold we've picked, verify with alt thresholds
3.) conversly it may be a problem the visualization, though this seems unlikely
4.) samsies- let's all pray it's not a problem with the travel times. Fucking can't see why it should be, but we just don't know hey. 

```{r}
block_list <- unique(block_distance$Id2)
block_mover$Access_Index <- 0
access_indexices <- as.data.frame(setNames(replicate(16,character(0), simplify = F), colnames(block_distance))) #make an empty data frame identical to block distances
summary_index <- census_centroids
summary_index$Access_Index <- 0

for (i in 1:length(block_list)){
  block.i <- block_list[i]
  block_mover <- subset(block_distance,block_distance$Id2==block.i)
  for (x in 1:nrow(block_mover)){
    #Car <- block_mover$PCT_Vhcl_w[x]/100
    Car <- 1
    Tc <- block_mover$driving_morning[x]
    Tp <- block_mover$transit_morning[x]
    n <- block_mover$n[x]
    Ai <- Car*(fxn(Tc,driving_commute_denver))
    block_mover$Access_Index[x] <- Ai
    if( i %% 200 == 0){
      print(paste(i,x,"working..."))}
    }
  #block_mover$Access_Index <- rescale(block_mover$Access_Index,to = c(0, 100))
  access_indexices <- rbind(block_mover,access_indexices)
  summary_index$Access_Index[i] <- max(block_mover$Access_Index)}
```

###Plotting the Access Index and number of programs
Note that you need to have the reschool programs data set defined as "reschool_programs" for this to work. 
```{r}
library(leaflet)
library(RColorBrewer)
pal <- colorNumeric(
  palette = "Blues",
  domain = summary_index$Access_Index)

info <- paste("n",total_programs$n)

new <- leaflet(census_blocks) %>%
  addTiles() %>%
  addPolygons(color = "Blues",weight =1,smoothFactor = 0.5,opacity = 0.5,fillOpacity = 0.75,
              fillColor = ~pal(summary_index$Access_Index),
              popup=census_blocks$Id2,
              highlightOptions = highlightOptions(color = "white", weight = 0.1)) %>%
  addLegend(pal = pal, values = ~summary_index$Access_Index, opacity = 1) %>%
  addCircles(data = total_programs, opacity = 0.4,weight=10, color="red",popup=info) %>%
 setView(lat=39.7,lng=-104.9,zoom=10)
```

###Leaflet example using the access_indexices data frame
```{r}
pal <- colorNumeric(
  palette = "Blues",
  domain = summary_index$Access_Index)

info <- paste("n",access_indexices$n)

test <- leaflet(census_blocks) %>%
  addTiles() %>%
  addPolygons(color = "Blues",weight =1,smoothFactor = 0.5,opacity = 0.5,fillOpacity = 1,
              fillColor = ~pal(summary_index$Access_Index),
              popup=census_blocks$Id2,
              highlightOptions = highlightOptions(color = "white", weight = 0.1)) %>%
  addLegend(pal = pal, values = ~summary_index$Access_Index, opacity = 1) %>%
  addCircles(data = access_indexices,lat= access_indexices$lat,
             lng= access_indexices$long,
             opacity = 0.4,weight=10, color="red",
             popup=info) %>%
 setView(lat=39.7,lng=-104.9,zoom=10)
```

