---
title: "Denver_Crime_Reprocessing"
author: "Andrew Taylor"
date: "7/31/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###RDS Pull
```{r}
denver_crimes <- dbGetQuery(con, "select * from clean.crimes")
denver_hate_crimes <- dbGetQuery(con, "select * from clean.hate_crimes")
acs_demographics <- dbGetQuery(con, "select * from clean.acs_demographics")
```

###Aggregate by crime category
```{r}
denver_crimes$n <- 1
denver_crimes_agg <- aggregate(n ~ category+lat+long,data=denver_crimes,FUN=sum)
```

###Hard Code Fix coords to block groups function
```{r}
coords_to_blockgroups <- function(df) {
  # get block group polygons
  #setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
  mypath <- getwd()
  dataPath <- file.path(dirname(mypath), "osr_dssg2018-3", "data", "census_clean", "shape_census")
  blockGroups <- readOGR(dsn = "/Users/Andrew/osr_dssg2018-3/data/census_clean/shape_census") #can't seem to read the path thing 
  
  # get the lat / long coordinates to code as spatial points
  coords <- data.frame(df$long, df$lat)
  colnames(coords) <- c("long", "lat")
  spdf <- SpatialPoints(coords, proj4string = CRS(proj4string(blockGroups)))
  
  # figure out which block group each lat / long is in
  nests <- gIntersects(spdf, blockGroups, byid = TRUE)  # rows = bgroups, cols = df entries
  true_idxs <- which(nests==TRUE, arr.ind=TRUE)  # col1 = bgroup idx, col2 = df entries
  bgroup_idxs  <- true_idxs[,1]
  df_idxs <- true_idxs[,2]
  
  # make a dataframe with the coordinates and corresponding block groups
  df_bgroups <- data.frame(coords[df_idxs,], blockGroups$Id2[bgroup_idxs])
  colnames(df_bgroups) <- c("long", "lat", "bgroup_id2")
  
  # strip leading 0's in block groups numbers
  df_bgroups$bgroup_id2 <- as.character(as.numeric(as.character(df_bgroups$bgroup_id2))) 
  
  df_out <- merge(df, df_bgroups)
  return(df_out)
}
```

###Lat Long to blockgroups
```{r}
library(rgeos)
library(rgdal)
library(raster)
library(tidyverse)
#denver all crimes
denver_crimes$n <- 1
denver_crimes_agg <- aggregate(n ~ category+lat+long,data=denver_crimes,FUN=sum)
denver_crimes_agg <- coords_to_blockgroups(denver_crimes_agg)
denver_crimes_agg <- aggregate(n ~ bgroup_id2+category,data=denver_crimes_agg,FUN=sum)
colnames(denver_crimes_agg)[colnames(denver_crimes_agg)=="n"] <- "n_crimes"

#denver hate crimes
denver_hate_crimes$n <- 1
denver_hate_crimes_agg <- aggregate(n ~ long+lat+bias_type,data=denver_hate_crimes,FUN=sum)
denver_hate_crimes_agg <- coords_to_blockgroups(denver_hate_crimes_agg)
denver_hate_crimes_agg <- aggregate(n ~ bgroup_id2+bias_type,data=denver_hate_crimes_agg,FUN=sum)
colnames(denver_hate_crimes_agg)[colnames(denver_hate_crimes_agg)=="n"] <- "hate_crimes"
```

###Spread Long Dataframes
```{r}
test <- denver_hate_crimes_agg
test <- reshape(test, idvar=c("bgroup_id2"),timevar="bias_type",direction="wide")
head(test)
denver_hate_crimes_agg <- test
denver_hate_crimes_agg[is.na(denver_hate_crimes_agg)] <- 0
denver_hate_crimes_agg$total_hate <- rowSums(denver_hate_crimes_agg[,c(2:length(denver_hate_crimes_agg))])

test <- denver_crimes_agg
test <- reshape(test, idvar=c("bgroup_id2"),timevar="category",direction="wide")
head(test)
denver_crimes_agg <- test
denver_crimes_agg[is.na(denver_crimes_agg)] <- 0
denver_crimes_agg$total_crime <- rowSums(denver_crimes_agg[,c(2:length(denver_crimes_agg))])
```

###Merge Hate & All Crime 
```{r}
total_crimes <- merge(denver_crimes_agg,denver_hate_crimes_agg,by="bgroup_id2",all.x=TRUE)
total_crimes[is.na(total_crimes)] <- 0
```

###Merge Demographics for per 100,000 rates
```{r}
acs_pop <- acs_demographics[c("ttl_population","Id2")]
acs_pop$Id2 <- as.character(acs_pop$Id2)
colnames(total_crimes)[colnames(total_crimes)=="bgroup_id2"] <- "Id2"
total_crimes <- merge(total_crimes,acs_pop,by="Id2")
```

###Per 100 Population Rates
Big difference in scale here

A.) Crime is total per five years + current to 7/31. so divided by 5.583333

B.) Hate crimes is MUCH broader, 8.686813 years worth of data, based on the range from first to last date. 

So here we divide by population, multiply by 100 to get the per 100 rate, then divide by the n of years (including the aprox n of months)

Note: This returns one NA, because there is a block group with population. 
```{r}
crime_rates <- total_crimes
crime_rates[,2:15] <- ((total_crimes[,2:15]/total_crimes$ttl_population)*100)/5.58333
crime_rates[,16:length(crime_rates)] <- ((total_crimes[,16:length(crime_rates)]/total_crimes$ttl_population)*100)/8.6868
crime_rates <- subset(crime_rates,crime_rates$ttl_population!=0)
```

###Adding Aggregate Crimes
Violent Crime = Aggregate of murder, robbery, aggrvated assault rates (verify FBI incident status, I think there's some sexual assualt esq charge that goes into this?)
Property Crime = Aggregte of robbery, burglary, theft from motor vehicle (car prowling), larceny, (again, verify FBI)
```{r}
crime_rates$violent_crime <- rowSums(crime_rates[,c("n_crimes.murder","n_crimes.robbery","n_crimes.aggravated-assault")])
crime_rates$property_crime <- rowSums(crime_rates[,c("n_crimes.robbery","n_crimes.burglary","n_crimes.theft-from-motor-vehicle","n_crimes.larceny")])
```


###Validation visualizations
```{r}
library(leaflet)
library(RColorBrewer)

#order shit
census_blocks <- census_blocks[order(census_blocks$Id2),]
crime_rates <- crime_rates[order(crime_rates$Id2),]

#Pal
range <- range(crime_rates$total_crime)
pal <- colorNumeric(palette = "Reds",domain = range)

#Plot Overall
Ai_crime_overall <- leaflet(census_blocks) %>%
  addTiles() %>%
  addPolygons(color = "RdYlBu",weight =1,smoothFactor = 0.5,opacity = 0.5,fillOpacity = 0.9,
              fillColor = ~pal(crime_rates$total_crime),
              popup=paste("Crime Rate =",crime_rates$total_crime),
              highlightOptions = highlightOptions(color = "white", weight = 0.1)) %>%
  addLegend(pal = pal, values = ~crime_rates$total_crime, opacity = 1, title = "Crime Rate Per 100 Residents") %>%
  setView(lat=39.7,lng=-104.9,zoom=11.5)

#Plot Violent
range <- range(crime_rates$violent_crime)
pal <- colorNumeric(palette = "Reds",domain = range)

Ai_crime_violent <- leaflet(census_blocks) %>%
  addTiles() %>%
  addPolygons(color = "RdYlBu",weight =1,smoothFactor = 0.5,opacity = 0.5,fillOpacity = 0.9,
              fillColor = ~pal(crime_rates$violent_crime),
              popup=paste("Crime Rate =",crime_rates$violent_crime),
              highlightOptions = highlightOptions(color = "white", weight = 0.1)) %>%
  addLegend(pal = pal, values = ~crime_rates$violent_crime, opacity = 1, title = "Violent Crime Rate Per 100 Residents") %>%
  setView(lat=39.7,lng=-104.9,zoom=11.5)

#Plot Property
range <- range(crime_rates$property_crime)
pal <- colorNumeric(palette = "Reds",domain = range)

Ai_crime_property <- leaflet(census_blocks) %>%
  addTiles() %>%
  addPolygons(color = "RdYlBu",weight =1,smoothFactor = 0.5,opacity = 0.5,fillOpacity = 0.9,
              fillColor = ~pal(crime_rates$property_crime),
              popup=paste("Crime Rate =",crime_rates$property_crime),
              highlightOptions = highlightOptions(color = "white", weight = 0.1)) %>%
  addLegend(pal = pal, values = ~crime_rates$property_crime, opacity = 1, title = "Property Crime Rate Per 100 Residents") %>%
  setView(lat=39.7,lng=-104.9,zoom=11.5)

```

