---
title: "Denver_Crime_Reprocessing"
author: "Andrew Taylor"
date: "7/31/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###RDS Pull
```{r}
denver_crimes <- dbGetQuery(con, "select * from clean.crimes")
denver_hate_crimes <- dbGetQuery(con, "select * from clean.hate_crimes")
acs_demographics <- dbGetQuery(con, "select * from clean.acs_demographics")
```

###Aggregate by crime category
```{r}
denver_crimes$n <- 1
denver_crimes_agg <- aggregate(n ~ category+lat+long,data=denver_crimes,FUN=sum)
```

###Lat Long to blockgroups
```{r}
library(rgeos)
library(rgdal)
library(raster)
library(tidyverse)
#denver all crimes
denver_crimes$n <- 1
denver_crimes_agg <- aggregate(n ~ category+lat+long,data=denver_crimes,FUN=sum)
denver_crimes_agg <- coords_to_blockgroups(denver_crimes_agg)
denver_crimes_agg <- aggregate(n ~ bgroup_id2+category,data=denver_crimes_agg,FUN=sum)
colnames(denver_crimes_agg)[colnames(denver_crimes_agg)=="n"] <- "n_crimes"

#denver hate crimes
denver_hate_crimes$n <- 1
denver_hate_crimes_agg <- aggregate(n ~ long+lat+bias_type,data=denver_hate_crimes,FUN=sum)
denver_hate_crimes_agg <- coords_to_blockgroups(denver_hate_crimes_agg)
denver_hate_crimes_agg <- aggregate(n ~ bgroup_id2+bias_type,data=denver_hate_crimes_agg,FUN=sum)
colnames(denver_hate_crimes_agg)[colnames(denver_hate_crimes_agg)=="n"] <- "hate_crimes"
```

###Spread Long Dataframes
```{r}
test <- denver_hate_crimes_agg
test <- reshape(test, idvar=c("bgroup_id2"),timevar="bias_type",direction="wide")
head(test)
denver_hate_crimes_agg <- test
denver_hate_crimes_agg[is.na(denver_hate_crimes_agg)] <- 0
denver_hate_crimes_agg$total_hate <- rowSums(denver_hate_crimes_agg[,c(2:length(denver_hate_crimes_agg))])

test <- denver_crimes_agg
test <- reshape(test, idvar=c("bgroup_id2"),timevar="category",direction="wide")
head(test)
denver_crimes_agg <- test
denver_crimes_agg[is.na(denver_crimes_agg)] <- 0
denver_crimes_agg$total_crime <- rowSums(denver_crimes_agg[,c(2:length(denver_crimes_agg))])
```

###Merge Hate & All Crime 
```{r}
total_crimes <- merge(denver_crimes_agg,denver_hate_crimes_agg,by="bgroup_id2",all.x=TRUE)
total_crimes[is.na(total_crimes)] <- 0
```

###Merge Demographics for per 100,000 rates
```{r}
acs_pop <- acs_demographics[c("ttl_population","Id2")]
acs_pop$Id2 <- as.character(acs_pop$Id2)
colnames(total_crimes)[colnames(total_crimes)=="bgroup_id2"] <- "Id2"
total_crimes <- merge(total_crimes,acs_pop,by="Id2")
```

###Per 100 Population Rates
Big difference in scale here

A.) Crime is total per five years + current to 7/31. so divided by 5.583333

B.) Hate crimes is MUCH broader, 8.686813 years worth of data, based on the range from first to last date. 

So here we divide by population, multiply by 100 to get the per 100 rate, then divide by the n of years (including the aprox n of months)

```{r}
crime_rates <- total_crimes
crime_rates[,2:15] <- ((total_crimes[,2:15]/total_crimes$ttl_population)*100)/5.58333
crime_rates[,16:length(crime_rates)] <- ((total_crimes[,16:length(crime_rates)]/total_crimes$ttl_population)*100)/8.6868
crime_rates <- subset(crime_rates,crime_rates$ttl_population!=0)
```

###Adding Aggregate Crimes
Violent Crime = murder, nonnegligent manslaughter, robbery, aggravated assault, burglary, larceny-theft, and motor vehicle theft.

Note that sexual assualt charges are not listed in the Denver Open Data, potentially out of concern for the victim. Therefore our violent crime defintion is one category short of the FBI definition, but a very close approximation nonetheless. 

Property Crime = burglary, larceny-theft, motor vehcile theft, arson

Identical to FBI definition. 

Because there's rollover here there's more or less no difference in areas with high violent & high property crime. 
```{r}
crime_rates$violent_crime <- rowSums(crime_rates[,c("n_crimes.murder","n_crimes.robbery","n_crimes.aggravated-assault","n_crimes.burglary","n_crimes.larceny","n_crimes.auto-theft","n_crimes.theft-from-motor-vehicle","n_crimes.auto-theft")])
crime_rates$property_crime <- rowSums(crime_rates[,c("n_crimes.burglary","n_crimes.theft-from-motor-vehicle","n_crimes.larceny","n_crimes.auto-theft","n_crimes.arson")])
```

###Block Groups to Neighborhoods
Small problem with block group level population crime rates. There are few, but noticable block groups with more crimes than people, which leads to a pretty weird interpretation. This problem persisits with neighborhoods, but hey, what can you do. 
```{r}
#source("open_data_functions.R")
#source("block_groups_to_nbhds.R")

#Prep data for neighborhoods
colnames(dfFinal)[colnames(dfFinal)=="bgroup_id2"] <- "Id2"
denver_crimes_nbhd <- aggregate(n ~ category+lat+long,data=denver_crimes,FUN=sum)
denver_crimes_nbhd <- coords_to_blockgroups(denver_crimes_nbhd)
colnames(denver_crimes_nbhd)[colnames(denver_crimes_nbhd)=="bgroup_id2"] <- "Id2"

#merge and reshape crime neigbhorhood level
crime_rates_nbhd <- merge(denver_crimes_nbhd,dfFinal,by="Id2") 
crime_rates_nbhd <- aggregate(n ~ category+nbhd_id+nbhd_name,data=crime_rates_nbhd,FUN=sum)
crime_rates_nbhd <- reshape(crime_rates_nbhd, idvar=c("nbhd_id","nbhd_name"),timevar="category",direction="wide")
crime_rates_nbhd[is.na(crime_rates_nbhd)] <- 0

#get neighborhood population
nbhd_pop <- merge(acs_demographics,dfFinal,by="Id2")
nbhd_pop <- aggregate(ttl_population ~ nbhd_id+nbhd_name,data=nbhd_pop,FUN=sum)

#make aggregate crimes
crime_rates_nbhd$violent_crimes <- rowSums(crime_rates_nbhd[,c("n.murder","n.robbery","n.aggravated-assault","n.burglary","n.larceny","n.auto-theft","n.theft-from-motor-vehicle","n.auto-theft")])
crime_rates_nbhd$property_crimes <- rowSums(crime_rates_nbhd[,c("n.burglary","n.theft-from-motor-vehicle","n.larceny","n.auto-theft","n.arson")])

#population rates per 100 rates average annual rates
crime_rates_nbhd <- merge(crime_rates_nbhd,nbhd_pop,by=c("nbhd_name","nbhd_id"))
crime_rates_nbhd[,3:(length(crime_rates_nbhd)-1)] <- ((crime_rates_nbhd[,3:(length(crime_rates_nbhd)-1)] / crime_rates_nbhd$ttl_population)*100)/5.5833
```


###Push to RDS
```{r}
require("RPostgreSQL")

# loads the PostgreSQL driver
drv <- dbDriver("PostgreSQL")

# creates a connection to the postgres database
# note that "con" will be used later in each connection to the database
con <- dbConnect(drv, dbname = "dssg2018uw",
                 host = "localhost", port = 9000,
                 user = user, password = password)

# specify data and schema and table names
data <- crime_rates  # must be already loaded in your environment
schemaName <- "clean"
tableName <- "crime_rates"

dbWriteTable(con,
             c(schemaName, tableName),
             value = data,
             row.names = FALSE,
             overwrite = TRUE  # overwrite an existing table
)

data <- crime_rates_nbhd
schemaName <- "clean"
tableName <- "crime_rates_nbhd"

dbWriteTable(con,
             c(schemaName, tableName),
             value = data,
             row.names = FALSE,
             overwrite = TRUE  # overwrite an existing table
)

# when you're done, close the connection and unload the driver 
dbDisconnect(con) 
dbUnloadDriver(drv)
```

###Block Group Validation visualizations
```{r}
library(leaflet)
library(RColorBrewer)

#order shit
census_blocks <- census_blocks[order(census_blocks$Id2),]
crime_rates <- crime_rates[order(crime_rates$Id2),]

#Pal
range <- range(crime_rates$total_crime)
pal <- colorNumeric(palette = "Reds",domain = range)

#Plot Overall
bg_crime_overall <- leaflet(census_blocks) %>%
  addTiles() %>%
  addPolygons(color = "RdYlBu",weight =1,smoothFactor = 0.5,opacity = 0.5,fillOpacity = 0.9,
              fillColor = ~pal(crime_rates$total_crime),
              popup=paste("Crime Rate =",crime_rates$total_crime),
              highlightOptions = highlightOptions(color = "white", weight = 0.1)) %>%
  addLegend(pal = pal, values = ~crime_rates$total_crime, opacity = 1, title = "Crime Rate Per 100 Residents") %>%
  setView(lat=39.7,lng=-104.9,zoom=11.5)

#Plot Violent
range <- range(crime_rates$violent_crime)
pal <- colorNumeric(palette = "Reds",domain = range)

bg_crime_violent <- leaflet(census_blocks) %>%
  addTiles() %>%
  addPolygons(color = "RdYlBu",weight =1,smoothFactor = 0.5,opacity = 0.5,fillOpacity = 0.9,
              fillColor = ~pal(crime_rates$violent_crime),
              popup=paste("Crime Rate =",crime_rates$violent_crime),
              highlightOptions = highlightOptions(color = "white", weight = 0.1)) %>%
  addLegend(pal = pal, values = ~crime_rates$violent_crime, opacity = 1, title = "Violent Crime Rate Per 100 Residents") %>%
  setView(lat=39.7,lng=-104.9,zoom=11.5)

#Plot Property
range <- range(crime_rates$property_crime)
pal <- colorNumeric(palette = "Reds",domain = range)

bg_crime_property <- leaflet(census_blocks) %>%
  addTiles() %>%
  addPolygons(color = "RdYlBu",weight =1,smoothFactor = 0.5,opacity = 0.5,fillOpacity = 0.9,
              fillColor = ~pal(crime_rates$property_crime),
              popup=paste("Crime Rate =",crime_rates$property_crime),
              highlightOptions = highlightOptions(color = "white", weight = 0.1)) %>%
  addLegend(pal = pal, values = ~crime_rates$property_crime, opacity = 1, title = "Property Crime Rate Per 100 Residents") %>%
  setView(lat=39.7,lng=-104.9,zoom=11.5)

```

###Neighborhood Validation Visuals
There's still, downtown neighborhoods in particular, where n of crimes > n of people 
```{r}
crime_rates_nbhd <- crime_rates_nbhd[order(crime_rates_nbhd$nbhd_id),]
nbhds <- nbhds[order(nbhds$NBHD_ID),]

#Violent
range <- range(crime_rates_nbhd$violent_crime)
pal <- colorNumeric(palette = "Reds",domain = range)

nbhd_crime_violent <- leaflet(nbhds) %>%
  addTiles() %>%
  addPolygons(color = "White",weight=5,smoothFactor = 0.5,opacity = 1,fillOpacity = 0.9,
              fillColor = ~pal(crime_rates_nbhd$violent_crime),
              popup=paste(round(crime_rates_nbhd$violent_crime,digits=0),crime_rates_nbhd$nbhd_name),
              highlightOptions = highlightOptions(color = "white", weight = 0.1)) %>%
  addLegend(pal = pal, values = ~crime_rates_nbhd$violent_crime, opacity = 1, title = "Violent Crime Rate Per 100 Residents") %>%
  setView(lat=39.7,lng=-104.9,zoom=11.5)

#property
range <- range(crime_rates_nbhd$property_crimes)
pal <- colorNumeric(palette = "Reds",domain = range)

nbhd_crime_property <- leaflet(nbhds) %>%
  addTiles() %>%
  addPolygons(color = "White",weight=5,smoothFactor = 0.5,opacity = 1,fillOpacity = 0.9,
              fillColor = ~pal(crime_rates_nbhd$property_crimes),
              popup=paste(round(crime_rates_nbhd$property_crimes,digits=0),crime_rates_nbhd$nbhd_name),
              highlightOptions = highlightOptions(color = "white", weight = 0.1)) %>%
  addLegend(pal = pal, values = ~crime_rates_nbhd$property_crimes, opacity = 1, title = "Property Crime Rate Per 100 Residents") %>%
  setView(lat=39.7,lng=-104.9,zoom=11.5)
```

