---
title: "Access_Index_SA_Slots"
author: "Andrew Taylor"
date: "8/3/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##By Number of Sessions
###Using a subset of sessions w/ slots
```{r}
#Merge sensitive
Merge_Set_Sensitive <- function(dataframe){
  relevant_columns <- c("Id2","session_address_1","driving_morning","walking_morning","transit_morning","kilometers","block_lat","block_long")
  block_distance_new <- block_distance[,relevant_columns] #here we're cutting duplicate program_addresses, and other antiquated info (previous aggregations, PCT vhcl ownership) that was use in previous calculations
  temp <- merge(block_distance_new,dataframe,all.y=TRUE)
  return(temp)
}

#make correct subset
reschool_programs_sa <- subset(reschool_programs,reschool_programs$session_size!=0)

#redfine agg set
Aggregate_Subset_Sensitive <- function(dataframe){
  return(aggregate(session_size ~ session_address_1+lat+long, data=dataframe,FUN=sum)) #note that "n" in this case ==1 per row, which n this case corresponds to one unique program session
}

#redefine soft gravity
decay_fxn_sensitiveGravity <- function(dataframe,mode){
  if (mode=="transit") {
    scale = 10
    column="transit_morning"
  }
  if (mode=="drive"){
    scale = 10
    column="driving_morning"
  }
  mode=mode
  dataframe[,"session_size"]*(1/(1+(dataframe[,column]/scale))^2)
}

#Create driving frame
driving_index <- base_frame

#Creating categories list for loop

categories <- c("academic","art","has_sports","has_nature")

#Creating cost list for loop

costs <- c("free","lowcost","anycost")

#Loop for all categories and costs

for (i in 1:length(categories)){
  type <- categories[i]
  for (x in 1:length(costs)){
    cost <- costs[x]
    user_set <- Make_Subset(reschool_programs_sa,type,cost)
    agg_set <- Aggregate_Subset(user_set)
    merged_set <- Merge_Set(agg_set)
    merged_set$AccessIndex <- decay_fxn_softGravity(merged_set,"drive") 
    summary_index_user <- aggregate(AccessIndex ~ Id2,data=merged_set,FUN=sum) 
    max_access <- max(summary_index_user$AccessIndex)
    #update colnames for merge
    summary_index_user$AccessIndex <- normalize(summary_index_user$AccessIndex,max_access)
    colnames(summary_index_user)[colnames(summary_index_user)=="AccessIndex"] <- paste("AI",type,cost,sep = "_") #flexible name writing for loop
    summary_index_user$Id2 <- as.numeric(summary_index_user$Id2) #matching str to baseframe 
    driving_index <- merge(summary_index_user,driving_index,by=c("Id2"))
  }
  print(paste(i,"working"))
  if(length(driving_index)>11){
    print(colnames(driving_index))
  }
}

#Create an overall average by category with no cost

driving_index$AI_overall <- rowMeans(driving_index[,c("AI_has_nature_anycost",'AI_has_sports_anycost','AI_art_anycost','AI_academic_anycost')])
driving_index$AI_overall_free <- rowMeans(driving_index[,c("AI_has_nature_free","AI_has_sports_free","AI_art_free","AI_academic_free")])
```

###Get Base Ranks
First we need orders for blockgroup ranks for every condition
```{r}
driving_index_sa <- driving_index
indecies <- colnames(driving_index[2:length(driving_index)]) #get the cols for our loop, substract the ID column

for (i in 1:length(indecies)){
  x <- indecies[i] #index col names for loop
  driving_index_sa <- driving_index_sa[order(driving_index_sa[paste0(x)]),] #order by indexed col name
  rownames(driving_index_sa) <- NULL #delete old row names (numbers)
  driving_index_sa$rank <- as.numeric(rownames(driving_index_sa)) #make "rank" = to order
  colnames(driving_index_sa)[colnames(driving_index_sa)=="rank"] <- paste0("rank_",x) #copy rank order to a new col
  print(paste("working...",i)) #verify working
  if (i >13){
    print(colnames(driving_index_sa)) #print colnames for verification
  }
}
```

##By Slots
###Subset of sessions w/ Slots
Next we need to recalculate the Access Index using slots instead of sessions, for the 50% of programs with slots. 
```{r}
#Create driving frame
driving_index_slots <- base_frame

#Creating categories list for loop

categories <- c("academic","art","has_sports","has_nature")

#Creating cost list for loop

costs <- c("free","lowcost","anycost")

#Loop for all categories and costs
for (i in 1:length(categories)){
  type <- categories[i]
  for (x in 1:length(costs)){
    cost <- costs[x]
    user_set <- Make_Subset(reschool_programs_sa,type,cost)
    agg_set <- Aggregate_Subset_Sensitive(user_set)
    merged_set <- Merge_Set_Sensitive(agg_set)
    merged_set$AccessIndex <- decay_fxn_sensitiveGravity(merged_set,"drive") 
    summary_index_user <- aggregate(AccessIndex ~ Id2,data=merged_set,FUN=sum) 
    max_access <- max(summary_index_user$AccessIndex)
    #update colnames for merge
    summary_index_user$AccessIndex <- normalize(summary_index_user$AccessIndex,max_access)
    colnames(summary_index_user)[colnames(summary_index_user)=="AccessIndex"] <- paste("SA_AI",type,cost,sep = "_") #flexible name writing for loop
    summary_index_user$Id2 <- as.numeric(summary_index_user$Id2) #matching str to baseframe 
    driving_index_slots <- merge(summary_index_user,driving_index_slots,by=c("Id2"))
  }
  print(paste(i,"working"))
  if(length(driving_index_slots)>11){
    print(colnames(driving_index_slots))
  }
}

#Create an overall average by category with no cost

driving_index_slots$SA_AI_overall <- rowMeans(driving_index_slots[,c("SA_AI_has_nature_anycost",'SA_AI_has_sports_anycost','SA_AI_art_anycost','SA_AI_academic_anycost')])
driving_index_slots$SA_AI_overall_free <- rowMeans(driving_index_slots[,c("SA_AI_has_nature_free","SA_AI_has_sports_free","SA_AI_art_free","SA_AI_academic_free")])
```

###Get Test Ranks
```{r}
driving_index_slots_sa <- driving_index_slots
indecies <- colnames(driving_index_slots_sa[2:length(driving_index_slots_sa)]) #get the cols for our loop, substract the ID column

for (i in 1:length(indecies)){
  x <- indecies[i] #index col names for loop
  driving_index_slots_sa <- driving_index_slots_sa[order(driving_index_slots_sa[paste0(x)]),] #order by indexed col name
  rownames(driving_index_slots_sa) <- NULL #delete old row names (numbers)
  driving_index_slots_sa$rank <- as.numeric(rownames(driving_index_slots_sa)) #make "rank" = to order
  colnames(driving_index_slots_sa)[colnames(driving_index_slots_sa)=="rank"] <- paste0("SA_rank_",x) #copy rank order to a new col
  print(paste("working...",i)) #verify working
  if (i >13){
    print(colnames(driving_index_slots_sa)) #print colnames for verification
  }
}
```

###Merge
```{r}
Slots_Test <- merge(driving_index_sa,driving_index_slots_sa,by="Id2")
colnames(Slots_Test)
```

###Plot Loop
```{r}
library(ggplot2)
library(gridExtra)
tests <- colnames(Slots_Test[-c(1:15,30:43)]) #subset to be only ranks, we don't care about scores here
for (i in 1:14){
  test <- tests[i]
  test_sa <- tests[i+14]
  p <- ggplot(Slots_Test, aes(y=Slots_Test[,test],x=Slots_Test[,test_sa])) +
         geom_point() +
         labs(title=paste("Rank Test",test)) + 
  ylab("AI Rank by Session") + 
  xlab("AI Rank by Slot")
  print(p)
  print(paste(i,"working..."))
}
```

##Same But For Transit
###Get Transit SA Frame
```{r}
#Create driving frame
transit_index <- base_frame

#Creating categories list for loop

categories <- c("academic","art","has_sports","has_nature")

#Creating cost list for loop

costs <- c("free","lowcost","anycost")

#Loop for all categories and costs

for (i in 1:length(categories)){
  type <- categories[i]
  for (x in 1:length(costs)){
    cost <- costs[x]
    user_set <- Make_Subset(reschool_programs_sa,type,cost)
    agg_set <- Aggregate_Subset(user_set)
    merged_set <- Merge_Set(agg_set)
    
    #replace na and 0 transit times with wallking times
    merged_set$transit_morning[is.na(merged_set$transit_morning)] <- merged_set$walking_morning[is.na(merged_set$transit_morning)] #there shouldn't be NAs or 0s
    merged_set$transit_morning[merged_set$transit_morning==0] <- merged_set$walking_morning[merged_set$transit_morning==0] #but if there is, as sometimes occurs with the API calls,        here we replace non-existant transit times with walking times, under the assumption the purpose of this index is to view access for individuals without a ca
    
    merged_set$AccessIndex <- decay_fxn_softGravity(merged_set,"transit") 
    merged_set$CarAccess <- decay_fxn_softGravity(merged_set,"drive")
    
    summary_index_user_transit_graphic <- aggregate(CarAccess ~ Id2,data=merged_set,FUN=sum)
    summary_index_user_transit <- aggregate(AccessIndex ~ Id2,data=merged_set,FUN=sum)

    #normalize, note here we make two columns, one to compare to driving, one for visual rep
    max_access <- max(summary_index_user_transit_graphic$CarAccess)
    summary_index_user_transit$AccessIndex <- normalize(summary_index_user_transit$AccessIndex,max_access)
    
    #update colnames for merge
    colnames(summary_index_user_transit)[colnames(summary_index_user_transit)=="AccessIndex"] <- paste("AI",type,cost,sep = "_") #flexible name writing for loop
    summary_index_user_transit$Id2 <- as.numeric(summary_index_user_transit$Id2) #matching str to baseframe 
    transit_index <- merge(summary_index_user_transit,transit_index,by=c("Id2"))
  }
  print(paste(i,"working"))
  if(length(transit_index)>11){
    print(colnames(transit_index))
  }
}

#Create an overall average by category with no cost

transit_index$AI_overall <- rowMeans(transit_index[,c("AI_has_nature_anycost",'AI_has_sports_anycost','AI_art_anycost','AI_academic_anycost')])
transit_index$AI_overall_free <- rowMeans(transit_index[,c("AI_has_nature_free","AI_has_sports_free","AI_art_free","AI_academic_free")])
```

###Get Transit Base Ranks
```{r}
transit_index_sa <- transit_index
indecies <- colnames(transit_index[2:length(transit_index)]) #get the cols for our loop, substract the ID column

for (i in 1:length(indecies)){
  x <- indecies[i] #index col names for loop
  transit_index_sa <- transit_index_sa[order(transit_index_sa[paste0(x)]),] #order by indexed col name
  rownames(transit_index_sa) <- NULL #delete old row names (numbers)
  transit_index_sa$rank <- as.numeric(rownames(transit_index_sa)) #make "rank" = to order
  colnames(transit_index_sa)[colnames(transit_index_sa)=="rank"] <- paste0("rank_",x) #copy rank order to a new col
  print(paste("working...",i)) #verify working
  if (i >13){
    print(colnames(transit_index_sa)) #print colnames for verification
  }
}
```

###Transit by Slots
```{r}
#Create driving frame
transit_index_slots <- base_frame

#Creating categories list for loop

categories <- c("academic","art","has_sports","has_nature")

#Creating cost list for loop

costs <- c("free","lowcost","anycost")

#Loop for all categories and costs

for (i in 1:length(categories)){
  type <- categories[i]
  for (x in 1:length(costs)){
    cost <- costs[x]
    user_set <- Make_Subset(reschool_programs_sa,type,cost)
    agg_set <- Aggregate_Subset_Sensitive(user_set)
    merged_set <- Merge_Set_Sensitive(agg_set)
    
    #replace na and 0 transit times with wallking times
    merged_set$transit_morning[is.na(merged_set$transit_morning)] <- merged_set$walking_morning[is.na(merged_set$transit_morning)] #there shouldn't be NAs or 0s
    merged_set$transit_morning[merged_set$transit_morning==0] <- merged_set$walking_morning[merged_set$transit_morning==0] #but if there is, as sometimes occurs with the API calls,        here we replace non-existant transit times with walking times, under the assumption the purpose of this index is to view access for individuals without a ca
    
    merged_set$AccessIndex <- decay_fxn_sensitiveGravity(merged_set,"transit") 
    merged_set$CarAccess <- decay_fxn_sensitiveGravity(merged_set,"drive")
    
    summary_index_user_transit_graphic <- aggregate(CarAccess ~ Id2,data=merged_set,FUN=sum)
    summary_index_user_transit <- aggregate(AccessIndex ~ Id2,data=merged_set,FUN=sum)

    #normalize, note here we make two columns, one to compare to driving, one for visual rep
    max_access <- max(summary_index_user_transit_graphic$CarAccess)
    summary_index_user_transit$AccessIndex <- normalize(summary_index_user_transit$AccessIndex,max_access)
    
    #update colnames for merge
    colnames(summary_index_user_transit)[colnames(summary_index_user_transit)=="AccessIndex"] <- paste("AI",type,cost,sep = "_") #flexible name writing for loop
    summary_index_user_transit$Id2 <- as.numeric(summary_index_user_transit$Id2) #matching str to baseframe 
    transit_index_slots <- merge(summary_index_user_transit,transit_index_slots,by=c("Id2"))
  }
  print(paste(i,"working"))
  if(length(transit_index_slots)>11){
    print(colnames(transit_index_slots))
  }
}

#Create an overall average by category with no cost

transit_index_slots$SA_AI_overall <- rowMeans(transit_index_slots[,c("AI_has_nature_anycost",'AI_has_sports_anycost','AI_art_anycost','AI_academic_anycost')])
transit_index_slots$SA_AI_overall_free <- rowMeans(transit_index_slots[,c("AI_has_nature_free","AI_has_sports_free","AI_art_free","AI_academic_free")])
```

###Transit Slots Ranks
```{r}
transit_index_slots_sa <- transit_index_slots
indecies <- colnames(transit_index_slots_sa[2:length(transit_index_slots_sa)]) #get the cols for our loop, substract the ID column

for (i in 1:length(indecies)){
  x <- indecies[i] #index col names for loop
  transit_index_slots_sa <- transit_index_slots_sa[order(transit_index_slots_sa[paste0(x)]),] #order by indexed col name
  rownames(transit_index_slots_sa) <- NULL #delete old row names (numbers)
  transit_index_slots_sa$rank <- as.numeric(rownames(transit_index_slots_sa)) #make "rank" = to order
  colnames(transit_index_slots_sa)[colnames(transit_index_slots_sa)=="rank"] <- paste0("Slots_rank_",x) #copy rank order to a new col
  print(paste("working...",i)) #verify working
  if (i >13){
    print(colnames(transit_index_slots_sa)) #print colnames for verification
  }
}
```

###Merge Transit
```{r}
Transit_Slots_Test <- merge(transit_index_sa,transit_index_slots_sa,by="Id2")
colnames(Transit_Slots_Test)
```

###Plot Loop
```{r}
library(ggplot2)
library(gridExtra)
tests <- colnames(Transit_Slots_Test[-c(1:15,30:43)]) #subset to be only ranks, we don't care about scores here
for (i in 1:14){
  test <- tests[i]
  test_sa <- tests[i+14]
  p <- ggplot(Transit_Slots_Test, aes(y=Transit_Slots_Test[,test],x=Transit_Slots_Test[,test_sa])) +
         geom_point() +
         labs(title=paste("Transit Rank Test",test)) + 
  ylab(paste("AI Rank by Session")) + 
  xlab(paste("AI Rank by Slot"))
  print(p)
  print(paste(i,"working..."))
}
```

###Local Moran's I, again
Do it to confirm that the same areas are clustered as in the base case, bro.
```{r}

```

###Leaflet
```{r}
library(RColorBrewer)
library(leaflet)
#Slots Agg
agg_slots <- aggregate(session_size ~ lat+long,data=reschool_programs_sa,FUN=sum)

#Ordering
census_blocks$Id2 <- as.numeric(census_blocks$Id2)
driving_index_sa$Id2 <- as.numeric(driving_index_sa$Id2)

#specifiying color pallet
colourCount <- length(unique(reschool_programs_sa$session_size))
getPalette <- colorRampPalette(brewer.pal(9,"Reds"))

range <- range(driving_index_sa$AI_overall)

pal <- colorNumeric(palette = "Blues",domain = range)

pal2 <- colorNumeric(
  palette = rev(getPalette(colourCount)),
  domain = range(agg_slots$session_size))

#indicating one popup info
info <- paste("n",reschool_programs$session_size)

#SA creating leaflet map
Ai_driving_SA <- leaflet(census_blocks) %>%
  addTiles() %>%
  addPolygons(color = "RdYlBu",weight =1,smoothFactor = 0.5,opacity = 0.5,fillOpacity = 0.75,
              fillColor = ~pal(driving_index_sa$AI_overall),
              popup=paste("AI =",round(driving_index_sa$AI_overall,digits=2)),
              highlightOptions = highlightOptions(color = "white", weight = 0.1)) %>%
  addLegend(pal = pal2, values = ~reschool_programs$session_size, opacity = 1,title = "Sessions") %>%
  addCircles(data = agg_slots, 
             opacity = 1,weight=7, popup=paste("n of sessions =",agg_slots$session_size),
             color = ~pal2(agg_slots$session_size) ) %>%
  setView(lat=39.7,lng=-104.9,zoom=11.5) %>%
  addLegend(pal = pal, values = ~driving_index_sa$AI_overall, opacity = 1, title = "SA Access Index")

#Slots leaflet map
colourCount <-  length(unique(reschool_programs$session_size))
getPalette <- colorRampPalette(brewer.pal(9,"Reds"))

range <- range(driving_index_slots_sa$SA_AI_overall)

pal <- colorNumeric(palette = "Blues",domain = range)

#indicating one popup info
info <- paste("n",reschool_programs$session_size)

#SA-Slots creating leaflet map
Ai_driving_Slots <- leaflet(census_blocks) %>%
  addTiles() %>%
  addPolygons(color = "RdYlBu",weight =1,smoothFactor = 0.5,opacity = 0.5,fillOpacity = 0.75,
              fillColor = ~pal(driving_index_slots_sa$SA_AI_overall),
              popup=paste("AI =",round(driving_index_slots_sa$SA_AI_overall,digits=2)),
              highlightOptions = highlightOptions(color = "white", weight = 0.1)) %>%
  addLegend(pal = pal2, values = ~agg_slots$session_size, opacity = 1,title = "Slots") %>%
  addCircles(data = agg_slots, 
             opacity = 1,weight=7, popup=paste("n of sessions =",agg_slots$session_size),
             color = ~pal2(agg_slots$session_size) ) %>%
  setView(lat=39.7,lng=-104.9,zoom=11.5) %>%
  addLegend(pal = pal, values = ~driving_index_slots_sa$SA_AI_overall, opacity = 1, title = "SA Access Index")
```

