---
title: "Denver Open Data Processing - Programs and Resources"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This file processes the Denver Open Data contained in the OSR RShiny app. Namely, the data in the "other resources" tab, on parks, playgrounds, rec centers, libraries, and athletic fields, pools, licensed child care. If you need to update the data in a subsequent year, you should be able to just run this file. If any of the structure in the Denver Open Data (link format, variable names in files, etc.) changes, though, some updates may be required.

First, set up the workspace.

```{r}
# Uncomment the line below if you're using RStudio to run the file
# (don't use if you're running knitr!)
# (makes sure data files are saved in same location as this file)
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
```

```{r libs, message=FALSE, results="hide"}
library(tidyverse)
library(rgdal)  # for working with spatial data frames
library(rgeos)  # for working with spatial data frames
library(splitstackshape)  # for splitting strings and turning into binary columns
library(gsheet)  # only needed for the afterschool programs, to merge with annotated data

source('open_data_functions.R')  # our functions
```

### Rec Centers

First, we look at recreation centers.

```{r, results="hide"}
recCenters <- GetOpenData("recreation_centers")
```

```{r}
colnames(recCenters@data)
head(recCenters)
```

Delete unuseful columns (urls, links to pdfs and photos, address info, contact info, hours) (Note: 'FACILITIES' is an old version of 'MARKETED_F' which is marketed facilities.)

```{r}
recSmall <- recCenters[, c('REC_NAME', 'REC_TYPE', 'MARKETED_F', 'MARKETED_P', 'YEAR_BUILT', 'YEAR_REMOD', 'BLDG_SQFT', 'LABEL')]
colnames(recSmall@data) <- c('name', 'type', 'FACILITIES', 'PROGRAMS', 'year_built', 'year_last_remodeled', 'bldg_sqft', 'short_name')
head(recSmall)

# there are some weird entries in year_built and year_remodeled - fix these
levels(recSmall@data$year_built) <- gsub("Unknown", NA, levels(recSmall@data$year_built))
levels(recSmall@data$year_built) <- gsub("1900's", "1900", levels(recSmall@data$year_built))
levels(recSmall@data$year_built) <- gsub("1960's", "1960", levels(recSmall@data$year_built))

levels(recSmall@data$year_last_remodeled) <- gsub("Unknown", NA, levels(recSmall@data$year_last_remodeled))
levels(recSmall@data$year_last_remodeled) <- gsub("None", NA, levels(recSmall@data$year_last_remodeled))
levels(recSmall@data$year_last_remodeled) <- gsub("\\d{4}, ", "", levels(recSmall@data$year_last_remodeled))
```

Split up the facility categories (currently contains lists of facilities in a single column) into separate, binary columns. Only keep the meaningful columns.

```{r}
# Turn each facility type into a column
recFinal <- SplitCommas(recSmall, 'FACILITIES')
colnames(recFinal@data)

# Compute interesting facility distinctions: HAS_CARDIO, HAS_WEIGHTS, HAS_POOL
recFinal@data$has_cardio <- pmax(recFinal@data$FACILITIES_Aerobics, recFinal@data$FACILITIES_Cardio.Eqpmnt, recFinal@data$FACILITIES_Cardio.Eqpmt)

recFinal@data$has_weights <- recFinal@data$FACILITIES_Weight.Room

recFinal@data$has_gym <- pmax(recFinal@data$FACILITIES_Gym..Large., recFinal@data$FACILITIES_Gym..Reg..Size.,recFinal@data$FACILITIES_Gym..Small.)

# Delete unnecessary/redundant facilities columns
colnames(recFinal@data)
recFinal@data <- recFinal@data[, -grep("FACILITIES", colnames(recFinal@data))]
```

Similarly, split up the program categories into separate, binary columns.

```{r}
# Turn each facility type into a column
recFinal <- SplitCommas(recFinal, 'PROGRAMS')
colnames(recFinal@data)

# Combine aquatics programs into one column
recFinal@data$has_aquatics <- pmax(recFinal@data$PROGRAMS_Aquatics, recFinal@data$PROGRAMS_Aquatics..using.Wash.Park.Pool.)
recFinal@data[,"PROGRAMS_Aquatics"] <- NULL
recFinal@data[,"PROGRAMS_Aquatics..using.Wash.Park.Pool."] <- NULL

# Rename the rest of the binary columns
colnames(recFinal@data)[10:17] <- c("has_arts_culture", "has_day_camps", "has_educ_programs", "has_fitness_health_programs", "has_senior_programs", "has_social_enrich_clubs", "has_special_events", "has_sports")

# Cast binary columns as TRUE/FALSE
for (colnum in 7:18) {
  recFinal@data[,colnum] <- as.logical(recFinal@data[,colnum])
}
head(recFinal)

# delete unnecessary/redundant columns
colnames(recFinal@data)
```

### Athletic fields

Next, look at athletic fields.

```{r, results="hide"}
fields <- GetOpenData("athletic_fields")
```

```{r}
colnames(fields@data)
head(fields@data)
```

Subset to useful variables and rename columns for simplicity.

```{r}
fieldsFinal <- fields[, c('FEATURE', 'LOCATION', 'FIELD_TIER', 'CLASS_CATE')]
colnames(fieldsFinal@data) <- c('sport', 'location', 'tier', 'class')
head(fieldsFinal@data)
```

### Playgrounds

Next, look at playgrounds.

```{r, results="hide"}
playgrounds <- GetOpenData("playgrounds")
```

```{r}
colnames(playgrounds@data)
head(playgrounds@data)
```

Subset to useful variables and rename columns for simplicity.

```{r}
playgroundsSmall <- playgrounds[, c('LOCATION', 'YEAR_REHAB', 'CLASS_CATE')]
colnames(playgroundsSmall@data) <- c('location', 'year_rehab', 'class')
head(playgroundsSmall@data)

# There are weird values in year_rehab - fix them
levels(playgroundsSmall@data$year_rehab)
levels(playgroundsSmall@data$year_rehab) <- gsub("unknown", NA, levels(playgroundsSmall@data$year_rehab))
levels(playgroundsSmall@data$year_rehab) <- gsub("2015 - partial upgrade", "2015", levels(playgroundsSmall@data$year_rehab))

playgroundsFinal <- playgroundsSmall
```

### Libraries

Next we look at libraries.

```{r, results="hide"}
libraries <- GetOpenData("libraries")
```

A lot of this info is redundant, such as abbreviations, address/state/zip (which is contained in the shapefile metadata). Other info is unnecessary, such as the status (whether under construction, temporary construction, etc). We also will rename the columns to understandable names.

Look at the data, subset, retype columns, and then save result.

```{r}
head(libraries)

librariesFinal <- libraries[,c("LIBRARY_NA","DOOR_COUNT","CIRCULATIO","TOTAL_BLDG")]
names(librariesFinal) <- c("name","patron_count","circulation_volume","sqft")

# make sure the types are correct
librariesFinal[["name"]] <- as.character(librariesFinal[["name"]])
librariesFinal[["patron_count"]] <- as.numeric(as.character(librariesFinal[["patron_count"]]))
librariesFinal[["circulation_volume"]] <- as.numeric(as.character(librariesFinal[["circulation_volume"]]))
librariesFinal[["sqft"]] <- as.numeric(as.character(librariesFinal[["sqft"]]))

head(librariesFinal)
```

### Swimming Pools

Next we look at swimming pools.

```{r, results="hide"}
pools <- GetOpenData("swimming_pools")
```

Look at and subset the data.

```{r}
poolsFinal <- pools[,c("POOL_NAME","POOL_TYPE","PARK_LOCAT")]
names(poolsFinal) <- c("name","type","location")
head(poolsFinal)
```

### Licensed Childcare Facilities

Next we look at childcare facilities.

```{r, results="hide"}
care <- GetOpenData("licensed_child_care_facilities")
```

Look at and subset data, then save.

```{r}
summary(care)

# get rid of places whose license has expired (as of today)
expired_rows <- which( as.Date(care@data$END_DATE) < Sys.Date() )

summary(as.Date(care@data[-expired_rows,"END_DATE"]))

careFinal <- care[-expired_rows,c("BUS_PROF_N","LIC_TYPE")]
names(careFinal) <- c("name", "license_type")
levels(careFinal@data$license_type) <- c("center", "home")

head(careFinal)
```

### Parks

```{r, results="hide"}
parks <- GetOpenData("parks")
```

Look at and subset parks data.

```{r}
head(parks@data)
parksFinal <- parks[,c("FORMAL_NAM","PARK_CLASS","FACILITIES")]
names(parksFinal) <- c("name","class","facilities")
parksFinal[["name"]] <- as.character(parksFinal[["name"]])
```

Split facilities lists into separate columns. Only keep some of the details.

```{r}
parksFinal <- SplitCommas(parksFinal, "facilities")
colSums(parksFinal@data[,3:194])

# What columns do we have now?
colnames(parksFinal@data)

# Add variable about whether the park has some kind of natural area
natureBool <- grepl("Nature|nature|NATURE|Natural|natural|NATURAL", colnames(parksFinal@data))
colnames(parksFinal@data)[natureBool]  # look at what columns you're getting
parksFinal@data$has_nature <- apply(parksFinal@data[,natureBool], 1, max)

# Add variable about whether the park has a garden
gardenBool <- grepl("Garden|garden|GARDEN", colnames(parksFinal@data))
colnames(parksFinal@data)[gardenBool]  # look at what columns you're getting
parksFinal@data$has_garden <- apply(parksFinal@data[,gardenBool], 1, max)

# Add variable about whether the park has a trail.
trailBool <- grepl("Trail|trail|TRAIL", colnames(parksFinal@data))
colnames(parksFinal@data)[trailBool]  # look at what columns you're getting
parksFinal@data$has_trail <- apply(parksFinal@data[,trailBool], 1, max)
parksFinal@data$has_trail <- NULL  # actually ignore this variable because they all have trails (lol)

# bike paths
bikeBool <- grepl("Bike|bike|BIKE|cycle|Cycle|CYCLE", colnames(parksFinal@data))
colnames(parksFinal@data)[bikeBool]  # look at what columns you're getting
parksFinal@data$has_biking <- apply(parksFinal@data[,bikeBool], 1, max)

# delete the extra variables
colnames(parksFinal@data)
parksFinal@data <- parksFinal@data[, -grep("facilities", colnames(parksFinal@data))]

# cast 0/1 as true/false
for (colnum in 3:5){
  parksFinal@data[,colnum] <- as.logical(parksFinal@data[,colnum])
}

head(parksFinal@data)
```

### Saving to csv files:

```{r}
preprefix <- dirname(dirname(getwd()))
prefix <- file.path(preprefix, 'DATA', 'Open_Denver')

SavePointsAsCSV(recFinal, prefix, "rec_centers.csv")
SavePolygonsAsCSV(fieldsFinal, prefix, "fields.csv")
SavePolygonsAsCSV(playgroundsFinal, prefix, "playgrounds.csv")
SavePointsAsCSV(librariesFinal, prefix, "libraries.csv")
SavePointsAsCSV(poolsFinal, prefix, "pools.csv")
SavePointsAsCSV(careFinal, prefix, "licensed_child_care.csv")
SavePolygonsAsCSV(parksFinal, prefix, "parks.csv")
```

### Adding block groups to each dataset, from lat / longs

Not doing this anymore. Neighborhood name no longer in data tab for other resources.

```{r}
# Clear the environment
#rm(list=ls())

# # Source the necessary files
# mypath <- getwd()
# source( file.path(dirname(mypath), "nest_geo_areas", "coords_to_blockgroups.R") )
# bgDataPath <- file.path(dirname(dirname(mypath)), "shape_census")
# 
# filenameList <- c("rec_centers.csv", "fields.csv", "playgrounds.csv", 
#                   "libraries.csv", "pools.csv", 
#                   "licensed_child_care.csv", "parks.csv")
# 
# for (filename in filenameList) {
#   # load csv into workspace
#   file <- file.path(prefix, filename)
#   df <- read.csv(file)
#   
#   # add the block groups
#   dfnew <- coords_to_blockgroups(df, bgDataPath)
#   
#   # overwrite the csv 
#   write.csv(dfnew, file, row.names = FALSE, na = "")
# }
```

## Make the codebook

Build the codebook, i.e. get variable names for each csv saved above in clean_data. Save the results as a csv for easy referencing later.

```{r}
filenameList <- c("rec_centers.csv", "fields.csv", "playgrounds.csv", 
                  "libraries.csv", "pools.csv", 
                  "licensed_child_care.csv", "parks.csv")

maxVars <- 25
codebook <- data.frame(matrix(nrow=maxVars, ncol=0))

for (filename in filenameList) {
  # load csv into workspace
  file <- read.csv(file.path(prefix,filename) )
  
  vars <- rep(NA, maxVars)
  vars[1:length(names(file))] <- names(file)
  
  # save column names to dataframe
  codebook[[filename]] <- vars
}

write.csv(codebook, file=file.path(prefix,"codebook.csv"), row.names=FALSE)
```

## Get neighborhood level demographics (ACS data from Denver Open Data)

```{r}
spdf <- GetNbhdDemog()
outputdir <- file.path(prefix, "nbhd_dem_shapes")
writeOGR(spdf, outputdir, "nbhd_dem_shapes", driver="ESRI Shapefile")
```