---
title: "Create_census"
author: "Haowen Zheng"
date: "6/20/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include = FALSE}
library(tidyverse)
library(rgdal)
# shape_afterschool <- readOGR("/Users/luna/Desktop/DSSG/data/afterschool_programs", "afterschool_programs")
```

### 1. Neighborhood level demographics from Open Denver 

I used it as a partial template for the construction of some variables. 

```{r}
demographics <- read.csv(file = "american_community_survey_nbrhd_2011_2015.csv")

str(demographics)
demographics$MED_FAMILY_INCOME
head(demographics)
```

### 2. Census data on block group level, Denver County. From ACS 2016. Estimates are based on 5 year average (2012-2016)

We will end up with one big file besides the seperate files we intentionally left as seperate files. "Census" contains all demographic estimates. 

#### Population counts 
```{r}
# get data 
population <- read.csv("ACS_16_5YR_B01003_with_ann.csv", skip = 1)

colnames(population)[which(names(population_ME) == "Estimate..Total")] <- "TTL_population" #rename columns
colnames(population)[which(names(population_ME) == "Margin.of.Error..Total")] <- "Margin.of.Error..TP"

population <- select(population, -5) #get rid of Margins of Error
```

#### Sex
```{r}
#the file from AFF is age by sex
sexage <- read.csv("ACS_16_5YR_B01001_with_ann.csv", skip = 1)

#first look at sex
sex <- sexage %>% 
  select(c("Id","Estimate..Male.", "Estimate..Female."))

colnames(sex_ME)[2:3] <- c("Male", "Female")

#merge sex into census
census <- left_join(population, sex, by = "Id")
head(census)
```

#### Age
```{r}
#age groups
age <- sexage 
age$Estimate..Male. <- NULL
age$Estimate..Female. <- NULL

age <- age%>% 
  select(-contains("Margin")) 

#Combine the age groups 
 #1. combine male and females under the same age group
#Extracting the names of the columns that need to be modified
column_names = colnames(age)[5:27]
#Extracting the appropriate string to name the modified columns
modified_column_name = sapply(1:length(column_names), function(x) sub(".*Male....", "", column_names[x]))
#Adding the columns that are necessary and giving them appropriate names
added_columns = cbind(age[,5:27] + cbind(age[,28:50]))
added_columns = as.data.frame(added_columns)
colnames(added_columns) = modified_column_name
#Delete the columns not required and adding the modified columns to the main censusage dataset
age2 <- age[-c(2:3, 5:50)]
age2 <- cbind(age2,added_columns)

colnames(age2)[2:6] <- c("Total", "Age_Less_5", "Age_5_to_9", "Age_10_to_14", "Age_15_to_17") #rename columns

age3 <- age2 %>%  #add up columns to get new age groups with bigger age ranges 
  mutate(Age_Less_18 = rowSums(.[3:6])) %>%
  mutate(Age_0_to_9 = rowSums(.[3:4]), Age_10_to_19 = rowSums(.[5:7]), Age_20_to_29 = rowSums(.[8:11]), 
         Age_30_to_39 = rowSums(.[12:13]), Age_40_to_49 = rowSums(.[14:15]), Age_50_to_59 = rowSums(.[16:17]),
         Age_60_to_69 = rowSums(.[18:21]), Age_70_to_79 = rowSums(.[22:23]), Age_80_and_over = rowSums(.[24:25]),
         PCT_Less_18 = Age_Less_18/Total * 100) 
age_final <- age3
head(age_final)

census <- left_join(census, age_final, by = "Id") #merge age into census

head(census)
```

#### Race

We want both the counts and percentages of each race.

```{r}
#race
race_ME <- read.csv("ACS_16_5YR_B02001_with_ann.csv", skip = 1)

race <- race_ME%>% 
  select(-contains("Margin")) 

#Extracting the names of the columns that need to be modified
column_names <- colnames(race)[5:13]
#Extracting and manupulating the appropriate string to name the modified columns
modified_column_name1 <- sapply(1:length(column_names), function(x) gsub(".*Total....", "", column_names[x]))
modified_column_name <- sapply(1:length(modified_column_name1), function(x) paste0("PCT ", modified_column_name1[x]))

#Obtaining the percentages
additional_columns <- (race[,5:13]/race[,4])*100
#Renaming the column names
additional_columns <- as.data.frame(additional_columns)
colnames(additional_columns) <- modified_column_name

#Delete the columns not required and adding the modified columns to the main censusage dataset
race_final <- cbind(race, additional_columns) 
race_final <- race_final[-c(2:4)]
#rename columns

race_final$Two.or.more.races <- rowSums(race_final[8:10])
race_final[,8:10] <- NULL
race_final$PCT_Two.or.more.races <- rowSums(race_final[14:16])
race_final[,14:16] <- NULL

colnames(race_final)[2:7] <- c("White", "Black", "AmericanIndian", "Asian", "PacificIslanders", "OtherRaces")
colnames(race_final)[8:13] <- c("PCT_White", "PCT_African", "PCT_AmericanIndian", "PCT_Asian", "PCT_PacificIslander", "PCT_OtherRaces")

race_final <- race_final %>%
              select(Id:OtherRaces, Two.or.more.races, PCT_White:PCT_OtherRaces, PCT_Two.or.more.races)

head(race_final)

#get ethnicity "Hispanics" 
hispanic <- read.csv("ACS_16_5YR_B03003.csv", skip = 1)
hispanic <- hispanic %>% select(Id, Estimate..Total....Hispanic.or.Latino)
colnames(hispanic)[2] <- "Hispanic"

head(hispanic)

race.ethnicity <- left_join(race_final, hispanic, by = "Id")
census <- left_join(census, race.ethnicity, by = "Id")
census$PCT_Hispanic <- census$Hispanic/census$TTL_population * 100

head(census)  
```

#### Education levels

Categories include TTPOP_25PLUS_EDU, LESS_THAN_HS_DIPLOMA, HSGREAD_OR_EQUIV_EDU, SOMECOLLEGE_OR_AA_EDU, BACHELORS_OR_HIGHER_EDU

``` {r}
census_edu_ME <- read.csv("ACS_16_5YR_B15003.csv", skip = 1)
census_edu <- census_edu_ME %>% select(-contains("Margin")) 
colnames(census_edu)[which(names(census_edu) == "Estimate..Total.")] <- "TTPOP_25PLUS_EDU" #rename "estimate total" column to "total population aged 25 years and above"

census_edu$LESS_THAN_HS_DIPLOMA <- rowSums(census_edu[, 5:19]) #recategorize educational levels
census_edu$HSGREAD_OR_EQUIV_EDU <- rowSums(census_edu[,c(20, 21)])  
census_edu$SOMECOLLEGE_OR_AA_EDU <- rowSums(census_edu[, 22:24])
census_edu$BACHELORS_OR_HIGHER_EDU <- rowSums(census_edu[, 25:28])

census_edu.final <- census_edu %>% select(-c(2:3,5:28))

head(census_edu.final)
census <- left_join(census, census_edu.final, by = "Id") #merge 

head(census)
```

#### Income

Census blog group level, median household income (in the past 12 months, adjusted in 2016 inflation money), Source: U.S. Census Bureau, 2012-2016 American Community Survey 5-Year Estimates. [Source](https://factfinder.census.gov/faces/tableservices/jsf/pages/productview.xhtml?pid=ACS_16_5YR_B19013&prodType=table)

```{r}
median_hh_income <- read.csv("ACS_16_5YR_B19013.csv", skip = 1)
median_hh_income[, c(2,3,5)] <- NULL #get rid of the marigin of error
colnames(median_hh_income)[2] <- "Median_HH_income"

head(median_hh_income)
census <- left_join(census, median_hh_income, by = "Id")

head(census)
```

#### Nativity
```{r}
nativity_ME <- read_csv("ACS_16_5YR_B99051.csv", skip = 1)
nativity <- nativity_ME %>% 
            select(-contains("llocated")) %>% #I got rid of the missing data computation; and only kept the total estimates
            select(-c(2,3,4)) #get rid of the Margins of Error
nativity <- nativity[,1:2]  #remove the repetitive columns   
colnames(nativity)[2] <- "Native"

head(nativity)
census <- left_join(census, nativity, by = "Id")
head(census)
```

#### Housing tenure
```{r}
housing <- read.csv("ACS_16_5YR_B25044.csv", skip = 1)

housing <- housing %>% select(-contains("Margin"))

housing$PCT_Housing_rent <- housing$Estimate..Renter.occupied./housing$Estimate..Total. * 100
housing$PCT_Vehicle_none <- (housing$Estimate..Owner.occupied....No.vehicle.available +          
                            housing$Estimate..Renter.occupied....No.vehicle.available)/
                            housing$Estimate..Total. * 100
housing$PCT_Vehicle_own <- (housing$Estimate..Owner.occupied....1.vehicle.available + 
                            housing$Estimate..Owner.occupied....2.vehicles.available +
                            housing$Estimate..Owner.occupied....3.vehicles.available +
                            housing$Estimate..Owner.occupied....4.vehicles.available +
                            housing$Estimate..Owner.occupied....5.or.more.vehicles.available +
                            housing$Estimate..Renter.occupied....1.vehicle.available +
                            housing$Estimate..Renter.occupied....2.vehicles.available +
                            housing$Estimate..Renter.occupied....3.vehicles.available +
                            housing$Estimate..Renter.occupied....4.vehicles.available +
                            housing$Estimate..Renter.occupied....5.or.more.vehicles.available)/
                            housing$Estimate..Total. * 100
housing <- housing %>%
           select(Id, PCT_Housing_rent, PCT_Vehicle_none, PCT_Vehicle_own)

head(housing)

census <- left_join(census, housing, by = "Id")

head(census)
```

#### Language
``` {r}
########################################################################################################
#Sree's code to tidy the data set.
languagedata = read.csv(file = "ACS_16_5YR_B16004.csv", skip = 1, na.strings = "")

#Selecting only the 'estimate' columns and deleting all the 'margin of erro'r columns
languagedata <- languagedata[,c(1,2,3,4,seq(50,ncol(languagedata),2))]

#Selecting only the 'estimate' columns and deleting all the 'margin of erro'r columns
languagedata_1 <- languagedata[,-c(1:7,12, 17, 22, 27:29, 34, 39, 44)]
languagedata_2 <- languagedata[, c(1:7,12, 17, 22, 27:29, 34, 39, 44)]

modified_columns = cbind(languagedata_1[c(seq(1,31,2))] + languagedata_1[c(seq(2,32,2))])

languagedata_final = cbind(languagedata_2, modified_columns)

#Making the column names much better
library(magrittr)

x = colnames(languagedata_final)
x %<>% gsub("Estimate..", "", .) %>% gsub(".years....", "", .) %>% gsub("over....", "", .) %>% gsub("Speak", "", .)

#Assigning the column names
colnames(languagedata_final) = x
########################################################################################################

language <- languagedata_final
language$TTPOP_18PLUS_LNG <- language$X18.to.64.years. + language$X65.over. #combine across the two age groups
language$LNG_English <- language$X18.to.64.only.English + language$X65..only.English
language$LNG_Spanish <- language$X18.to.64.Spanish. + language$X65..Spanish.
language$LNG_Spanish_English_bad <- language$X18.to.64.Spanish.....English..not.well. + language$X65..Spanish.....English..not.well.
language$LNG_IndoEuropean <- language$X18.to.64.other.Indo.European.languages. + language$X65..other.Indo.European.languages.
language$LNG_IndoEuropean_English_bad <- language$X18.to.64.other.Indo.European.languages.....English..not.well. + 
                                         language$X65..other.Indo.European.languages.....English..not.well.
language$LNG_AsianPacific <- language$X18.to.64.Asian.and.Pacific.Island.languages. + language$X65..Asian.and.Pacific.Island.languages.
language$LNG_AsianPacific_English_bad <- language$X18.to.64.Asian.and.Pacific.Island.languages.....English..not.well. +
                                         language$X65..Asian.and.Pacific.Island.languages.....English..not.well.
language$LNG_Others <- language$X18.to.64.other.languages. + language$X65..other.languages.
language$LNG_Others_English_bad <- language$X18.to.64.other.languages.....English..not.well. + 
                                   language$X65..other.languages.....English..not.well.
language_final <- language %>% select(Id, TTPOP_18PLUS_LNG:LNG_Others_English_bad)
    
head(language_final)

census <- left_join(census, language_final, by = "Id") #merge into "census"
head(census)
```

#### Poverty

POVERTY STATUS IN THE PAST 12 MONTHS OF FAMILIES BY FAMILY TYPE BY PRESENCE OF RELATED CHILDREN UNDER 18 YEARS BY AGE OF RELATED CHILDREN
 
```{r}
#read in data
homecomposition <- read.csv("ACS_16_5YR_B17010.csv", skip = 1)
#first get rid of the families that do not have children under 18 years old.
poverty <- homecomposition
poverty$Poverty_TTL_HH <- poverty$Estimate..Total. -
  poverty$Estimate..Income.in.the.past.12.months.below.poverty.level....Married.couple.family....No.related.children.of.the.householder.under.18.years -
  poverty$Estimate..Income.in.the.past.12.months.below.poverty.level....Other.family....Male.householder..no.wife.present....No.related.children.of.the.householder.under.18.years -
  poverty$Estimate..Income.in.the.past.12.months.below.poverty.level....Other.family....Female.householder..no.husband.present....No.related.children.of.the.householder.under.18.years -
  poverty$Estimate..Income.in.the.past.12.months.at.or.above.poverty.level....Married.couple.family....No.related.children.of.the.householder.under.18.years -
  poverty$Estimate..Income.in.the.past.12.months.at.or.above.poverty.level....Other.family....Male.householder..no.wife.present....No.related.children.of.the.householder.under.18.years -
  poverty$Estimate..Income.in.the.past.12.months.at.or.above.poverty.level....Other.family....Female.householder..no.husband.present....No.related.children.of.the.householder.under.18.years

poverty$Poverty_HH <-   homecomposition$Estimate..Income.in.the.past.12.months.below.poverty.level. -
           homecomposition$Estimate..Income.in.the.past.12.months.below.poverty.level....Married.couple.family....No.related.children.of.the.householder.under.18.years -
           homecomposition$Estimate..Income.in.the.past.12.months.below.poverty.level....Other.family....Male.householder..no.wife.present....No.related.children.of.the.householder.under.18.years -
           homecomposition$Estimate..Income.in.the.past.12.months.below.poverty.level....Other.family....Female.householder..no.husband.present....No.related.children.of.the.householder.under.18.years

#get the percentage of families that have children under 18 years old and are under poverty line 
poverty$PCT_HH_Poverty <- poverty$Poverty_HH/poverty$Poverty_TTL_HH * 100

poverty <- poverty %>% select(Id, Poverty_TTL_HH, Poverty_HH, PCT_HH_Poverty)
colnames(poverty)[2] <- "TTL_HH_Children_under_18" 
colnames(poverty)[4] <- "Poverty_PCT_HH"

head(poverty)
census <- left_join(census, poverty, by = "Id")
head(census)
```

#### Home composition
```{r}
homecomposition <- homecomposition %>%
                   select(-contains("Margin"))

#married-couple hhs, single-dad hhs, single-mom hhs with children under 18 years old
homecomposition$MarriedCouple_HH <-
  homecomposition$Estimate..Income.in.the.past.12.months.below.poverty.level....Married.couple.family....With.related.children.of.the.householder.under.18.years. +
  homecomposition$Estimate..Income.in.the.past.12.months.at.or.above.poverty.level....Married.couple.family....With.related.children.of.the.householder.under.18.years.

homecomposition$SingleDad_HH <- 
  homecomposition$Estimate..Income.in.the.past.12.months.below.poverty.level....Other.family....Male.householder..no.wife.present....With.related.children.of.the.householder.under.18.years. +
  homecomposition$Estimate..Income.in.the.past.12.months.at.or.above.poverty.level....Other.family....Male.householder..no.wife.present....With.related.children.of.the.householder.under.18.years.

homecomposition$SingleMom_HH <- 
  homecomposition$Estimate..Income.in.the.past.12.months.below.poverty.level....Other.family....Female.householder..no.husband.present....With.related.children.of.the.householder.under.18.years. +
  homecomposition$Estimate..Income.in.the.past.12.months.at.or.above.poverty.level....Other.family....Female.householder..no.husband.present....With.related.children.of.the.householder.under.18.years.
  
#PCT
homecomposition$TTL_HH <- poverty$TTL_HH_Children_under_18
homecomposition$PCT_MarriedCouple_HH <- homecomposition$MarriedCouple_HH/homecomposition$TTL_HH * 100
homecomposition$PCT_SingleDad_HH <- homecomposition$SingleDad_HH/homecomposition$TTL_HH * 100
homecomposition$PCT_SingleMom_HH <- homecomposition$SingleMom_HH/homecomposition$TTL_HH * 100

homecomposition <- homecomposition %>% 
  select(Id, MarriedCouple_HH, SingleDad_HH, SingleMom_HH, PCT_MarriedCouple_HH, PCT_SingleDad_HH, PCT_SingleMom_HH)

head(homecomposition)
census <- left_join(census, homecomposition, by = "Id")
head(census)

write.csv(census, file = "Denver Demographics on block group level 2016.csv")
census <- read.csv("/Users/luna/Desktop/DSSG/ors/data/clean data/Denver Demographics on block group level 2016.csv")
```

### 3. Separate files

#### Commuting

Means of transportation to work by travel time to work for workers 16 years and over who did not work at home.
Leave the file for future manipulation. 

```{r}
commute <- read_csv("/Users/luna/dssg2018/osr_dssg2018/data/raw data/Commute data/ACS_16_5YR_B08134.csv")
# commute <- commute %>% select(-contains("Margin"))

colnames(commute) = dbSafeNames(colnames(commute))

head(commute)
```

### 4. Shapefile 

I got the shapefile from Tiger/Line and merged it into the census dataset. 

```{r}
shape_conorado <- readOGR("/Users/luna/Desktop/tl_2016_08_bg", "tl_2016_08_bg")

census$Id2 <- as.character(census$Id2)
census$Id2 <- paste0("0", census$Id2)

#merge 
library(tigris)
shape_census <- geo_join(shape_conorado, census, "GEOID", "Id2", how = "inner")
shape_census@data$Id2 <- gsub("0(\\d{11})","\\1",shape_census@data$Id2)

#write the shapefile
writeOGR(obj=shape_census, dsn="shape_census", layer="shape_census", driver="ESRI Shapefile")

#try if it works
# shape_census <- readOGR(dsn="/path/shape_census", layer="shape_census")
```

### 5. Reformatting

```{r}
# change column names
# make names db safe: no '.' or other illegal characters,
# all lower case and unique
dbSafeNames = function(names) {
  names = gsub('[^a-z0-9]+','_',tolower(names))
  names = make.names(names, unique=TRUE, allow_=TRUE)
  names = gsub('.','_',names, fixed=TRUE)
  names
}

colnames(census) = dbSafeNames(colnames(census))

#write a new csv
write.csv(census, file ="Denver Demographics on block group level 2016.csv")

#get the codebook
filenameList <- c("Denver Demographics on block group level 2016.csv")

maxVars <- 70
codebook <- data.frame(matrix(nrow=maxVars, ncol=0))

for (filename in filenameList) {
  # load csv into workspace
  file <- read.csv(file.path("/Users/luna/dssg2018/osr_dssg2018/data/clean data",filename) )
  
  vars <- rep(NA, maxVars)
  vars[1:length(names(file))] <- names(file)
  
  # save column names to dataframe
  codebook[[filename]] <- vars
}

write.csv(codebook, file=file.path("/Users/luna/dssg2018/osr_dssg2018/data/clean data","codebook_resources.csv"), row.names=FALSE)
```

### 6. Upload

```{r}
# install.packages("RPostgreSQL")

require("RPostgreSQL")

# loads the PostgreSQL driver
drv <- dbDriver("PostgreSQL")

# creates a connection to the postgres database
# note that "con" will be used later in each connection to the database
source('/Users/luna/Desktop/DSSG/ors/other/keyforRDS')

con <- dbConnect(drv, dbname = "dssg2018uw",
                 host = "localhost", port =9000,
                 user = user, password = password)

#here you will do what you need to do in the database.. like
dbWriteTable(con, c("clean", "acs_demographics"), value = census, row.names = FALSE)
dbWriteTable(con, c("raw", "commute"), value = commute, row.names = FALSE)
## dbWriteTable(con, c("raw", "acs_commute"), value = commute, row.names = FALSE) #unsuccessful
## did not upload the shapefile, waiting for the new database 

#or: 
#sql_df = dbGetQuery(con, "select * from _schema.table_")

#when you're done, close the connection and unload the driver 
dbDisconnect(con) 
dbUnloadDriver(drv)
```

