---
title: "Data_access_cleaning"
output: html_document
---

```{r}
#installing the relavent packages
library(tidyverse)

#install relavent packages to access aws
install.packages("aws.s3", repos = c("cloudyr" = "http://cloudyr.github.io/drat"))
library(aws.s3)

source('C:/Users/Sreekanth/Desktop/DSSG Project/key.txt')

Sys.setenv("AWS_ACCESS_KEY_ID" = access_key,
           "AWS_SECRET_ACCESS_KEY" = secret_key,
           "AWS_DEFAULT_REGION" = "us-west-2")

get_bucket('dssg2018')

data <- s3read_using(FUN = read.csv, object = "s3://dssg2018/rawdata/DPS_Choice_1314-1718.csv")



```




****************************************Looking at the Attendence dataset********************************************

```{r}

#Skipping the first row inorder to g et the appropriate headers
DPS_Attendance <- s3read_using(FUN = read.csv, object = "s3://dssg2018/rawdata/DPS_Attendance_1112-1718.csv", skip = 1, na.strings = "") 

#Assigning appropriate column names
colnames(DPS_Attendance) = c("StudentNumber", "2011-2012", "2012-2013", "2013-2014", "2014-2015", "2015-2016", "2016-2017", "2017-2018")
colnames(DPS_Attendance) <- make.names(colnames(DPS_Attendance))


DPS_Attendance = as.data.frame(DPS_Attendance)


head(DPS_Attendance)
```

Dimensions of the dataset
```{r}
dim(DPS_Attendance)
```


Checking if the student attendance data is distinct on the studentnumber level.
```{r}
nrow(DPS_Attendance)
length(unique(DPS_Attendance$StudentNumber))
length(which(is.na(DPS_Attendance$StudentNumber)))

```

The data is unique at the stuent number level and also no missing student numbers. 
```{r}
summary(DPS_Attendance)
```

We need discuss if the number of missing records in each of the years is a cause of concern. 

```{r}
nrow(DPS_Attendance[DPS_Attendance$na_count > 6,])
```
There are 90 student records that have attendance records missing or all the years. We might not necessarily need to have them in the dataset but we can look at relavent information for these students in other datasets and then take a call. 

Finding the minimum and maximum of attendence in each of the years and noting any discrepencies
```{r}

#Ignoring the studentnumbers
x = DPS_Attendance[,-1]

#Storing the minimum and maximum of each column as a dataframe
min_values = apply(x, 2, min, na.rm = TRUE)
min_values = as.data.frame(min_values)

max_values = apply(x, 2, max, na.rm = TRUE)
max_values = as.data.frame(max_values)

na_count <-apply(x, 2, function(y) sum(length(which(is.na(y)))))
min_max_info = cbind(min_values, max_values)

min_max_info
```
As we can see academic years 2014-2015 and 2016-2017 have incorrect minimum values which needs to be addressed.

What are the finaldata types of each of the columns in the dataset?
```{r}

sapply(DPS_Attendance, typeof)

```

*******************************Looking at the rawdata/DPS_CMAS_1415-1617.csv dataset*****************************************

```{r}
#Skipping the first row inorder to g et the appropriate headers
DPS_CMAS_data <- s3read_using(FUN = read.csv, object = "s3://dssg2018/rawdata/DPS_CMAS_1415-1617.csv", skip = 2, na.strings = "")
#Assigning appropriate column names

#Converting it as a dataframe
DPS_CMAS_data = as.data.frame(DPS_CMAS_data)

#Ensuring that there are readable and understandable column names
colnames(DPS_CMAS_data) = c("StudentNumber", "ELAScaleScore_2014-15", "ELAProficiency_2014-15", "ELAScaleScore_2015-16",
                            "ELAProficiency_2015-16", "ELAScaleScore_2016-17", "ELAProficiency_2016-17",  "MathScaleScore_2014-15",
                            "MathProficiency_2014-15", "MathScaleScore_2015-16", "MathProficiency_2015-16",
                            "MathScaleScore_2016-17","MathProficiency_2016-17")
head(DPS_CMAS_data)
```

Looking at the dimensions
```{r}
dim(DPS_CMAS_data)
```

Checking if the student attendance data is distinct on the studentnumber level.
```{r}
nrow(DPS_CMAS_data)
length(unique(DPS_CMAS_data$StudentNumber))
length(which(is.na(DPS_CMAS_data$StudentNumber)))

```
There seems to be student numbers which are repeated. Lets look at those students.
```{r}
DPS_CMAS_data[duplicated(DPS_CMAS_data$StudentNumber),]
```
Extracting the rows with these student numbers.
```{r}
DPS_CMAS_data[DPS_CMAS_data$StudentNumber %in% c(255727, 325054, 344328, 377188,451556),  ]
```

Need to discuss what has to be done with these multiple student numbers.
```{r}
summary(DPS_CMAS_data)
```

The above shows the number of NA's and need to discuss if we the number of NA's makes sense. The min and max is 650 and 850 for all the columns. 

Let us look at the unique values for proficiency comments.
```{r}
unique(DPS_CMAS_data$`MathProficiency_2016-17`)
```


*******************************Looking at the rawdata/DPS_CMAS_1415-1617.csv dataset*****************************************

```{r}
#Skipping the first row inorder to g et the appropriate headers
DPS_Enrollment <- s3read_using(FUN = read.csv, object = "s3://dssg2018/rawdata/DPS_Enrollment_1112-1718.csv", na.strings = "")
#Assigning appropriate column names

#Converting it as a dataframe
DPS_CMAS_data = as.data.frame(DPS_CMAS_data)
head(DPS_Enrollment)
```

```{r}
nrow(DPS_Enrollment)
length(unique(DPS_Enrollment$StudentNumber))
length(which(is.na(DPS_Enrollment$StudentNumber)))
```
```{r}
dim(DPS_Enrollment)
```
```{r}
unique(DPS_Enrollment$SchoolYear)
unique(DPS_Enrollment$LRE)
unique(DPS_Enrollment$SwD)
unique(DPS_Enrollment$Grade)
```
```{r}
DPS_Enrollment[duplicated(DPS_Enrollment$StudentNumber),]
```

```{r}
DPS_Enrollment[DPS_Enrollment$StudentNumber %in% c(304932, 323085),  ]
```

