---
title: "DPS"
author: "Joe"
date: "6/28/2018"
output: html_document
---

#Reading in the data#

First we install the necessary packages

```{r eval=FALSE}
install.packages("aws.s3", repos = c("cloudyr" = "http://cloudyr.github.io/drat"))
install.packages('tidyverse')
install.packages('gridExtra')
```

and load these packages.

```{r message=FALSE}
library(aws.s3)
library(tidyverse)
library(gridExtra)
```

Now we set up to retrieve the data from AWS. 

```{r message=FALSE}
source('/Users/kelliemacphee/Desktop/dssg2018/cred.txt')

Sys.setenv("AWS_ACCESS_KEY_ID" = access_key,
           "AWS_SECRET_ACCESS_KEY" = secret_key,
           "AWS_DEFAULT_REGION" = "us-west-2")
```

##Discipline##

We will read in data as a tibble using the read_csv (NOT read.csv) function. 

```{r message=FALSE, warning=FALSE}
raw_df <- s3read_using(FUN = read_csv, object = "s3://dssg2018/rawdata/DPS_Discipline_1112-1617.csv")

summary(raw_df)
```

As we can see, the data is currently wide. We would like to go from the 19 columns to just 4: "StudentNumber" (an int), "DisciplineType" (a factor), "SpringYear" (a year), and "NumberOffenses" (an int). In other words, we want the data to be "tidy".

```{r}
#in school suspensions
iSS_inds <- 1:7
iSS <- raw_df[,iSS_inds]
colnames(iSS) <- iSS[1,]
iSS <- iSS[-1,]

iSS <- iSS %>% gather('2011-2012','2012-2013','2013-2014','2014-2015','2015-2016','2016-2017',key = 'SpringYear' ,value='NumberOffenses')

# for making the table long
#iSS = add_column(iSS, DisciplineType = 'In School Suspension')

# for making ISS its own column
colnames(iSS)[3] <- "NumberISS"

#out of school suspensions
oSS_inds=c(1,8:13)
oSS = raw_df[oSS_inds]
colnames(oSS) = oSS[1,]
oSS = oSS[-1,]
oSS = oSS %>% gather('2011-2012','2012-2013','2013-2014','2014-2015','2015-2016','2016-2017',key = 'SpringYear' ,value='NumberOffenses')

# for making the table long
# oSS = add_column(oSS, DisciplineType = 'Out of School Suspension')

# for making ISS its own column
colnames(oSS)[3] <- "NumberOSS"

#expulsions
expulsion_inds=c(1,14:19)
expulsion = raw_df[expulsion_inds]
colnames(expulsion) = expulsion[1,]
expulsion = expulsion[-1,]
expulsion = expulsion %>% gather('2011-2012','2012-2013','2013-2014','2014-2015','2015-2016','2016-2017',key = 'SpringYear' ,value='NumberOffenses')

# for making the table long
# expulsion = add_column(expulsion, DisciplineType = 'Expulsion')

# for making expulsions its own column
colnames(expulsion)[3] <- "NumberExpulsions"

discipline = bind_rows(iSS, oSS, expulsion)
glimpse(discipline)
```

Finally, we do simple touch-ups like type conversions and dealing with NA.

```{r}
discipline$SpringYear <- sub("\\d{4}-","",discipline$SpringYear)
discipline$SpringYear <- as.integer(discipline$SpringYear)

# for making the data long
# discipline$DisciplineType <- as.factor(discipline$DisciplineType)
#discipline <- discipline %>% replace_na(list(NumberOffenses=0))
#discipline$NumberOffenses <- as.integer(discipline$NumberOffenses)

# for separate columns for ISS, OSS, expulsions
discipline <- discipline %>% replace_na(list(NumberISS=0))
discipline <- discipline %>% replace_na(list(NumberOSS=0))
discipline <- discipline %>% replace_na(list(NumberExpulsions=0))

discipline$StudentNumber <- as.integer(discipline$StudentNumber)
discipline <- discipline %>% arrange(StudentNumber)
#discipline <- discipline[,c(1,2,4,3)]

discipline <- as.data.frame(discipline)
head(discipline)

write.csv(discipline, file="discipline.csv", row.names=FALSE)
```

##ELA Test##

```{r}
ela <- s3read_using(FUN = read.csv, object = "s3://dssg2018/rawdata/DPS_EOCELA_1112-1617.csv")
```

Change SchoolYear to be the spring year only.
```{r}
levels(ela$SchoolYear) <- gsub("\\d{4}-", "", levels(ela$SchoolYear))
colnames(ela)[1] <- "SpringYear"
```

CourseFinal entries look unnecessarily complicated. Do years agree with years in the other column? 

```{r}
for (course in levels(ela$CourseFinal)){
  temp <- ela[ela$CourseFinal==course,]
  years <- unique(temp$SpringYear)
  print(paste(course, "happened in", years))
}
```

Answer: yes. Let's simplify these CourseFinal names, then, to just the course name (and Part I, II, or III, if necessary). Note that in cases when the year isn't specified, that means the exam occurred in 2012.

```{r}
levels(ela$CourseFinal) <- gsub(" Course Final \\d{4}", "", levels(ela$CourseFinal))
levels(ela$CourseFinal) <- gsub(" Course Final", "", levels(ela$CourseFinal))

# delete any trailing white spaces
levels(ela$CourseFinal) <- gsub(" $", "", levels(ela$CourseFinal))

levels(ela$CourseFinal)
```

Finally, look for missing or absurd values.

```{r}
sapply(ela, function(x) sum( is.na(x) | x=="" | x==" " ))
```

Make some plots, too, to see if distributions make sense or if there are any strange outliers.

```{r, echo=FALSE}
plot(ela$SpringYear, main="Years")
plot(ela$CourseFinal, main="ELA Subjects")
plot(ela$StudentNumber, main="Student Numbers")
hist(ela$PointsEarned, main="Points Earned")
hist(ela$PercentageEarned, main="Percentage Earned")
plot(ela$Proficiency, main="Proficiency")
```

```{r}
ela <- ela[,c(3,1,2,4,5,6)]
ela %>% arrange(StudentNumber)
head(ela)
```

##Math##

```{r}
math <- s3read_using(FUN = read.csv, object = "s3://dssg2018/rawdata/DPS_EOCMath_1112-1617.csv")
```

Overview of entries in each column:

```{r}
summary(math)
```

Change SchoolYear to be the spring year only.
```{r}
levels(math$SchoolYear) <- gsub("\\d{4}-", "", levels(math$SchoolYear))
colnames(math)[1] <- "SpringYear"
summary(math)
```

CourseFinal entries look unnecessarily complicated. Do years agree with years in the other column? 

```{r}
for (course in levels(math$CourseFinal)){
  temp <- math[math$CourseFinal==course,]
  years <- unique(temp$SpringYear)
  print(paste(course, "happened in", years))
}
```

Answer: yes. Let's simplify these CourseFinal names, then, to just the course name (and Part 1 or 2, if necessary). Note that in cases when the year isn't specified, that means the exam occurred in 2012.

```{r}
levels(math$CourseFinal) <- gsub(" Course Final \\d{4}", "", levels(math$CourseFinal))
levels(math$CourseFinal) <- gsub(" Course Final \\d{2}-\\d{2}", "", levels(math$CourseFinal))
levels(math$CourseFinal) <- gsub(" Course Final", "", levels(math$CourseFinal))

# delete any trailing white spaces
levels(math$CourseFinal) <- gsub(" $", "", levels(math$CourseFinal))

# make "Probabilty & Statistics" be "Probability and Statistics" for consistency 
levels(math$CourseFinal) <- gsub("\\&", "and", levels(math$CourseFinal))

levels(math$CourseFinal)
```

Finally, look for missing or absurd values.

```{r}
sapply(data, function(x) sum( is.na(x) | x=="" | x==" " ))
summary(data)
```

Make some plots, too, to see if distributions make sense or if there are any strange outliers.

```{r, echo=FALSE}
plot(math$SpringYear, main="Years")
plot(math$CourseFinal, main="Math Subjects")
plot(math$StudentNumber, main="Student Numbers")
hist(math$PointsEarned, main="Points Earned")
hist(math$PercentageEarned, main="Percentage Earned")
plot(math$Proficiency, main="Proficiency")
```

```{r}
math <- math[,c(3,1,2,4,5,6)]
math %>% arrange(StudentNumber)
head(math)
```

## Graduation  ##

```{r}
graduation <- s3read_using(FUN = read.csv, object = "s3://dssg2018/rawdata/DPS_Grad_1112-1617.csv", skip = 1)
transportation <- s3read_using(FUN = read.csv, object = "s3://dssg2018/rawdata/DPS_Transport.csv")
STARreading <- s3read_using(FUN = read.csv, object = "s3://dssg2018/rawdata/DPS_STAR_1112-1617.csv", skip = 1)
```

```{r}
summary(graduation)
```

We make graduation into a simple Graduation Year lookup by Student Number.

```{r results = 'hide'}
# change column names
colnames(graduation) <- sub("X20\\d\\d.", "", colnames(graduation))
graduationYear = rep(NA,length(graduation))
for (i in 2:7) {
  graduationYear[which(graduation[i]==1)] <- colnames(graduation)[i]
}
graduation$GraduationYear <- graduationYear
graduation <- graduation[c("StudentNumber", "GraduationYear")]
graduation %>% arrange(StudentNumber)
head(graduation, 6)
```

## Transportation ##

```{r results='hide'}
summary(transportation)
length(unique(transportation$StudentNumber))
 #StudentNumber is unique 
unique(transportation$Transportation)
 #Only one level under column Transportation: Yes 
unique(transportation$X)
 #This X column is totally usefuless - let's remove it
transportation <- transportation[,c("StudentNumber","Transportation")]
colnames(transportation)[2] <- "HasTransportation"
```

# STARreading

```{r results='hide'}
head(STARreading,6)
```

```{r}
colnames(STARreading) <- sub("X20\\d\\d.", "", colnames(STARreading))
STARreading = STARreading %>% gather('2012','2013','2014','2015','2016','2017',key = 'SpringYear' ,value='ReadingLevel')
STARreading <- STARreading %>% arrange(StudentNumber)
unique(STARreading$ReadingLevel)
```

We have some blank lines we should get rid of. 

```{r}
STARreading$ReadingLevel <- sub("^$",NA,STARreading$ReadingLevel)
STARreading <- STARreading %>% na.omit()
head(STARreading,6)
```

##Enrollment##
```{r}
enrollment <- s3read_using(FUN = read.csv, object = "s3://dssg2018/rawdata/DPS_Enrollment_1112-1718.csv")

head(enrollment)
summary(enrollment)
```

We want locations rather than just school names, so we use the Google maps API. This takes 5-10 minutes. We also want to ultimately separate out student from school information in the RDS, so we will have school information as one dataframe and student enrollment information in another. 

```{r eval=FALSE}
library(ggmap)
school_locations = data.frame(unique(enrollment$SchoolName))
colnames(school_locations) <- c("SchoolName")
school_locations$SchoolName = paste(school_locations$SchoolName, "denver", "colorado", sep=", ")

for(i in 1:nrow(school_locations)) {
  result <- geocode(school_locations$SchoolName[i], output="latlona", source="google")
  school_locations$lon[i] <- as.numeric(result[1])
  school_locations$lat[i] <- as.numeric(result[2])
  Sys.sleep(1)
  print(i)
}
unique_indices <- order(enrollment$SchoolName)[!duplicated(sort(enrollment$SchoolName))]
school_locations$SchoolName <- enrollment$SchoolName[unique_indices]
school_locations$SchoolNum <- enrollment$SchoolNum[unique_indices]
write.csv(school_locations, file="school_locations.csv", row.names=FALSE)
```

This is just so we don't have to rerun the above code. 

### School Locations ###

```{r}
school_locations <- read.csv(file="school_locations.csv")
unique(school_locations$lon)
```

Looks like we have some NA values.

```{r}
school_locations$SchoolName[which(is.na(school_locations$lon))]
```

We'll just manually fix that by typing into Google maps. STRIVE-Federal is at (39.67819,-105.0209) and DS Innovation & Sustainable Design is at (39.7377,-104.9793).

```{r}
school_locations[which(is.na(school_locations$lon)),]$lat <- c(39.67819,39.7377)
school_locations[which(is.na(school_locations$lon)),]$lon <- c(-105.0209,-104.9793)
```

Let's look at these locations to make sure nothing insane is going on. 

```{r}
library("leaflet")
library("leaflet.extras")
leaflet(school_locations) %>% 
  addProviderTiles("CartoDB.Positron") %>%
  addCircleMarkers(
    stroke = FALSE, label=school_locations$SchoolName,
    radius = 4, opacity = 1.0
  )
```

Scrolling over the sites outside of Denver and checking their names against the GreatSchools website, it looks like google mis-geocoded "DELTA HS" (the extremely southwestern point). "Satellite Program" and "Ridge View Academy Charter School" are also outside of Denver. Delta HS is alright, but Satellite Program and Ridge View Academy Charter School are outside of Denver limits so we will drop those. 

```{r}
school_locations[school_locations$SchoolName=="DELTA HS",]$lat <- 39.6866
school_locations[school_locations$SchoolName=="DELTA HS",]$lon <- -104.9565
school_locations <- school_locations %>% filter(!SchoolName %in% c("Satellite Program", "Ridge View Academy Charter School"))
enrollment$SchoolName <- as.character(enrollment$SchoolName)
enrollment <- enrollment %>% filter(!SchoolName %in% c("Satellite Program", "Ridge View Academy Charter School"))
```

Now we check the counts of students in schools. 

```{r}
head(sort(table(enrollment$SchoolName)))
```

We'll remove "Sandra Todd-Williams Academy" since there are only two students on record for them. GED High School sounds like students who got their GED (since it doesn't show up when googled). Looking at the data,

```{r}
enrollment %>% filter(SchoolName=="GED High School")
```

It looks like it was something they just did in 2011. So we'll exclude this also.

```{r}
enrollment <- enrollment %>% filter(!SchoolName %in% c("GED High School", "Sandra Todd-Williams Academy"))
school_locations <- school_locations %>% filter(!SchoolName %in% c("GED High School", "Sandra Todd-Williams Academy"))
```

Now we do some sanity checks via plots. 
```{r}
hist(enrollment$Grade,col = "black",main = "Grade Level")
barplot(summary(as.factor(enrollment$Gender)),main="Student Gender")
barplot(summary(as.factor(enrollment$PrimaryDisability!="No Disability")),main="Student Disability")
barplot(summary(as.factor(enrollment$RaceEthnicity)),main="Student Race/Ethnicity",las=2,cex.names=.6)
barplot(summary(as.factor(enrollment[enrollment$PrimaryDisability!="No Disability",]$PrimaryDisability)),main="Student Disability (Ignore `No Disability`)",las=2,cex.names=.4)
barplot(summary(as.factor(enrollment$LRE)),main="Student Least Restrictive Environment Setting")
```

Finally, we'll remove excess information. First, we see that "SwD" (student with disability) has the same info as "PrimaryDisability":

```{r}
length(enrollment$SwD==0) == length(enrollment$PrimaryDisability=="No Disability")
all((enrollment$SwD==0) == (enrollment$PrimaryDisability=="No Disability"))
```

After rearranging columns and eliminating the unnecessary ones:

```{r}
enrollment <- enrollment[,c(4,1,3,2,5,6,7,8,9,10,11)]
# School name is in the school information dataframe, so we can remove it
enrollment$SchoolName <- NULL
# Whether student has disability is contained in PrimaryDisability as "No Disability" for 0 and 1 otherwise.
enrollment$SwD <- NULL
enrollment <- enrollment %>% arrange(StudentNumber)
colnames(enrollment)[2] <- "SpringYear"
head(enrollment,6)
```

```{r}
school_locations <- school_locations[,c(1,4,3,2)]
head(school_locations,6)
```

##Attendance##

```{r}

#Skipping the first row inorder to g et the appropriate headers
attendance <- s3read_using(FUN = read.csv, object = "s3://dssg2018/rawdata/DPS_Attendance_1112-1718.csv", skip = 1, na.strings = " ") 

#Assigning appropriate column names
colnames(attendance) <- gsub("X\\d{4}.","",colnames(attendance))

#Changing percentage datapoints (currently read as factors) to numerics:
convert_percent_to_numeric <- function(y) {
  if (as.character(y)=="")
    return(NA)
  else
    return(as.numeric(substr(y,0,nchar(as.character(y))-1)))
}

attendance[,-1] <- apply(attendance[,-1],1:2,convert_percent_to_numeric)

attendance = as.data.frame(attendance)
head(attendance)
```

Checking if the student attendance data is distinct on the studentnumber level.
```{r}
nrow(attendance)
length(unique(attendance$StudentNumber))
length(which(is.na(attendance$StudentNumber)))
```

The data is unique at the student number level and also no missing student numbers. 
```{r}
summary(attendance)
```

NA's make sense if students have yet to matriculate or have already graduated in that year. But if a student has NA for all years, that's an issue. It looks like there are 90 such students. 

```{r}
nrow(attendance[apply(attendance, 1, function(x) sum(is.na(x))) > 6,])
```

```{r}
always_absent_students <- attendance[apply(attendance, 1, function(x) sum(is.na(x))) > 6,]$StudentNumber
always_absent_student_info <- enrollment %>% filter(StudentNumber %in% always_absent_students)
sort(table((enrollment %>% filter(StudentNumber %in% always_absent_students))[,"SchoolName"]))
```

Googling these schools, it looks like most are just errors, or perhaps students left the school system mid-year. However, ASCENT (upon googling) is a transitional program for students that need just a few more credits to graduate. So it's not too surprising that we have missing attendance here. So we'll delete these in the attendance record (when we merge with other datasets later, they will be refilled with all NAs anyway).

```{r}
attendance$StudentNumber <- as.character(attendance$StudentNumber)
attendance <- attendance %>% filter(!StudentNumber %in% always_absent_students)
```


Let's check to makes sure all values are between 0 and 100 (inclusive)
```{r}
#Storing the minimum and maximum of each column as a dataframe
min_values = apply(attendance[,-1], 2, min, na.rm = TRUE)
min_values = as.data.frame(min_values)

max_values = apply(attendance[,-1], 2, max, na.rm = TRUE)
max_values = as.data.frame(max_values)

cbind(min_values, max_values)
```
As we can see academic years 2014-2015 and 2016-2017 have incorrect minimum values. Let's see the Student Numbers for 

```{r}
problem_indices <- apply(attendance[,-1], 2, function(y) which(y<0 | y>100))
```

Let's see what these students' enrollment records look like. 

```{r}
attendance[c(117203,97075,50504, 137538, 149214, 152694),]
```

And let's also see what the enrollment looks like for these students

```{r}
enrollment[c(117203,97075,50504, 137538, 149214, 152694),]
```

It's unfortunate this data is missing, but for fear of injecting false numbers into our data set we will remove these 6 students. 

```{r}
attendance <- attendance[-c(117203,97075,50504, 137538, 149214, 152694),]
```

Finally, we do some touchups and visualize the attendance data.

```{r}
attendance <- attendance %>% gather('2012','2013','2014','2015','2016','2017','2018',key = 'SpringYear',value='AttendanceRate', na.rm=TRUE)
head(attendance)
```

##CMAS Scores##

```{r}
#Skipping the first row inorder to get the appropriate headers
cmas <- s3read_using(FUN = read.csv, object = "s3://dssg2018/rawdata/DPS_CMAS_1415-1617.csv", na.strings = " ") 
head(cmas)
```

```{r}
ela_score_inds <- c(2,4,6)
ela_proficiency_inds <- c(3,5,7)
math_score_inds <- c(8,10,12)
math_proficiency_inds <- c(9,11,13)

df_to_numeric <- function(y) {
  if (as.character(y)=="")
    return(NA)
  else
    return(as.numeric(as.character(y)))
}

df_to_character <- function(y) {
  if (as.character(y)=="")
    return(NA)
  else
    return(as.character(y))
}

cmas[-(1:2),c(ela_score_inds,math_score_inds)] <- apply(cmas[-(1:2),c(ela_score_inds,math_score_inds)],1:2,df_to_numeric)
cmas[-(1:2),c(ela_proficiency_inds,math_proficiency_inds)] <- apply(cmas[-(1:2),c(ela_proficiency_inds,math_proficiency_inds)],1:2,df_to_character)

ela_scores <- cmas[-(1:2),c(1,ela_score_inds)]
colnames(ela_scores) <- c("StudentNumber","2014-2015","2015-2016","2016-2017")
ela_scores <- ela_scores %>% gather("2014-2015","2015-2016","2016-2017",key="SpringYear",value="ELAScores")
ela_proficiency <- cmas[-(1:2),c(1,ela_proficiency_inds)]
colnames(ela_proficiency) <- c("StudentNumber","2014-2015","2015-2016","2016-2017")
ela_proficiency <- ela_proficiency %>% gather("2014-2015","2015-2016","2016-2017",key="SpringYear",value="ELAProficiency")
math_scores <- cmas[-(1:2),c(1,math_score_inds)]
colnames(math_scores) <- c("StudentNumber","2014-2015","2015-2016","2016-2017")
math_scores <- math_scores %>% gather("2014-2015","2015-2016","2016-2017",key="SpringYear",value="MathScores")
math_proficiency <- cmas[-(1:2),c(1,math_proficiency_inds)]
colnames(math_proficiency) <- c("StudentNumber","2014-2015","2015-2016","2016-2017")
math_proficiency <- math_proficiency %>% gather("2014-2015","2015-2016","2016-2017",key="SpringYear",value="MathProficiency")

cmas <- merge(merge(merge(ela_scores, ela_proficiency),math_scores),math_proficiency)
all_na_rows <- apply(cmas, 1, function(x) sum(is.na(x)) > 3)
cmas <- cmas[!all_na_rows,]
cmas$SpringYear <- sub("\\d{4}-","",cmas$SpringYear)

head(cmas)
```

##Student Locations##

```{r}
student_locations <- s3read_using(FUN = read.csv, object = "s3://dssg2018/cleaned/student_block.csv")
head(student_locations)
```

We have lots of NA values. Let's map these to see what's going on. 

```{r}
leaflet(student_locations[is.na(student_locations$blocklon),]) %>% 
  addProviderTiles("CartoDB.Positron") %>%
  addCircleMarkers(
    stroke = FALSE, label=student_locations$StudentNumber,
    radius = 4, opacity = 1.0
  )
```

Looks like these are all locations outside of Denver. We will remove these for now, and consider that we can use them as a check for robustness later, or perhaps for discontinuity analysis. The ones spread across the country are likely just errors. In any case, for now we'll delete those NA values.

```{r}
student_locations <- student_locations[!is.na(student_locations$blocklon),]
```

Now we delete the sensitive student locations and clean up the column names.

```{r}
student_locations <- student_locations[,c("StudentNumber","block","geoid_num","blocklon","blocklat")]
colnames(student_locations)[4:5] <- c("lon","lat")
```

## Student Language and School Choice ##
```{r}
langAndChoice <- s3read_using(FUN = read.csv, object = "s3://dssg2018/cleaned/dps_choice_dems.csv")
head(langAndChoice)
```

Let's properly format the school year column as always. 

```{r}
levels(langAndChoice$SchoolYear) <- gsub("\\d{2}-","",levels(langAndChoice$SchoolYear))
colnames(langAndChoice)[1] <- "SpringYear"
```

Now we'll deal with language.

```{r}
lang <- langAndChoice[,c("StudentNumber","SpringYear","PrimaryHomeLanguage")]
lang <- lang[!is.na(lang$primary_home_language),]
levels(langAndChoice$PrimaryHomeLanguage)
```

There are different years for langauge, so let's make sure they're all consistent. This will take a minute to run.

```{r}
repeats <- c()
for (student_num in unique(lang$StudentNumber)) {
  if (length(unique(lang$PrimaryHomeLanguage[lang$StudentNumber==student_num]))>1)
    repeats <- c(repeats,student_num)
}
length(repeats)
```

No repeats, so we'll just take the first occurrence as the value:

```{r}
lang <- lang[!duplicated(lang$StudentNumber),]
lang$SpringYear <- NULL
```


And school choice.

```{r}
head(levels(langAndChoice$RequestedSchoolName))
```

It looks like these school names have a lot more detail than our table on school locations, so we'll keep both the school name (in case we need the info) and school number (so we can easily combine this with other info) here. 

```{r}
choice <- langAndChoice[,c("StudentNumber","SpringYear","RequestedSchoolNum","RequestedSchoolName","ChoiceSelection","ChoiceStatus")]
head(choice)
```

All said and done, we now have a ton of tables to consider:

```{r}
head(discipline)
head(ela)
head(math)
head(graduation)
head(transportation)
head(STARreading)
head(enrollment)
head(school_locations)
head(attendance)
head(cmas)
head(student_locations)
head(lang)
head(choice)
```

Finally, change all column names from camel case to lower case with hyphens, make lon into long, and remove sensitive student info from the data:

```{r}
#camel case to lower case with hyphens
changed_col_names <- function(df) {
  return(tolower(gsub("(?<=[a-z0-9])(?=[A-Z])", "_", colnames(df), perl = TRUE)))
}
colnames(discipline) <- changed_col_names(discipline)
colnames(ela) <- changed_col_names(ela)
colnames(math) <- changed_col_names(math)
colnames(graduation) <- changed_col_names(graduation)
colnames(transportation) <- changed_col_names(transportation)
colnames(STARreading) <- changed_col_names(STARreading)
colnames(enrollment) <- changed_col_names(enrollment)
colnames(school_locations) <- changed_col_names(school_locations)
colnames(attendance) <- changed_col_names(attendance)
colnames(cmas) <- changed_col_names(cmas)
colnames(student_locations) <- changed_col_names(student_locations)
colnames(lang) <- changed_col_names(lang)
colnames(choice) <- changed_col_names(choice)

#corrections
colnames(cmas)[colnames(cmas)=="elascores"]<-"ela_scores"
colnames(cmas)[colnames(cmas)=="elaproficiency"]<-"ela_proficiency"

#lon to long
colnames(school_locations)[colnames(school_locations)=="lon"] <- "long"

#remove sensitive
student_locations <- student_locations[,c("student_number","block","geoid_num")]
```

Also, we later found that one of the school names has a non-UTF-8 character in it:

```{r}
school_locations[school_locations$school_num==490,]
```

So we'll just change that manually to "Colorado High School Charter GES":

```{r}
school_locations$school_name[school_locations$school_num==490] <- "Colorado High School Charter GES"
```


And finally let's save them.

```{r}
write.csv(discipline, file="discipline.csv", row.names=FALSE, na="")
write.csv(ela, file="ela.csv", row.names=FALSE, na="")
write.csv(math, file="math.csv", row.names=FALSE, na="")
write.csv(graduation, file="graduation.csv", row.names=FALSE, na="")
write.csv(transportation, file="transportation.csv", row.names=FALSE, na="")
write.csv(STARreading, file="star_reading.csv", row.names=FALSE, na="")
write.csv(enrollment, file="enrollment.csv", row.names=FALSE, na="")
write.csv(school_locations, file="school_locations.csv", row.names=FALSE, na="")
write.csv(attendance, file="attendance.csv", row.names=FALSE, na="")
write.csv(cmas, file="cmas.csv", row.names=FALSE, na="")
write.csv(student_locations, file="student_locations.csv", row.names=FALSE, na="")
write.csv(lang, file="lang.csv", row.names=FALSE, na="")
write.csv(choice, file="choice.csv", row.names=FALSE, na="")
```