---
title: "Temporary- Incorporating Program Time"
author: "Andrew Taylor"
date: "8/3/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Intro
This is going to be merged with the dataprocessing for reschool data and deleted PRN. This only exists so we can get work done for now. **know that and move on**

##Incoporating new time data from Selam
```{r}
reschool_programs <- reschool_programs[order(reschool_programs$session_id),]
test <- read.csv("/Users/Andrew/Downloads/Detailed_B4S_Colorado_Camp_Export.csv",stringsAsFactors = FALSE,header=TRUE)
test <- test[order(test$session_id),]
reschool_programs$first_start_time <- test$first_session_start_time
reschool_programs$first_end_time <- test$first_session_end_time
reschool_programs$Posfirst_start_time <- strptime(reschool_programs$first_start_time, "%I:%M:%S %p")
reschool_programs$Posfirst_end_time <- strptime(reschool_programs$first_end_time, "%I:%M:%S %p")
reschool_programs$Hours_Day <- (reschool_programs$Posfirst_end_time - reschool_programs$Posfirst_start_time)/(60*60) #differences are stored numerically in seconds, counting up from a static date. Thus we do last minus first for the difference in seconds, divide by 60*60 for to convert to hours
```

###Spot Checking
Reveals five strange ones. There are three 0s, and two start at midnight end at 11:59 (>23 hours per day). The 0s (0 hours per day) are a for a camp, so we treat those as 24 hours. The 11:59s are for an online course, so we drop em. 
```{r}
summary(as.numeric(reschool_programs$Hours_Day))
reschool_programs$Hours_Day[reschool_programs$Hours_Day>23] <- NA
reschool_programs$Hours_Day[reschool_programs$Hours_Day==0] <- 24
summary(as.numeric(reschool_programs$Hours_Day))
```

###Get Raw Days
For sessions that are less than one week
```{r}
reschool_programs$session_date_start <- as.Date(reschool_programs$session_date_start, formate = "%Y /%m /%d")
reschool_programs$session_date_end <- as.Date(reschool_programs$session_date_end, formate = "%Y /%m /%d")
reschool_programs$raw_days <- reschool_programs$session_date_end - reschool_programs$session_date_start
summary(as.numeric(reschool_programs$raw_days))
```

###Get Working Days
Here we do some tricky things. First, for programs with "raw days" == 0, meaning they start and stop on the same day, we substitute the 0 for 1, in other treat them as being one day long. Second, we calculate number of weeks for programs with "raw days" < 7. Finally, we generate a working days col, with 1 day for 1 day long programs, the "raw days" for programs which are less than a week long, and the number of weeks * 5 (i.e, number of working days)  for all other programs.

If we get days of the week information, we'll come up with a way to amend n of days per week for both <1 week and >1 week parameters then. 
```{r}
reschool_programs$raw_days <- ifelse(reschool_programs$raw_days==0,1,reschool_programs$raw_days)
reschool_programs$raw_weeks <- ifelse(reschool_programs$raw_days<7,0,reschool_programs$raw_days/7)
reschool_programs$days <- ifelse(reschool_programs$raw_week==0,reschool_programs$raw_days,reschool_programs$raw_weeks*5)
summary(reschool_programs$days)
```

###Approximate Cost Per Day
Still waiting on days of week to refine calculation. But here we go for now.
```{r}
reschool_programs$cost_per_day <- reschool_programs$session_cost/reschool_programs$days
```

###RDS Push
yea ok, to be fair we'll need to update the markdown, but for now we're gonna push and worry about that later. 
```{r}
library("RPostgreSQL")

# loads the PostgreSQL driver
drv <- dbDriver("PostgreSQL")

con <- dbConnect(drv, dbname = "dssg2018uw",
                 host = "localhost", port = 9000,
                 user = user, password = password)

# specify data and schema and table names
data <- reschool_programs  # must be already loaded in your environment
schemaName <- "clean"
tableName <- "reschool_programs"

dbWriteTable(con,
             c(schemaName, tableName),
             value = data,
             row.names = FALSE,
             overwrite = TRUE  # overwrite an existing table
)

dbListTables(con) #check both are saved

dbDisconnect(con) 
dbUnloadDriver(drv)
```


