---
title: "Testing Blueprint4Summer clean data"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(leaflet)
library(dplyr)

# Load the data - saved locally in the same folder as this Rmd
date <- "2018-07-16"  # input this
cleandata_filename <- paste("clean-reschool-data-", date, ".csv", sep="")
cleanData <- read.csv(cleandata_filename)
```

### Basic data quality check

Are camp and session ids unique?

```{r, echo=FALSE, results='hold'}
sessionIds_are_unique <- (nrow(cleanData) == length(unique(cleanData$session_id)))
print(paste("Session ids are unique:", sessionIds_are_unique))

campIds_are_unique <- (length(unique(cleanData$camp_id)) == 
                         length(unique(cleanData$camp_name)))
print(paste("Camp ids are unique:", campIds_are_unique))
```

Are any session addresses repeated, with different zipcodes (or city/state)?

```{r, echo=FALSE, results='hold'}
session_addresses <- unique(cleanData[c("session_address_1",  "session_city",
                                          "session_state", "session_zip")])

AreDuplicateAddresses <- function(session_addresses) {
  num_repeats <- (length(session_addresses$session_address_1) -
                    length(unique(session_addresses$session_address_1)))
  print(paste("Number of session addresses that are repeated:", num_repeats)) 
}

AreDuplicateAddresses(session_addresses)
```

Look at the duplicated addresses:

```{r, echo=FALSE, results='hold'}
session_addresses[duplicated(session_addresses$session_address_1)|duplicated(session_addresses$session_address_1, fromLast=TRUE),]
```

From a Google Maps search, 2390 Havana St has the zipcode 80010, 3054 S Laredo St has the zip code 80013 and 800 Telluride St has the zipcode 80011. Manually replace these.

```{r}
cleanData[cleanData$session_address_1=="2390 Havana St", "session_zip"] <- 80010
cleanData[cleanData$session_address_1=="3054 S Laredo St", "session_zip"] <- 80013
cleanData[cleanData$session_address_1=="800 Telluride St", "session_zip"] <- 80011
```

Are there any bad duplicated addresses left? If so, add additional rows to the code above.

```{r, echo=FALSE, results='hold'}
session_addresses <- unique(cleanData[c("session_address_1",  
                                               "session_city", 
                                               "session_state", 
                                               "session_zip")])

AreDuplicateAddresses(session_addresses)
```

### Geocoding quality check

Did any geocodings fail? First we look for NAs in the added latitudes and longitudes.

```{r, echo=FALSE, results='hold'}
print(paste(sum(is.na(cleanData$lat)), "failed latitudes"))
print(paste(sum(is.na(cleanData$long)), "failed longitudes"))
```

Next we look at the geocoded program locations in an interactive map, to see if any are obviously wrong (e.g. the wrong state or city). **Zoom in and look around!!! You can click on points in the map to see the session name, camp name, and address.**

```{r, echo=FALSE}
denver_center_lat <- 39.7392
denver_center_long <- -104.9903

leaflet(cleanData, width = "100%", height = 400) %>%
  setView(lng = denver_center_long, lat = denver_center_lat, zoom = 0) %>%
  addTiles() %>%
  addCircleMarkers(~long, ~lat, stroke = FALSE,
                   popup = paste(cleanData$session_name, 
                                 cleanData$camp_name, 
                                 cleanData$session_address_1,
                                 paste(cleanData$session_city,
                                       cleanData$session_state, sep=", "), 
                                 sep="<br>")
                   )
```

For example, the session named "Summer in The Park - Fred Thomas Park" is supposed to be in Denver but has been geocoded into Arkansas for some reason. From a Google search, the correct latitude for Fred Thomas Park is 39.7525 and the correct longitude is -104.9016. We manually correct this:

```{r}
cleanData[cleanData$session_name=="Summer in The Park - Fred Thomas Park", "lat"] <- 39.7525
cleanData[cleanData$session_name=="Summer in The Park - Fred Thomas Park", "long"] <- -104.9016
```

Look at the map again:

```{r, echo=FALSE}
denver_center_lat <- 39.7392
denver_center_long <- -104.9903

leaflet(cleanData, width = "100%", height = 400) %>%
  setView(lng = denver_center_long, lat = denver_center_lat, zoom = 0) %>%
  addTiles() %>%
  addCircleMarkers(~long, ~lat, stroke = FALSE,
                   popup = paste(cleanData$session_name, 
                                 cleanData$camp_name, 
                                 cleanData$session_address_1,
                                 paste(cleanData$session_city,
                                       cleanData$session_state, sep=", "), 
                                 sep="<br>")
                   )
```

**You should emulate the code above to manually fix any other noticeably wrong geocodings.**

### Save to csv

Once everything looks good, we overwrite the previous version of the clean data with our even cleaner data.

```{r}
# Uncomment the line below to actually save the csv
# write.csv(cleanData, file = cleandata_filename, row.names = FALSE, na = "")
```

