---
title: "Update Blueprint4Summer Program Data"
author: "Kellie"
date: "8/15/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Specify where the raw input data on Blueprint4Summer programs is:

```{r input}
# get local location of github directory: 3 levels up from this file
# only works if you click "run" or "run all" in rstudio
github_dir <- dirname(dirname(dirname(rstudioapi::getActiveDocumentContext()$path)))

# name/location where raw input data lives
raw_file <- file.path(github_dir, "data", "PUT_RAW_B4S_PROGRAM_DATA_HERE", "Blueprint4Summer.csv")
```

Make sure the updated data is in the "PUT_RAW_B4S_PROGRAM_DATA_HERE" folder, and is a csv called "Blueprint4Summer.csv", otherwise this will not work!

Load in the raw data to R, as a dataframe:

```{r load-in}
raw_df <- read.csv(file = raw_file, na.strings = "", stringsAsFactors = FALSE)
```

Process the data, so that it's in the right format for the Shiny dashboard (note that this will take several minutes, as it needs to geocode the program addresses into latitudes and longitudes):

```{r process}
source( file.path(github_dir, "data_processing", "reschool", "raw_to_shiny_b4s.R") )
shiny_df <- raw_to_shiny_b4s(raw_df)
```

Specify where to save the Shiny data (don't change this or the Shiny app won't contain updated data!):

```{r output}
shiny_file <- file.path( github_dir, "data", "shiny_tables", "b4s_programs.csv" )
```

Save the Shiny data as a csv in the right location.

```{r save}
write.csv(shiny_df, file = shiny_file, row.names = FALSE)
```
