---
title: "Google Analytics API"
author: "Andrew Taylor"
date: "6/29/2018"
output: html_document
---

###API Codebook
Metrics and Dimensions codebook: https://developers.google.com/analytics/devguides/reporting/core/dimsmets

Load in libraries
```{r}
library(googleAnalyticsR)
```

###Authenticate
#VERY IMPORTANT:
The "ga_auth()" command will prompt an internet dialogue to authenticate our access to google analytics. In this case, let's use the DSSG OSR one. You have to use the browser for this, at least with this packege. 
```{r}
##Authenticate
ga_auth()

##Get accounts
account_list <- ga_account_list()

##Every account has a view ID
account_list$viewId

##Here we extract the view ID we want. One seems to be the homepage and two seems to be the search engine?
ga_id_one <- account_list$viewId[1]
ga_id_two <- account_list$viewId[2]
```

###Understanding "PathLevel" Dimension 
Of all the dimensions we could have picked why PathLevel? Why not something more for n of searches or search traffic data? 

In short- because the former too return primarily null values because they are not correctly specified for the internal search engine on the site. Or so the internet tells me. "PathLevel" is the only metric that returns details related to user searches (in this case at the user level, though we could perform "session" analysis with say, page views), in the form of the url which contains the search features.

PagePath is a "Page Level" dimensions, e.g PathLevel by "Level" risks double counting if not "Paths" are discrete categories, which it appears they're not. In other words, PathLevelOne & Two are not actually about the person, like, not about their "path", but rather about aggergating at levels like, www.example.com/path1/page1, vs. /path2/page2. Hence, why path3 == only user profiles. But why in the hell then, path1 & path2 both give us different ns of search results, and that this is also different from general path, is confusing. Do some more research. 

See the following links for more explanation:

https://www.quora.com/In-Google-analytics-custom-Reporting-what-is-Page-path-level-definition

https://analyticscanvas.com/session-vs-page-level-dimensions-google-analytics/

https://www.optimizesmart.com/complete-guide-to-dimensions-and-metrics-in-google-analytics/

###Pulling Path Level Data
```{r}
todays_date <- as.character(Sys.Date())

#By Path
PathLevel <- google_analytics(
  ga_id_two, 
  date_range=c("2018-01-01",(todays_date)),
  metrics=c('Users','timeOnPage','pageviews','uniquePageviews'),
  dimensions =c('Latitude','Longitude',"PagePath","dateHourMinute"),max=50000)

Pathtest <- google_analytics(
  ga_id_two, 
  date_range=c("2018-01-01",(todays_date)),
  metrics=c('Users','timeOnPage','pageviews','uniquePageviews'),
  dimensions =c("PagePath"),max=50000)
```

###Parsing the Clean Paths
Using our functions to fill out categories
```{r}
library(stringr)

PathLevel$CleanPath <- ParsePath(PathLevel$PagePath) #critical first stage, standardizes the different features of the html hash

PathLevel$gender <- sapply(PathLevel$CleanPath,ReturnGender)
PathLevel$distance <- sapply(PathLevel$CleanPath,ReturnDistance)
PathLevel$category <- sapply(PathLevel$CleanPath,ReturnCategory)
PathLevel$mincost <- sapply(PathLevel$CleanPath,ReturnMinCost)
PathLevel$maxcost <- sapply(PathLevel$CleanPath,ReturnMaxCost)
PathLevel$sessiontimes <- sapply(PathLevel$Clean,ReturnTime)
PathLevel$maxage <- sapply(PathLevel$CleanPath,ReturnMaxAge)
PathLevel$minage <- sapply(PathLevel$CleanPath,ReturnMinAge)
PathLevel$sort <- sapply(PathLevel$CleanPath,ReturnSort)
PathLevel$location <- sapply(PathLevel$CleanPath,ReturnLocation)
PathLevel$scholarships <- sapply(PathLevel$CleanPath,Returnscholarships)
PathLevel$specialneeds <- sapply(PathLevel$CleanPath,ReturnspecialNeeds)
PathLevel$beforeaftercare <- sapply(PathLevel$CleanPath,ReturnBeforeCare)
PathLevel$gifted <- sapply(PathLevel$CleanPath,ReturnGifted)
PathLevel$keywords <- sapply(PathLevel$CleanPath,ReturnKeyWords)
```

###Cleaning Paths
```{r}
library(stringr)
PathLevel$gender <- sapply(PathLevel$gender,CleanHash)
PathLevel$distance <-sapply(PathLevel$distance,CleanHash)
PathLevel$category <- sapply(PathLevel$category,CleanHash)
PathLevel$mincost <- sapply(PathLevel$mincost,CleanHash)
PathLevel$maxcost <- sapply(PathLevel$maxcost,CleanHash)
PathLevel$sessiontimes <- sapply(PathLevel$sessiontimes,CleanHash)
PathLevel$minage <- sapply(PathLevel$minage,CleanHash)
PathLevel$maxage <- sapply(PathLevel$maxage,CleanHash)
PathLevel$sort <- sapply(PathLevel$sort,CleanHash)
PathLevel$location <- sapply(PathLevel$location,CleanHash)
PathLevel$scholarships <- sapply(PathLevel$scholarships,CleanHash)
PathLevel$specialneeds <- sapply(PathLevel$specialneeds,CleanHash)
PathLevel$beforeaftercare <- sapply(PathLevel$beforeaftercare,CleanHash)
PathLevel$gifted <- sapply(PathLevel$gifted,CleanHash)
PathLevel$keywords <- sapply(PathLevel$keywords,CleanHash)
```

###Subset Pathlevel to searchs only
Our pull returns lots of pages that are not related to search results When we start to aggregate features, we'll notice a lot of empy searches. Here we use a function that only returns pages with search specifications, then subset for those. 
```{r}
PathLevel$cut <- ""
PathLevel$cut <- sapply(PathLevel$PagePath,FlagNonSearch)
PathLevel <- subset(PathLevel,PathLevel$cut!="cut")
```


###Aggregating Users By Category & Lat/Lon
While we may have started our report with several different metrics and uniquepaths, here we're only interested in the number of users searching for "x" features at "y" lat/lon. So we make a simpler data.frame from here on out. 
```{r}
library(haven)
Search_Summary  <- aggregate(Users ~ gender+distance+category+mincost+maxcost+sessiontimes+minage+maxage+sort+location+scholarships+specialneeds+beforeaftercare+gifted+keywords+Latitude+Longitude,data=PathLevel,sum,na.rm=TRUE,na.action=NULL)
colnames(Search_Summary) <- c("gender","distance","category","mincost"
,"maxcost","sessiontimes","minage","maxage"
,"sort","location","scholarships","specialneeds","beforeaftercare","gifted","keywords","lat","long","users")
```

```{r}
keep_only_numerics <- function(vec) {
  as.numeric(gsub("^.*[^0-9]+.*$", NA, vec))
}
relevant_columns <- c("mincost", "maxcost", "minage", "maxage") 
for (relevant_column in relevant_columns) {
  Search_Summary[,relevant_column] <- keep_only_numerics(Search_Summary[,relevant_column])
}
```

###Writing to the database
```{r}
source("../update_rds.R",chdir=TRUE)
update_rds(Search_Summary, "clean", "google_analytics", overwrite = TRUE)
```
