---
title: "Analysis NoteBook"
author: "Andrew Taylor"
date: "7/25/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Intro
This is notebook for calculations for the analysis in the report. In theory, you should just be able to run these code chunks and return the correlations requested per whatever specific access index we want. 

###RDS Pull
```{r}
library(rgeos)
library(rgdal)
library(raster)
library(RPostgreSQL)
# loads the PostgreSQL driver
drv <- dbDriver("PostgreSQL")

# creates a connection to the postgres database
# note that "con" will be used later in each connection to the database
con <- dbConnect(drv, dbname = "dssg2018uw",
                 host = "localhost", port = 9000,
                 user = user, password = password) #local source credentials

#get reschool_programs
reschool_programs <- dbGetQuery(con, "select * from clean.reschool_summer_programs")
acs_demographics <- dbGetQuery(con, "select * from clean.acs_demographics")
dps_demographics <- dbGetQuery(con, "select * from clean.dps_students")
dps_block_locations <- dbGetQuery(con, "select * from clean.dps_block_locations")
crime_rates <- dbGetQuery(con, "select * from clean.crime_rates")
driving_index <- dbGetQuery(con, "select * from clean.driving_index")
transit_index <- dbGetQuery(con, "select * from clean.transit_index")
driving_index_disabillity <- dbGetQuery(con, "select * from clean.driving_index_disabillity")
transit_index_disabillity <- dbGetQuery(con, "select * from clean.transit_index_disabillity")

#disconnect
dbDisconnect(con) 
dbUnloadDriver(drv)
```

###Census Merge
```{r}
#pre acs for merge
acs_demographics$id2 <- as.numeric(acs_demographics$id2)
acs_demographics <- acs_demographics[order(acs_demographics$id2),]
colnames(acs_demographics)[colnames(acs_demographics)=="id2"] <- "Id2"

#some variable modififcation for analysis
acs_demographics$singleparent_hh <- acs_demographics$singledad_hh + acs_demographics$singlemom_hh
acs_demographics$somecollege_or_aa_edu <- acs_demographics$somecollege_or_aa_edu/acs_demographics$ttpop_25plus_edu
acs_demographics$hsgread_or_equiv_edu <- acs_demographics$hsgread_or_equiv_edu/acs_demographics$ttpop_25plus_edu
acs_demographics$bachelors_or_higher_edu <- acs_demographics$bachelors_or_higher_edu/acs_demographics$ttpop_25plus_edu
acs_demographics$singleparent_hh <- acs_demographics$singleparent_hh/acs_demographics$ttpop_25plus_edu
acs_demographics$less_than_hs_diploma <- acs_demographics$less_than_hs_diploma/acs_demographics$ttpop_25plus_edu

acs_demographics$hsgread_or_less <- acs_demographics$hsgread_or_equiv_edu+acs_demographics$less_than_hs_diploma

#merge access index with census block data
acs_driving_AI <- merge(driving_index,acs_demographics,by="Id2")
acs_transit_AI <- merge(transit_index,acs_demographics,by="Id2")
acs_driving_disabAI <- merge(driving_index_disabillity,acs_demographics,by="Id2")
acs_transit_disabAI <- merge(transit_index_disabillity,acs_demographics,by="Id2")
```

###Census Correlations Loop for Driving
```{r}
cor_tests <-  colnames(acs_driving_AI)[c(46:85)]#prepare dataframe
index_tests <-  colnames(acs_driving_AI)[c(2:18)]#prepare things we want to loop
cor_table <- data.frame(cor_tests)

for (i in 1:length(index_tests)){ #here we loop the aggregation of students we want to test for correlations, store p_values and results in a dataframe. 
  cor_table$temp <- 0
  cor_table$p_temp <- 0
  cor_table$confit_utemp <- 0
  cor_table$confit_ltemp <- 0
  y <- index_tests[i]
  colnames(cor_table)[colnames(cor_table)=="temp"] <- paste0(y)
  colnames(cor_table)[colnames(cor_table)=="p_temp"] <- paste0("p_",y)
  colnames(cor_table)[colnames(cor_table)=="confit_utemp"] <- paste0("upper_",y)
  colnames(cor_table)[colnames(cor_table)=="confit_ltemp"] <- paste0("lower_",y)
  for (z in 1:length(cor_tests)){
    x <- cor_tests[z]
    test <- cor.test(acs_driving_AI[,paste(x)],acs_driving_AI[,paste(y)],methods=c("pearson"))
    cor_table[z,paste0(y)] <- round(as.numeric(test$estimate),3)
    cor_table[z,paste0("p_",y)] <- round(as.numeric(test$p.value),3)
    cor_table[z,paste0("upper_",y)] <- round(as.numeric(test$conf.int)[2],3)
    cor_table[z,paste0("lower_",y)] <- round(as.numeric(test$conf.int)[1],3)
  }
  print(paste(i,"working..."))
}

cor_driving_acs <- cor_table
```

###Census Correlations Loop for Transit
```{r}
cor_tests <-  colnames(acs_transit_AI)[c(46:84)]#prepare dataframe
index_tests <-  colnames(acs_transit_AI)[c(2:18)]#prepare things we want to loop
cor_table <- data.frame(cor_tests)

for (i in 1:length(index_tests)){ #here we loop the aggregation of students we want to test for correlations, store p_values and results in a dataframe. 
  cor_table$temp <- 0
  cor_table$p_temp <- 0
  cor_table$confit_utemp <- 0
  cor_table$confit_ltemp <- 0
  y <- index_tests[i]
  colnames(cor_table)[colnames(cor_table)=="temp"] <- paste0(y)
  colnames(cor_table)[colnames(cor_table)=="p_temp"] <- paste0("p_",y)
  colnames(cor_table)[colnames(cor_table)=="confit_utemp"] <- paste0("upper_",y)
  colnames(cor_table)[colnames(cor_table)=="confit_ltemp"] <- paste0("lower_",y)
  for (z in 1:length(cor_tests)){
    x <- cor_tests[z]
    test <- cor.test(acs_transit_AI[,paste(x)],acs_transit_AI[,paste(y)],methods=c("pearson"))
    cor_table[z,paste0(y)] <- round(as.numeric(test$estimate),3)
    cor_table[z,paste0("p_",y)] <- round(as.numeric(test$p.value),3)
    cor_table[z,paste0("upper_",y)] <- round(as.numeric(test$conf.int)[2],3)
    cor_table[z,paste0("lower_",y)] <- round(as.numeric(test$conf.int)[1],3)
  }
  print(paste(i,"working..."))
}

cor_transit_acs <- cor_table

```

###Census Loop Plots
####Race 
```{r}
library(ggplot2)
plot_list_t <- list()
cor_race_acs <- cor_driving_acs[c(1:7,9),]
for (i in 1:length(index_tests)){
  test <- index_tests[i]
  p <- ggplot(cor_race_acs, aes(y=cor_race_acs[,test],x=cor_tests)) +
         geom_errorbar(data=cor_race_acs,aes(ymin=cor_race_acs[,(paste0("lower_",test))],ymax=cor_race_acs[,(paste0("upper_",test))]),width=.1,color="blue3",alpha=.5) + 
         geom_point(color="blue3")+
         labs(title=paste("Student Correlations",test)) + 
  geom_hline(yintercept = 0,color="grey3") +
  ylab("Overall Driving Access Index") + 
  xlab("Percent of Students in Block Group") +
    ylim(-0.75,0.75)
  plot_list_t[[i]] <- p
  print(p)
  print(paste(i,"working..."))
}
```

####Education / Income
```{r}
library(ggplot2)
plot_list_t <- list()
cor_inc_acs <- cor_driving_acs[c(11:15,32),]
for (i in 1:length(index_tests)){
  test <- index_tests[i]
  p <- ggplot(cor_inc_acs, aes(y=cor_inc_acs[,test],x=cor_tests)) +
         geom_errorbar(data=cor_inc_acs,aes(ymin=cor_inc_acs[,(paste0("lower_",test))],ymax=cor_inc_acs[,(paste0("upper_",test))]),width=.1,color="blue3",alpha=.5) + 
         geom_point(color="blue3")+
         labs(title=paste("Student Correlations",test)) + 
  geom_hline(yintercept = 0,color="grey3") +
  ylab("Overall Driving Access Index") + 
  xlab("Percent of Students in Block Group") +
    ylim(-0.75,0.75)
  plot_list_t[[i]] <- p
  print(p)
  print(paste(i,"working..."))
}
```


##DPS demographics
###Merging block groups to DPS demographic data
```{r}
dps_demographics <- subset(dps_demographics,is.na(dps_demographics$block)==FALSE)
dps_demographics <- merge(dps_demographics,dps_block_locations,by="block")
dps_demographics$Id2 <- dps_demographics$block_group
#dps_demographics$Id2 <- as.numeric(substr(dps_demographics$block,1,11)) #FYI if you ever don't have a relational file, cutting the block level digits to this length should return block groups
```

###Aggregating to the block group level
There's def a better way to do this, but gathering by this many factors proved to be a pretty big headache for dcast and gather functions. 
```{r}
#Aggregate things we're interested in
#doing this the inefficient way

#Gender
dps_demographics$male <- ifelse(dps_demographics$gender=="M",1,0)
dps_demographics$female <- ifelse(dps_demographics$gender=="M",0,1)

male <- aggregate(male ~ Id2,data=dps_demographics,FUN=sum)
female <- aggregate(female ~ Id2,data=dps_demographics,FUN=sum)

#Race
dps_demographics$white <- ifelse(dps_demographics$race=="White",1,0)
dps_demographics$black <- ifelse(dps_demographics$race=="Black",1,0)
dps_demographics$hispanic <- ifelse(dps_demographics$race=="Hispanic",1,0)
dps_demographics$native <- ifelse(dps_demographics$race=="Nat. American",1,0)
dps_demographics$pci <- ifelse(dps_demographics$race=="Pac. Islander",1,0)
dps_demographics$asian <- ifelse(dps_demographics$race=="Asian",1,0)

white <- aggregate(white ~ Id2,data=dps_demographics,FUN=sum)
black <- aggregate(black ~ Id2,data=dps_demographics,FUN=sum)
hispanic <- aggregate(hispanic ~ Id2,data=dps_demographics,FUN=sum)
native <- aggregate(native ~ Id2,data=dps_demographics,FUN=sum)
pci <- aggregate(pci ~ Id2,data=dps_demographics,FUN=sum)
asian <- aggregate(asian ~ Id2,data=dps_demographics,FUN=sum)
other <- merge(asian,pci,by="Id2")
other <- merge(other,native,by="Id2")
other$Other <- other$asian+other$pci+other$native
other <- other[,c("Other","Id2")]

dps_agg <- merge(male,female)
dps_agg <- merge(dps_agg,white)
dps_agg <- merge(dps_agg,black)
dps_agg <- merge(dps_agg,hispanic)
dps_agg <- merge(dps_agg,other)
#dps_agg <- merge(dps_agg,native)
#dps_agg <- merge(dps_agg,pci)
#dps_agg <- merge(dps_agg,asian)

#Disability

#autism, emotional disability, intellectual disability, other health impairment, SLD, SLI
#the rest of them can be labeled as "low incidence disabilities"
dps_demographics$autism <- ifelse(dps_demographics$primary_disability=="Autism",1,0)
dps_demographics$emotional <- ifelse(dps_demographics$primary_disability=="Emotional Disability",1,0)
dps_demographics$intellectual <- ifelse(dps_demographics$primary_disability=="Intellectual Disability",1,0)
dps_demographics$SLD <- ifelse(dps_demographics$primary_disability=="SLD",1,0) #specific learning disability
dps_demographics$SLI <- ifelse(dps_demographics$primary_disability=="SLI",1,0) #speech or language impariment
dps_demographics$low_incidence <- ifelse(dps_demographics$primary_disability=="Other Health Impairment" | dps_demographics$primary_disability=="Multiple Disabilities" | dps_demographics$primary_disability=="Physical Disability" | dps_demographics$primary_disability=="Visual Impairment" | dps_demographics$primary_disability=="TBI",1,0)
dps_demographics$developdelay <- ifelse(dps_demographics$primary_disability=="Developmental Delay",1,0)
dps_demographics$any_disability <- ifelse(dps_demographics$primary_disability=="No Disability",0,1)

#dps_demographics$hearing <- ifelse(dps_demographics$primary_disability=="Hearing Impairment",1,0)
#dps_demographics$physical <- ifelse(dps_demographics$primary_disability=="Physical Disability",1,0)

any_disability <- aggregate(any_disability ~ Id2,data=dps_demographics,FUN=sum)
autism <- aggregate(autism ~ Id2, data=dps_demographics,FUN=sum)
developdelay <- aggregate(developdelay ~ Id2, data=dps_demographics,FUN=sum)
emotional <- aggregate(emotional ~ Id2,data=dps_demographics,FUN=sum)
#hearing <- aggregate(hearing ~ Id2, data=dps_demographics,FUN=sum)
#physical <- aggregate(physical ~ Id2, data=dps_demographics,FUN=sum)
SLD <- aggregate(SLD ~ Id2, data=dps_demographics,FUN=sum)
SLI <- aggregate(SLI ~ Id2, data=dps_demographics,FUN=sum)
low_incidence <- aggregate(low_incidence ~ Id2, data=dps_demographics,FUN=sum)

dps_agg <- merge(dps_agg,any_disability)
dps_agg <- merge(dps_agg,autism)
dps_agg <- merge(dps_agg,developdelay)
dps_agg <- merge(dps_agg,emotional)
#dps_agg <- merge(dps_agg,hearing)
#dps_agg <- merge(dps_agg,physical)
dps_agg <- merge(dps_agg,SLD)
dps_agg <- merge(dps_agg,SLI)
dps_agg <- merge(dps_agg,low_incidence)

#Mean LRE Scores
#lre <- aggregate(lre ~ Id2,data=dps_demographics,FUN=mean) #WRONG, LRE = LEAST RESTRICTIVE ENVIR WITH A NUMERIC CODE, SEE CODEBOOK, REDO
#dps_agg <- merge(dps_agg,lre,all.x=TRUE)

#EL Status
dps_demographics$ESL <- ifelse(dps_demographics$el_status=="EL",1,0)
ESL <- aggregate(ESL ~ Id2,data=dps_demographics,FUN=sum)

dps_agg <- merge(dps_agg,ESL)

#Primary Language
#THERE ARE SO MANY OPTIONS
#would love to do more
#let's start with english spanish and other?

dps_demographics$english <- ifelse(dps_demographics$primary_home_language=="English",1,0)
dps_demographics$spanish <- ifelse(dps_demographics$primary_home_language=="Spanish",1,0)
dps_demographics$other <- ifelse(dps_demographics$primary_home_language=="English" | dps_demographics$primary_home_language=="Spanish",0,1)

english <- aggregate(english ~ Id2, data=dps_demographics,FUN=sum)
spanish <- aggregate(spanish ~ Id2, data=dps_demographics,FUN=sum)
other <-aggregate(other ~ Id2, data=dps_demographics,FUN=sum)

dps_agg <- merge(dps_agg,english,all.x=TRUE)
dps_agg <- merge(dps_agg,spanish,all.x=TRUE)
dps_agg <- merge(dps_agg,other,all.x=TRUE)

#convert NAs to 0s

dps_agg[is.na(dps_agg)] <- 0

#get total students
dps_agg$students <- dps_agg$male+dps_agg$female

#Making proper names for plotting the access index, note that transit & driving use the same names
driving_names <- colnames(driving_index)
driving_names <- gsub("AI_","Access Index ",driving_names)
driving_names <- gsub("_"," ",driving_names)
driving_names <- gsub("has","",driving_names)
driving_names <- gsub("overall"," Overall",driving_names)
driving_names <- gsub("cost","-cost",driving_names)
driving_names <- capwords(driving_names)

colnames(driving_index) <- driving_names
colnames(transit_index) <- driving_names

driving_names_disab <- colnames(driving_index_disabillity)
driving_names_disab <- gsub("AI_","Access Index ",driving_names_disab)
driving_names_disab <- gsub("_"," ",driving_names_disab)
driving_names_disab <- gsub("has"," has",driving_names_disab)
driving_names_disab <- gsub("overall"," Overall",driving_names_disab)
driving_names_disab <- gsub("cost","-cost",driving_names_disab)
driving_names_disab <- capwords(driving_names_disab)

colnames(driving_index_disabillity) <- driving_names_disab
colnames(transit_index_disabillity) <- driving_names_disab

#merge in AI scores
dps_agg_transit <- merge(dps_agg,transit_index)
dps_special_needs_drive <- merge(dps_agg,driving_index_disabillity)
dps_special_needs_transit <- merge(dps_agg,transit_index_disabillity)
dps_agg <- merge(dps_agg,driving_index)

#merge in crime
crime <- crime_rates[,c("Id2","total_crime","violent_crime","property_crime")]
dps_agg <- merge(dps_agg,crime)
dps_agg_transit <- merge(dps_agg_transit,crime)
dps_special_needs_drive <- merge(dps_special_needs_drive,crime)
dps_special_needs_transit <- merge(dps_special_needs_transit,crime)
```

###Driving Correlations with DPS demographics

####Loop for cor table, driving
We wanna look at everything right? well there's a lot of things to look at, so here we return a dataframe of correlations for all ranges, and free vs. low and high cost. For driving.
```{r}
cor_tests <- colnames(dps_agg)[c(2:18,37:39)] #prepare dataframe
index_tests <- colnames(dps_agg)[19:36] #prepare things we want to loop
cor_table <- data.frame(cor_tests)

for (i in 1:length(index_tests)){ #here we loop the aggregation of students we want to test for correlations, store p_values and results in a dataframe. 
  cor_table$temp <- 0
  cor_table$p_temp <- 0
  cor_table$confit_utemp <- 0
  cor_table$confit_ltemp <- 0
  y <- index_tests[i]
  colnames(cor_table)[colnames(cor_table)=="temp"] <- paste0(y)
  colnames(cor_table)[colnames(cor_table)=="p_temp"] <- paste0("p_",y)
  colnames(cor_table)[colnames(cor_table)=="confit_utemp"] <- paste0("upper_",y)
  colnames(cor_table)[colnames(cor_table)=="confit_ltemp"] <- paste0("lower_",y)
  for (z in 1:length(cor_tests)){
    x <- cor_tests[z]
    test <- cor.test((dps_agg[,paste(x)]/dps_agg$students),dps_agg[,paste(y)],methods=c("pearson")) #note we are using % of students, not n of students here, to match census specifications
    cor_table[z,paste0(y)] <- round(as.numeric(test$estimate),3)
    cor_table[z,paste0("p_",y)] <- round(as.numeric(test$p.value),3)
    cor_table[z,paste0("upper_",y)] <- round(as.numeric(test$conf.int)[2],3)
    cor_table[z,paste0("lower_",y)] <- round(as.numeric(test$conf.int)[1],3)
  }
  print(paste(i,"working..."))
}

cor_driving_dps <- cor_table
```

###Loop for transit 
```{r}
cor_tests <- colnames(dps_agg)[c(2:18,37:39)] #prepare dataframe
index_tests <- colnames(dps_agg)[19:36] #prepare things we want to loop
cor_table <- data.frame(cor_tests)

for (i in 1:length(index_tests)){ #here we loop the aggregation of students we want to test for correlations, store p_values and results in a dataframe. 
  cor_table$temp <- 0
  cor_table$p_temp <- 0
  cor_table$confit_utemp <- 0
  cor_table$confit_ltemp <- 0
  y <- index_tests[i]
  colnames(cor_table)[colnames(cor_table)=="temp"] <- paste0(y)
  colnames(cor_table)[colnames(cor_table)=="p_temp"] <- paste0("p_",y)
  colnames(cor_table)[colnames(cor_table)=="confit_utemp"] <- paste0("upper_",y)
  colnames(cor_table)[colnames(cor_table)=="confit_ltemp"] <- paste0("lower_",y)
  for (z in 1:length(cor_tests)){
    x <- cor_tests[z]
    test <- cor.test((dps_agg_transit[,paste(x)]/dps_agg_transit$students),dps_agg_transit[,paste(y)],methods=c("pearson")) #note we are using % of students, not n of students here, to match census specifications
    cor_table[z,paste0(y)] <- round(as.numeric(test$estimate),3)
    cor_table[z,paste0("p_",y)] <- round(as.numeric(test$p.value),3)
    cor_table[z,paste0("upper_",y)] <- round(as.numeric(test$conf.int)[2],3)
    cor_table[z,paste0("lower_",y)] <- round(as.numeric(test$conf.int)[1],3)
  }
  print(paste(i,"working..."))
}

cor_transit_dps <- cor_table
```

###For Disabillity Index Driving
```{r}
cor_tests <- colnames(dps_special_needs_drive)[c(2:14)] #prepare dataframe
index_tests <- colnames(dps_special_needs_drive)[c(22:33)]   #prepare things we want to loop
cor_table <- data.frame(cor_tests)

for (i in 1:length(index_tests)){ #here we loop the aggregation of students we want to test for correlations, store p_values and results in a dataframe. 
  cor_table$temp <- 0
  cor_table$p_temp <- 0
  cor_table$confit_utemp <- 0
  cor_table$confit_ltemp <- 0
  y <- index_tests[i]
  colnames(cor_table)[colnames(cor_table)=="temp"] <- paste0(y)
  colnames(cor_table)[colnames(cor_table)=="p_temp"] <- paste0("p_",y)
  colnames(cor_table)[colnames(cor_table)=="confit_utemp"] <- paste0("upper_",y)
  colnames(cor_table)[colnames(cor_table)=="confit_ltemp"] <- paste0("lower_",y)
  for (z in 1:length(cor_tests)){
    x <- cor_tests[z]
    test <- cor.test((dps_special_needs_drive[,paste(x)]/dps_special_needs_drive$students),dps_special_needs_drive[,paste(y)],methods=c("pearson")) #note we are using % of students, not n of students here, to match census specifications
    cor_table[z,paste0(y)] <- round(as.numeric(test$estimate),3)
    cor_table[z,paste0("p_",y)] <- round(as.numeric(test$p.value),3)
    cor_table[z,paste0("upper_",y)] <- round(as.numeric(test$conf.int)[2],3)
    cor_table[z,paste0("lower_",y)] <- round(as.numeric(test$conf.int)[1],3)
  }
  print(paste(i,"working..."))
}

cor_driving_disab_dps <- cor_table
```

###plot loop
```{r}
library(ggplot2)
plot_list_t <- list()
for (i in 1:length(index_tests)){
  test <- index_tests[i]
  p <- ggplot(cor_table, aes(y=cor_table[,test],x=cor_tests)) +
         geom_errorbar(data=cor_table,aes(ymin=cor_table[,(paste0("lower_",test))],ymax=cor_table[,(paste0("upper_",test))]),width=.1,color="blue3",alpha=.5) + 
         geom_point(color="blue3")+
         labs(title=paste("Student Correlations",test)) + 
  geom_hline(yintercept = 0,color="grey3") +
  ylab("Overall Driving Access Index") + 
  xlab("Percent of Students in Block Group") +
    ylim(-0.75,0.75)
  plot_list_t[[i]] <- p
  print(p)
  print(paste(i,"working..."))
}
```


###Larger text
```{r}
Text12 <- element_text(size = 12)
Text16 <- element_text(size = 16)
Text24 <- element_text(size = 24)
```

###For Export - Race
```{r}
library(ggthemes)
library(ggplot2)
plot_list_t <- list()
race_cor <- cor_table[3:6,] #subset for only the variables we care about
race_cor$cor_tests <- as.character(race_cor$cor_tests)
race_cor$cor_tests <- capwords(race_cor$cor_tests)
for (i in 1:length(index_tests)){
  test <- index_tests[i]
  race <- ggplot(race_cor, aes(y=race_cor[,test],x=race_cor$cor_tests)) +
         geom_errorbar(data=race_cor,aes(ymin=race_cor[,(paste0("lower_",test))],ymax=race_cor[,(paste0("upper_",test))]),width=.1,color="blue3",alpha=.5) + 
         geom_point(color="blue3",size=3)+
         labs(title=paste(test)) + 
  geom_hline(yintercept = 0,color="grey3") +
  ylab("Driving Access Index") + 
  xlab("Percent of Students in Block Group") +
    ylim(-0.75,0.75) + theme_tufte()
  plot_list_t[[i]] <- race
  print(race + theme(axis.text.x = Text12, axis.text.y = Text12, axis.title.x = Text16, axis.title.y = Text16,title = Text24))
  print(paste(i,"working..."))
}
```

###For Export - Disabillity
```{r}
plot_list_t <- list()
disab_cor <- cor_table[c(7:13),]
disab_cor$cor_tests <- as.character(disab_cor$cor_tests)
disab_cor$cor_tests <- capwords(disab_cor$cor_tests)
for (i in 1:length(index_tests)){
  test <- index_tests[i]
  disab <- ggplot(disab_cor, aes(y=disab_cor[,test],x=disab_cor$cor_tests)) +
         geom_errorbar(data=disab_cor,aes(ymin=disab_cor[,(paste0("lower_",test))],ymax=disab_cor[,(paste0("upper_",test))]),width=.1,color="blue3",alpha=.5) + 
         geom_point(color="blue3",size=3)+
         labs(title=paste("Student Correlations",test)) + 
  geom_hline(yintercept = 0,color="grey3") +
  ylab("Driving Access Index for Special Needs Programs") + 
  xlab("Percent in Block Group") +
    ylim(-0.75,0.75)
  plot_list_t[[i]] <- disab
  print(disab)
  print(paste(i,"working..."))
}
```


###Avearge Access to Free Programs by Median HH Income
For quantiles of access, average median hh income. Also we drop block groups without students. 
```{r}
library(ggplot2)
#subset for students only
acs_driving_AI$students <- acs_driving_AI$age_less_18 - acs_driving_AI$age_less_5
acs_driving_AI <- subset(acs_driving_AI,acs_driving_AI$students>0)

#get access quantiles
access_1st <- quantile(acs_driving_AI$AI_overall_free)[[2]]
access_2nd <- quantile(acs_driving_AI$AI_overall_free)[[3]]
access_3rd <- quantile(acs_driving_AI$AI_overall_free)[[4]]

#create a frame to plot
AccessQuantiles <- c("Bottom 25%","Bottom 50%","Top 50%","Top 25%")

plot_frame$AccessQuantiles <- factor(plot_frame$AccessQuantiles, levels=c("Bottom 25%","Bottom 50%","Top 50%","Top 25%"),ordered = TRUE)
plot_frame <- data.frame(AccessQuantiles)
plot_frame$MedianHHIncome <- 0

#we probably should have done this in a loop, but because there's not many 
subset_1st <- subset(acs_driving_AI,acs_driving_AI$AI_overall_free<access_1st)

subset_belowm <- subset(acs_driving_AI,acs_driving_AI$AI_overall_free>access_1st & acs_driving_AI$AI_overall_free<access_2nd)

subset_abovem <- subset(acs_driving_AI,acs_driving_AI$AI_overall_free>access_2nd & acs_driving_AI$AI_overall_free<access_3rd)

subset_3rd <- subset(acs_driving_AI,acs_driving_AI$AI_overall_free>access_3rd)

#finishing plot frames
plot_frame[1,2] <- mean(subset_1st$median_hh_income,na.rm=TRUE)
plot_frame[2,2] <- mean(subset_belowm$median_hh_income,na.rm=TRUE)
plot_frame[3,2] <- mean(subset_abovem$median_hh_income,na.rm=TRUE)
plot_frame[4,2] <- mean(subset_3rd$median_hh_income,na.rm=TRUE)

#plot for access to free programs
ggplot(plot_frame, aes(plot_frame$AccessQuantiles,MedianHHIncome)) + geom_bar(stat = "identity",fill="#2166ac") + ylab("Median Household Income") + xlab("Access Index Quantiles") + theme_tufte() + ggtitle("Mean Household Income per Access Quantile")
```

###Datatable for average access index
###Creating logical columns for quantiles of hhincome
```{r}
median_hh_income <- median(acs_driving_AI$median_hh_income,na.rm = TRUE)
hh_1st <- quantile(acs_driving_AI$median_hh_income,na.rm=TRUE)[[2]]
hh_2nd <- quantile(acs_driving_AI$median_hh_income,na.rm=TRUE)[[3]]
hh_3rd <- quantile(acs_driving_AI$median_hh_income,na.rm=TRUE)[[4]]

acs_driving_AI$below_1st <- acs_driving_AI$median_hh_income<hh_1st

acs_driving_AI$lower_middle <- acs_driving_AI$median_hh_income>hh_1st & acs_driving_AI$median_hh_income<hh_3rd

acs_driving_AI$upper_middle <- acs_driving_AI$median_hh_income>hh_2nd & acs_driving_AI$median_hh_income<hh_3rd

acs_driving_AI$above_3rd <- acs_driving_AI$median_hh_income>hh_3rd
```

###Creating a summary access index table by hhincome quantiles
```{r}
#index_tests <- c("AI_overall","AI_overall_free","AI_art_anycost","AI_has_nature_anycost","AI_has_sports_anycost","AI_academic_anycost")
index_tests <- colnames(driving_index)[2:length(colnames(driving_index))]

summary_table <- data.frame(index_tests)
summary_table$below_1st <- 0
summary_table$lower_middle <- 0
summary_table$upper_middle <- 0
summary_table$above_3rd <- 0

for (i in 1:(length(index_tests))){
  index.i <- index_tests[i]
  subset_below1st <- subset(acs_driving_AI,acs_driving_AI$below_1st==TRUE)
  subset_lower_middle <- subset(acs_driving_AI,acs_driving_AI$lower_middle==TRUE)
  subset_upper_middle <- subset(acs_driving_AI,acs_driving_AI$upper_middle==TRUE)
  subset_above3rd <- subset(acs_driving_AI,acs_driving_AI$above_3rd==TRUE)
  summary_table[i,2] <- mean(subset_below1st[,index.i])
  summary_table[i,3] <- mean(subset_middle[,index.i])
  summary_table[i,4] <- mean(subset_above3rd[,index.i])
  summary_table[i,5] <- mean(subset_upper_middle[,index.i])
  print(paste(i,"working"))
}
```

