---
title: "Analysis NoteBook"
author: "Andrew Taylor"
date: "7/25/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Intro
This is notebook for calculations for the analysis in the report. In theory, you should just be able to run these code chunks and return the correlations requested per whatever specific access index we want. 

###RDS Pull
```{r}
library(rgeos)
library(rgdal)
library(raster)
library(tidyverse)
library(RPostgreSQL)
# loads the PostgreSQL driver
drv <- dbDriver("PostgreSQL")

# creates a connection to the postgres database
# note that "con" will be used later in each connection to the database
con <- dbConnect(drv, dbname = "dssg2018uw",
                 host = "localhost", port = 9000,
                 user = user, password = password) #local source credentials

#get reschool_programs
reschool_programs <- dbGetQuery(con, "select * from clean.reschool_summer_programs")
acs_demographics <- dbGetQuery(con, "select * from clean.acs_demographics")
dps_demographics <- dbGetQuery(con, "select * from clean.dps_students")
dps_block_locations <- dbGetQuery(con, "select * from clean.dps_block_locations")


#disconnect
dbDisconnect(con) 
dbUnloadDriver(drv)
```

###Census Merge
```{r}
#pre acs for merge
acs_demographics$id2 <- as.numeric(acs_demographics$id2)
acs_demographics <- acs_demographics[order(acs_demographics$id2),]
colnames(acs_demographics)[colnames(acs_demographics)=="id2"] <- "Id2"

#merge access index with census block data
driving_index <- driving_index[order(driving_index$Id2),]
acs_driving_AI <- merge(driving_index,acs_demographics,by="Id2")

#make an overall column
acs_driving_AI$AI_overall <- rowMeans(acs_driving_AI[,c("AI_has_nature_anycost",'AI_has_sports_anycost','AI_art_anycost','AI_academic_anycost')])
```

##Census Correlations
```{r}
library(haven)
#Stat Sig
white_cor <- cor.test(acs_driving_AI$pct_white, acs_driving_AI$AI_overall, method=c("pearson"))
black_cor <- cor.test(acs_driving_AI$pct_african, acs_driving_AI$AI_overall, method=c("pearson"))
hsp_cor <- cor.test(acs_driving_AI$pct_hispanic, acs_driving_AI$AI_overall, method=c("pearson"))


#Not Stat Sig
asn_cor <- cor.test(acs_driving_AI$pct_asian, acs_driving_AI$AI_overall, method=c("pearson"))
AmI_cor <- cor.test(acs_driving_AI$pct_americanindian, acs_driving_AI$AI_overall, method=c("pearson"))
PcI_cor <- cor.test(acs_driving_AI$pct_pacificislander,acs_driving_AI$AI_overall,method=c("pearson"))
```

###Income Correlations
Add single mom & dad? / Single parent? 
Consider some kind of transformation with median hh income so it's not $s? 
```{r}
#sig
median_income_cor <- cor.test(acs_driving_AI$median_hh_income,acs_driving_AI$AI_overall,method=c("pearson"))

#not sig
poverty_cor <- cor.test(acs_driving_AI$poverty_pct_hh,acs_driving_AI$AI_overall,method=c("pearson"))
```

###Language Correlations
```{r}
#make percent of ESLs
acs_driving_AI$pcnt_english <- acs_driving_AI$lng_english/acs_driving_AI$ttpop_18plus_lng
acs_driving_AI$pcnt_ESL <- 1-acs_driving_AI$pcnt_english

ESL_cor <- cor.test(acs_driving_AI$pcnt_ESL,acs_driving_AI$AI_overall,methods=c("pearson"))
```

##DPS demographics

###Merging block groups to DPS demographic data
```{r}
dps_demographics <- subset(dps_demographics,is.na(dps_demographics$block)==FALSE)
dps_demographics <- merge(dps_demographics,dps_block_locations,by="block")
dps_demographics$Id2 <- dps_demographics$block_group
#dps_demographics$Id2 <- as.numeric(substr(dps_demographics$block,1,11)) #FYI if you ever don't have a relational file, cutting the block level digits to this length should return block groups
```

###Aggregating to the block group level
There's def a better way to do this, but gathering by this many factors proved to be a pretty big headache for dcast and gather functions. 
```{r}
#Aggregate things we're interested in
#doing this the inefficient way

#Gender
dps_demographics$male <- ifelse(dps_demographics$gender=="M",1,0)
dps_demographics$female <- ifelse(dps_demographics$gender=="M",0,1)

male <- aggregate(male ~ Id2,data=dps_demographics,FUN=sum)
female <- aggregate(female ~ Id2,data=dps_demographics,FUN=sum)

#Race
dps_demographics$white <- ifelse(dps_demographics$race=="White",1,0)
dps_demographics$black <- ifelse(dps_demographics$race=="Black",1,0)
dps_demographics$hispanic <- ifelse(dps_demographics$race=="Hispanic",1,0)
dps_demographics$native <- ifelse(dps_demographics$race=="Nat. American",1,0)
dps_demographics$pci <- ifelse(dps_demographics$race=="Pac. Islander",1,0)
dps_demographics$asian <- ifelse(dps_demographics$race=="Asian",1,0)

white <- aggregate(white ~ Id2,data=dps_demographics,FUN=sum)
black <- aggregate(black ~ Id2,data=dps_demographics,FUN=sum)
hispanic <- aggregate(hispanic ~ Id2,data=dps_demographics,FUN=sum)
native <- aggregate(native ~ Id2,data=dps_demographics,FUN=sum)
pci <- aggregate(pci ~ Id2,data=dps_demographics,FUN=sum)
asian <- aggregate(asian ~ Id2,data=dps_demographics,FUN=sum)

dps_agg <- merge(male,female)
dps_agg <- merge(dps_agg,white)
dps_agg <- merge(dps_agg,black)
dps_agg <- merge(dps_agg,hispanic)
dps_agg <- merge(dps_agg,native)
dps_agg <- merge(dps_agg,pci)
dps_agg <- merge(dps_agg,asian)

#Disability
dps_demographics$any_disability <- ifelse(dps_demographics$primary_disability=="No Disability",0,1)
dps_demographics$autism <- ifelse(dps_demographics$primary_disability=="Autism",1,0)
dps_demographics$developdelay <- ifelse(dps_demographics$primary_disability=="Developmental Delay",1,0)
dps_demographics$emotional <- ifelse(dps_demographics$primary_disability=="Emotional Disability",1,0)
dps_demographics$hearing <- ifelse(dps_demographics$primary_disability=="Hearing Impairment",1,0)
dps_demographics$physical <- ifelse(dps_demographics$primary_disability=="Physical Disability",1,0)
dps_demographics$SLD <- ifelse(dps_demographics$primary_disability=="SLD",1,0) #specific learning disability
dps_demographics$SLI <- ifelse(dps_demographics$primary_disability=="SLI",1,0) #speech or language impariment


any_disability <- aggregate(any_disability ~ Id2,data=dps_demographics,FUN=sum)
autism <- aggregate(autism ~ Id2, data=dps_demographics,FUN=sum)
developdelay <- aggregate(developdelay ~ Id2, data=dps_demographics,FUN=sum)
emotional <- aggregate(emotional ~ Id2,data=dps_demographics,FUN=sum)
hearing <- aggregate(hearing ~ Id2, data=dps_demographics,FUN=sum)
physical <- aggregate(physical ~ Id2, data=dps_demographics,FUN=sum)
SLD <- aggregate(SLD ~ Id2, data=dps_demographics,FUN=sum)
SLI <- aggregate(SLI ~ Id2, data=dps_demographics,FUN=sum)

dps_agg <- merge(dps_agg,any_disability)
dps_agg <- merge(dps_agg,autism)
dps_agg <- merge(dps_agg,developdelay)
dps_agg <- merge(dps_agg,emotional)
dps_agg <- merge(dps_agg,hearing)
dps_agg <- merge(dps_agg,physical)
dps_agg <- merge(dps_agg,SLD)
dps_agg <- merge(dps_agg,SLI)

#Mean LRE Scores
#lre <- aggregate(lre ~ Id2,data=dps_demographics,FUN=mean) #WRONG, LRE = LEAST RESTRICTIVE ENVIR WITH A NUMERIC CODE, SEE CODEBOOK, REDO
#dps_agg <- merge(dps_agg,lre,all.x=TRUE)

#EL Status
dps_demographics$ESL <- ifelse(dps_demographics$el_status=="EL",1,0)
ESL <- aggregate(ESL ~ Id2,data=dps_demographics,FUN=sum)

dps_agg <- merge(dps_agg,ESL)

#Primary Language
#THERE ARE SO MANY OPTIONS
#would love to do more
#let's start with english spanish and other?

dps_demographics$english <- ifelse(dps_demographics$primary_home_language=="English",1,0)
dps_demographics$spanish <- ifelse(dps_demographics$primary_home_language=="Spanish",1,0)
dps_demographics$other <- ifelse(dps_demographics$primary_home_language=="English" | dps_demographics$primary_home_language=="Spanish",0,1)

english <- aggregate(english ~ Id2, data=dps_demographics,FUN=sum)
spanish <- aggregate(spanish ~ Id2, data=dps_demographics,FUN=sum)
other <-aggregate(other ~ Id2, data=dps_demographics,FUN=sum)

dps_agg <- merge(dps_agg,english,all.x=TRUE)
dps_agg <- merge(dps_agg,spanish,all.x=TRUE)
dps_agg <- merge(dps_agg,other,all.x=TRUE)

#convert NAs to 0s

dps_agg[is.na(dps_agg)] <- 0

#get total students
dps_agg$students <- dps_agg$male+dps_agg$female

#merge in AI scores
dps_agg <- merge(dps_agg,driving_index)

#merge in crime
crime <- crime_rates[,c("Id2","total_crime","violent_crime","property_crime")]
dps_agg <- merge(dps_agg,crime)
```

###Driving Correlations with DPS demographics

####Loop for cor table, driving
We wanna look at everything right? well there's a lot of things to look at, so here we return a dataframe of correlations for all ranges, and free vs. low and high cost. For driving.
```{r}
cor_tests <- colnames(dps_agg)[c(2:21,36:38)] #prepare dataframe
index_tests <- colnames(dps_agg)[23:35] #prepare things we want to loop
cor_table <- data.frame(cor_tests)

for (i in 1:length(index_tests)){ #here we loop the aggregation of students we want to test for correlations, store p_values and results in a dataframe. 
  cor_table$temp <- 0
  cor_table$p_temp <- 0
  cor_table$confit_utemp <- 0
  cor_table$confit_ltemp <- 0
  y <- index_tests[i]
  colnames(cor_table)[colnames(cor_table)=="temp"] <- paste0(y)
  colnames(cor_table)[colnames(cor_table)=="p_temp"] <- paste0("p_",y)
  colnames(cor_table)[colnames(cor_table)=="confit_utemp"] <- paste0("upper_",y)
  colnames(cor_table)[colnames(cor_table)=="confit_ltemp"] <- paste0("lower_",y)
  for (z in 1:length(cor_tests)){
    x <- cor_tests[z]
    test <- cor.test((dps_agg[,paste(x)]/dps_agg$students),dps_agg[,paste(y)],methods=c("pearson")) #note we are using % of students, not n of students here, to match census specifications
    cor_table[z,paste0(y)] <- round(as.numeric(test$estimate),3)
    cor_table[z,paste0("p_",y)] <- round(as.numeric(test$p.value),3)
    cor_table[z,paste0("upper_",y)] <- round(as.numeric(test$conf.int)[2],3)
    cor_table[z,paste0("lower_",y)] <- round(as.numeric(test$conf.int)[1],3)
  }
  print(paste(i,"working..."))
}

cor_driving_dps <- cor_table
```

###Loop for transit 
```{r}
dps_agg_transit <- dps_agg[-c(23:35)]
dps_agg_transit <- merge(dps_agg_transit,transit_index)

cor_tests <- colnames(dps_agg_transit)[c(2:21,23:25)] #prepare dataframe
index_tests <- colnames(dps_agg_transit)[26:length(colnames(dps_agg_transit))] #prepare things we want to loop
cor_table <- data.frame(cor_tests)

for (i in 1:length(index_tests)){ #here we loop the aggregation of students we want to test for correlations, store p_values and results in a dataframe. 
  cor_table$temp <- 0
  cor_table$p_temp <- 0
  cor_table$confit_utemp <- 0
  cor_table$confit_ltemp <- 0
  y <- index_tests[i]
  colnames(cor_table)[colnames(cor_table)=="temp"] <- paste0(y)
  colnames(cor_table)[colnames(cor_table)=="p_temp"] <- paste0("p_",y)
  colnames(cor_table)[colnames(cor_table)=="confit_utemp"] <- paste0("upper_",y)
  colnames(cor_table)[colnames(cor_table)=="confit_ltemp"] <- paste0("lower_",y)
  for (z in 1:length(cor_tests)){
    x <- cor_tests[z]
    test <- cor.test((dps_agg_transit[,paste(x)]/dps_agg_transit$students),dps_agg_transit[,paste(y)],methods=c("pearson")) #note we are using % of students, not n of students here, to match census specifications
    cor_table[z,paste0(y)] <- round(as.numeric(test$estimate),3)
    cor_table[z,paste0("p_",y)] <- round(as.numeric(test$p.value),3)
    cor_table[z,paste0("upper_",y)] <- round(as.numeric(test$conf.int)[2],3)
    cor_table[z,paste0("lower_",y)] <- round(as.numeric(test$conf.int)[1],3)
  }
  print(paste(i,"working..."))
}

cor_transit_dps <- cor_table
```

###plot loop
```{r}
library(ggplot2)
library(gridExtra)
plot_list_t <- list()
for (i in 1:length(index_tests)){
  test <- index_tests[i]
  p <- ggplot(cor_table, aes(y=cor_table[,test],x=cor_tests)) +
         geom_errorbar(data=cor_table,aes(ymin=cor_table[,(paste0("lower_",test))],ymax=cor_table[,(paste0("upper_",test))]),width=.1,color="blue3",alpha=.5) + 
         geom_point(color="blue3")+
         labs(title=paste("Student Correlations",test)) + 
  geom_hline(yintercept = 0,color="grey3") +
  ylab("Overall Driving Access Index") + 
  xlab("Percent of Students in Block Group")
  plot_list_t[[i]] <- p
  print(p)
  print(paste(i,"working..."))
}

```


