---
title: "Analysis NoteBook"
author: "Andrew Taylor"
date: "7/25/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Analysis Notebook
This is notebook for calculations for the analysis in the report. In theory, you should just be able to run these code chunks and return the correlations requested per whatever specific access index we want. 

###RDS Pull
```{r}
library(rgeos)
library(rgdal)
library(raster)
library(tidyverse)
library(RPostgreSQL)
# loads the PostgreSQL driver
drv <- dbDriver("PostgreSQL")

# creates a connection to the postgres database
# note that "con" will be used later in each connection to the database
con <- dbConnect(drv, dbname = "dssg2018uw",
                 host = "localhost", port = 9000,
                 user = user, password = password) #local source credentials

#get reschool_programs
#reschool_programs <- dbGetQuery(con, "select * from clean.reschool_summer_programs")
acs_demographics <- dbGetQuery(con, "select * from clean.acs_demographics")

#disconnect
dbDisconnect(con) 
dbUnloadDriver(drv)
```

###Census Merge
```{r}
acs_demographics$id2 <- as.numeric(acs_demographics$id2)
acs_demographics <- acs_demographics[order(acs_demographics$id2),]

#order access index dataframe by Id2 and extract Ai scores
summary_index_user <- summary_index_user[order(summary_index_user$Id2),]
AI <- summary_index_user$AccessIndex

#merge access index with census block data
acs_AI <- cbind(acs_demographics,AI)
colnames(acs_AI)
```

###Census Race Correlations
```{r}
library(haven)
acs_AI$pct_white <- acs_AI$pct_white/100
acs_AI$pct_african <- acs_AI$pct_african/100
acs_AI$pct_americanindian <- acs_AI$pct_americanindian/100
acs_AI$pct_asian <- acs_AI$pct_asian/100
acs_AI$pct_hispanic <- acs_AI$pct_hispanic/100

#Stat Sig
white_cor <- cor.test(acs_AI$pct_white, acs_AI$AI, method=c("pearson"))
black_cor <- cor.test(acs_AI$pct_african, acs_AI$AI, method=c("pearson"))
hsp_cor <- cor.test(acs_AI$pct_hispanic, acs_AI$AI, method=c("pearson"))


#Not Stat Sig
asn_cor <- cor.test(acs_AI$pct_asian, acs_AI$AI, method=c("pearson"))
AmI_cor <- cor.test(acs_AI$pct_americanindian, acs_AI$AI, method=c("pearson"))
PcI_cor <- cor.test(acs_AI$pct_pacificislander,acs_AI$AI,method=c("pearson"))
```

###Income Correlations
Add single mom & dad? / Single parent? 
```{r}
acs_AI$poverty_pct_hh <- acs_AI$poverty_pct_hh/100

#sig
median_income_cor <- cor.test(acs_AI$median_hh_income,acs_AI$AI,method=c("pearson"))

#not sig
poverty_cor <- cor.test(acs_AI$poverty_pct_hh,acs_AI$AI,method=c("pearson"))
```

###Language Correlations
```{r}
#make percent of ESLs
acs_AI$pcnt_english <- acs_AI$lng_english/acs_AI$ttpop_18plus_lng
acs_AI$pcnt_ESL <- 1-acs_AI$pcnt_english

ESL_cor <- cor.test(acs_AI$pcnt_ESL,acs_AI$AI,methods=c("pearson"))
```

###Geospatial Regression To Return Effect of race on access?
Why? Besides that you want to do it what is the value?
```{r}

```

