---
title: "Spatial_Autocorrelation"
author: "Andrew Taylor"
date: "8/1/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Intro
Step by step following these instructions: http://rspatial.org/analysis/rst/3-spauto.html

###Prep
```{r}
library(spdep)
p <- census_blocks #get shape file
w <- poly2nb(p, row.names = p$Id2)  #find adjacent polygons
xy <- coordinates(p) #get centroids
plot(p,col="grey",border="blue",lwd=2) #check borders with block groups
plot(w,xy,col="red",lwd=2,add=TRUE) #check connections between block groups
```

###Transform w into a spatial weights matrix
whatever the fuck that is. Whatever it is we use it later to set no adjacent pairs to 0. 
```{r}
wm <- nb2mat(w,style="B")
```


##Compute Moran's I
###Setting observations and outcomes
```{r}
#merge driving index values
p <- merge(p,driving_index,by="Id2")

#number of observations
n <- length(p)

#"y" and "y-bar" (mean value of y)
y <- p$AI_overall
ybar <- mean(y)
```

###Setting yi - ybar
```{r}
dy <- y - ybar

g <- expand.grid(dy,dy)
yiyj <- g[,1]*g[,2]
```

###Make a matrix of multiplied pairs
```{r}
pm <- matrix(yiyj,ncol=n)
```

###Multiply this matrix with the weights to set to 0 the values of pairs that art no adjacanet
```{r}
pmw <- pm*wm
```

###Here we sum the values to get part of Moran's I
```{r}
spmw <- sum(pmw)
spmw
```

###then we divide by the sum of weights
```{r}
smw <- sum(wm)
sw <- spmw/smw
```

###Then compute the inverse variance of y
```{r}
vr <- n / sum(dy^2)
```

###Final step to compute Moran's I
```{r}
MI <- vr*sw
```

###Compute Moran's I using a prebuilt function
Rather than doing everything by hand, we could just use this function. But it's good to verify we're doing things correctly. 
```{r}
ww <-  nb2listw(w, style='B')
moran(p$AI_overall,ww,n=length(ww$neighbours),S0=Szero(ww))
```

#Now for something totally different
Here we calculate local moran's I statistics per each block group in order to identify clusers of blockgroups with similar access systematically, rather than just by eyeballing the map. 

##Compute local Moran Driving
```{r}
oid <- order(p$Id2)
resI <- localmoran(p$AI_overall,ww)
results <- printCoefmat(data.frame(resI[oid,],row.names=p$Id2[oid]),check.names=false)
results$Id2 <- rownames(results)
```

###Plot Sig blocks
This map is primarily for validation, here we plot the p-values per each block group. There is really no need to repeat this per each specification. 
```{r}
library(leaflet)
library(RColorBrewer)

p <- p[order(p$Id2),]
results <- results[order(results$Id2),]

#specifiying color pallet
range <- range(results$Pr.z...0.)

pal <- colorNumeric(palette = "Greens",domain = range,reverse = TRUE)

#creating leaflet map
pLocal_Morans_I <- leaflet(census_blocks) %>%
  addTiles() %>%
  addPolygons(color = "RdYlBu",weight =1,smoothFactor = 0.5,opacity = 0.5,fillOpacity = 0.9,
              fillColor = ~pal(results$Pr.z...0.),
              popup=paste("p =",round(results$Pr.z...0.,digits=2)),
              highlightOptions = highlightOptions(color = "white", weight = 0.1)) %>%
  addLegend(pal = pal, values = ~results$Pr.z...0., opacity = 1, title = "p value") %>%
  setView(lat=39.7,lng=-104.9,zoom=11.5)
```

###Plot morans I of sig blocks Driving
**NOTE** that a positive LMI means the area is similar to it's neighbors, i.e part of spatial cluster. A higher positive LMI denotes that the blockgroup is more similar to it's neighbors, though the actual score is relative to the mean of the sample, and, according to at least one internet not guide, not inherently meaningful. 
```{r}
results_sig <- results
results_sig$Ii <- ifelse(results$Pr.z...0.>=0.05,NA,results$Ii)

p <- p[order(p$Id2),]
results_sig <- results_sig[order(results_sig$Id2),]

#specifiying color pallet
pal <- colorNumeric(palette = "Purples",domain = NULL,na.color="Grey",reverse=TRUE)

#creating leaflet map
Driving_Morans_I <- leaflet(p) %>%
  addTiles() %>%
  addPolygons(color = "RdYlBu",weight =1,smoothFactor = 0.5,opacity = 0.5,fillOpacity = 0.9,
              fillColor = ~pal(results_sig$Ii),
              popup=paste("lMI =",round(results_sig$Ii,digits=2)),
              highlightOptions = highlightOptions(color = "white", weight = 0.1)) %>%
  addLegend(pal = pal, values = ~results_sig$Ii, opacity = 1, title = "Driving LMI") %>%
  setView(lat=39.7,lng=-104.9,zoom=11.5)
```

##LMI For Transit
```{r}
p <- census_blocks #get shape file
p <- merge(p,transit_index,by="Id2")
oid <- order(p$Id2)
resI <- localmoran(p$AI_overall,ww)
results <- printCoefmat(data.frame(resI[oid,],row.names=p$Id2[oid]),check.names=false)
results$Id2 <- rownames(results)

results_sig <- results
results_sig$Ii <- ifelse(results$Pr.z...0.>=0.05,NA,results$Ii)

p <- p[order(p$Id2),]
results_sig <- results_sig[order(results_sig$Id2),]

#specifiying color pallet
pal <- colorNumeric(palette = "Purples",domain = NULL,na.color="Grey",reverse=TRUE)

#creating leaflet map
Transit_Morans_I <- leaflet(p) %>%
  addTiles() %>%
  addPolygons(color = "RdYlBu",weight =1,smoothFactor = 0.5,opacity = 0.5,fillOpacity = 0.9,
              fillColor = ~pal(results_sig$Ii),
              popup=paste("lMI =",round(results_sig$Ii,digits=2)),
              highlightOptions = highlightOptions(color = "white", weight = 0.1)) %>%
  addLegend(pal = pal, values = ~results_sig$Ii, opacity = 1, title = "Transit LMI") %>%
  setView(lat=39.7,lng=-104.9,zoom=11.5)
```

##LMI For Driving, Free
```{r}
p <- census_blocks #get shape file
p <- merge(p,driving_index,by="Id2")
oid <- order(p$Id2)
resI <- localmoran(p$AI_overall_free,ww)
results <- printCoefmat(data.frame(resI[oid,],row.names=p$Id2[oid]),check.names=false)
results$Id2 <- rownames(results)

results_sig <- results
results_sig$Ii <- ifelse(results$Pr.z...0.>=0.05,NA,results$Ii)

p <- p[order(p$Id2),]
results_sig <- results_sig[order(results_sig$Id2),]

#specifiying color pallet
pal <- colorNumeric(palette = "Oranges",domain = NULL,na.color="Grey",reverse=TRUE)

#creating leaflet map
DrivingFree_Morans_I <- leaflet(p) %>%
  addTiles() %>%
  addPolygons(color = "RdYlBu",weight =1,smoothFactor = 0.5,opacity = 0.5,fillOpacity = 0.9,
              fillColor = ~pal(results_sig$Ii),
              popup=paste("lMI =",round(results_sig$Ii,digits=2)),
              highlightOptions = highlightOptions(color = "white", weight = 0.1)) %>%
  addLegend(pal = pal, values = ~results_sig$Ii, opacity = 1, title = "Driving-Free LMI") %>%
  setView(lat=39.7,lng=-104.9,zoom=11.5)
```

##LMI For Transit Free
```{r}
p <- census_blocks #get shape file
p <- merge(p,transit_index,by="Id2")
oid <- order(p$Id2)
resI <- localmoran(p$AI_overall_free,ww)
results <- printCoefmat(data.frame(resI[oid,],row.names=p$Id2[oid]),check.names=false)
results$Id2 <- rownames(results)

results_sig <- results
results_sig$Ii <- ifelse(results$Pr.z...0.>=0.05,NA,results$Ii)

p <- p[order(p$Id2),]
results_sig <- results_sig[order(results_sig$Id2),]

#specifiying color pallet
pal <- colorNumeric(palette = "Oranges",domain = NULL,na.color="Grey",reverse=TRUE)

#creating leaflet map
TransitFree_Morans_I <- leaflet(p) %>%
  addTiles() %>%
  addPolygons(color = "RdYlBu",weight =1,smoothFactor = 0.5,opacity = 0.5,fillOpacity = 0.9,
              fillColor = ~pal(results_sig$Ii),
              popup=paste("lMI =",round(results_sig$Ii,digits=2)),
              highlightOptions = highlightOptions(color = "white", weight = 0.1)) %>%
  addLegend(pal = pal, values = ~results_sig$Ii, opacity = 1, title = "Transit-Free LMI") %>%
  setView(lat=39.7,lng=-104.9,zoom=11.5)
```

