---
title: "Analysis_correlation"
author: "Haowen Zheng"
date: "8/1/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include=FALSE}
library(tidyverse)
library(rgdal)
library(rgeos)
library(leaflet)
library(ggplot2)
library(tigris)
library(ggthemes)
```

## Get connected to RDS
```{r}
# install.packages("RPostgreSQL")
require("RPostgreSQL")

# loads the PostgreSQL driver
drv <- dbDriver("PostgreSQL")
source('/Users/luna/Desktop/DSSG/ors/other/keyforRDS')
# creates a connection to the postgres database
# note that "con" will be used later in each connection to the database
con <- dbConnect(drv, dbname = "dssg2018uw",
                 host = "localhost", port = 9000,
                 user = user, password = password)
```

## Get the data for analysis
```{r}
# get the indices
driving_index <- dbGetQuery(con, "select * from clean.driving_index")
transit_index <- dbGetQuery(con, "select * from clean.transit_index")

# get acs_demographics flat file
acs_demographics <- dbGetQuery(con, "select * from clean.acs_demographics")
# get acs_shapefile 
acs_shape <- readOGR(dsn="/Users/luna/Desktop/DSSG/ors/data/clean data/shape_census", layer="shape_census")

# get student data
students = dbGetQuery(con, "select * from clean.dps_students")
student_location <- dbGetQuery(con, "select * from clean.dps_block_locations")
student_new <- students %>% filter(!is.na(block)) %>% left_join(student_location) #keep the students who have address info
student_bg <- student_new %>% group_by(block_group) %>% summarise(student_n = n()) #the number of students in each block group
student_DPS_bg <- dbGetQuery(con, "select * from shiny.dps_student_aggregate_bgs")

# get neighborhood shapefile data
nbhd_shape <- readOGR(dsn="/Users/luna/dssg2018/osr_dssg2018/data/nbhd_dem_shapes", layer="nbhd_dem_shapes")
colnames(nbhd_shape@data)[1] <- "nbhd_name"

# get relationship file
bg_nbhd <- dbGetQuery(con, "select * from clean.blockgroup_nbhds")
```


## Analysis
### How equal/unequal is access to out-of-school resources in Denver? Is access to osr proportional to the number of students? 

```{r}
#compare numbers of census student-age population with DPS student 
# merge the acs_demographics and student_bg together
colnames(acs_demographics)[2] <- "Id2"
colnames(student_bg)[1] <- "Id2"
analysis1 <- merge(acs_demographics, student_bg, by = "Id2")
analysis1$student_age_n <- analysis1$age_less_18 - analysis1$age_less_5

hist(analysis1$student_age_n)

# dtk how to deal with the extreme values, see how it performs if we aggregate thing to the neighborhood level

acs_subset <- analysis1 %>% select(Id2,student_age_n)
colnames(bg_nbhd)[3] <- "Id2"
analysis_subset <- merge(acs_subset, bg_nbhd, by = "Id2")
 # aggregate by nbhd
analysis_subset_aggregated2 <- analysis_subset %>% group_by(nbhd_name) %>% summarise(student_age_a = sum(student_age_n))

# It seems like we have three options for the student number!
head(analysis_subset_aggregated1)
summary(nbhd_shape@data$AGE_5_T) #2015estimate, acs Year5-17
# summary(analysis_subset_aggregated1$student_n_a) #DPS choice students
summary(analysis_subset_aggregated2$student_age_a) #2016estimate, acs Year5-18
# We should probably use the last one as the DPS choice is biased. 

# how does the access index look like if we aggregate it on neighborhood level? 
driving_index_subset <- driving_index %>% select(Id2, AI_overall)
driving_index_subset <- merge(bg_nbhd, driving_index_subset, by = "Id2")
 # aggregate by nbhd 
driving_index_a <- driving_index_subset %>% group_by(nbhd_name) %>% summarise(AI_score = mean(AI_overall))

head(nbhd_merged_AI@data)
hist(driving_index_a$AI_score)
 # merge and plot on the map
nbhd_merged_AI <- geo_join(nbhd_shape, driving_index_a, by = "nbhd_name")
nbhd_merged_AI2 <- geo_join(nbhd_merged_AI, analysis_subset_aggregated2, by = "nbhd_name")

pal <- colorQuantile(
  palette = "Blues",
  domain = nbhd_merged_AI@data$AI_score, n = 10)
pop <- paste("Nbhd:", nbhd_merged_AI@data$nbhd_name,"<br/>",
             "AI:", nbhd_merged_AI@data$AI_score)

# map for access index on neighborhood level
leaflet(nbhd_merged_AI) %>%
  addTiles() %>%
  addPolygons(stroke = FALSE, weight = 1, smoothFactor = 0.5,
    opacity = 1.0, fillOpacity = 0.5, 
    color = ~pal(nbhd_merged_AI@data$AI_score), popup = pop) %>%
  setView(lat=39.7,lng=-104.9,zoom=10)

# let's drop the two zeros and see if Gini coefficient makes sense! 
nbhd_merged_AI_sn <- merge(driving_index_a, analysis_subset_aggregated2, by = "nbhd_name")  %>% 
  filter(student_age_a != 0)
nbhd_merged_AI_sn$AI_weighted <- nbhd_merged_AI_sn$AI_score/nbhd_merged_AI_sn$student_age_a
hist(nbhd_merged_AI_sn$AI_weighted)

library(ineq)

# Lorenz curve
plot(Lc(nbhd_merged_AI_sn$AI_weighted, nbhd_merged_AI_sn$student_age_a),
     col="darkred",lwd=2, xlab = "Percentage of Students", ylab = "Cumulative Share of 'Access'") 
ineq(nbhd_merged_AI_sn$AI_score,type="Gini")  #0.59

library(DescTools)
Gini(nbhd_merged_AI_sn$AI_weighted, nbhd_merged_AI_sn$student_age_a)

plot(Lc(nbhd_merged_AI_sn$AI_score,  nbhd_merged_AI_sn$student_age_a),
     col="darkred",lwd=2, xlab = "Percentage of Students", ylab = "Cumulative Share of 'Access'") 

# to examine the data closer, let's rank the three variables
nbhd_merged_AI_sn <- nbhd_merged_AI_sn %>% 
  arrange(-AI_score) %>% 
  mutate(rank_AI = row_number()) %>% 
  arrange(-student_age_a) %>% 
  mutate(rank_student_n = row_number()) %>%
  arrange(-AI_weighted) %>% 
  mutate(rank_AI_weighted = row_number())
# This is potentially problematic because the weighted AI score ended up driven purely by the number of students in the neighborhoods.

# Therefore we change the unit of analysis to student level 
plot(Lc(nbhd_merged_AI_sn$AI_weighted, nbhd_merged_AI_sn$student_age_a),col="darkred",lwd=2) 

# map for student-age population distribution on neighborhood level
leaflet(nbhd_merged_AI2) %>%
  addTiles() %>%
  addPolygons(color = "#444444", weight = 1, smoothFactor = 0.5,
   opacity = 1.0, fillOpacity = 0.5, 
   fillColor = ~colorQuantile("YlOrRd", nbhd_merged_AI2@data$student_age_a)(nbhd_merged_AI2@data$student_age_a)) %>%
  setView(lat=39.7,lng=-104.9,zoom=10)

head(nbhd_merged_AI_sn)
```

Let's make a scatter plot showing the number of programs and access scores. 
```{r}
# driving overall index
ggplot(nbhd_merged_AI_sn, aes(student_age_a, AI_score)) +
  geom_point(aes(color= rank_AI_weighted>66)) +
  scale_color_manual(values = c("black", "red")) +
  geom_text(aes(label=ifelse(rank_AI_weighted>66,as.character(nbhd_name),'')),hjust=0.7,vjust=1, size=3) +
  xlab("Number of Students in Each Neighborhood") +
  ylab("Access Score (Neighborhood)") +
  theme_bw() +
  theme(legend.position="none") +
  ggtitle("Neighborhoods with Low Access and High Student Age Population") 
```

We care about the following indices, AI for overall programs for driving and transit, AI for free programs for driving and transit. Therefore we want to see if there are any interesting patterns for them, too. 

```{r}
full_index <- merge(driving_index, transit_index, by = "Id2")
names(bg_nbhd)[3] <- "Id2"
selected_index <- full_index %>% 
  dplyr::select(Id2, AI_overall.x, AI_overall_free.x, AI_overall.y, AI_overall_free.y) %>% # x driving, y transit
  merge(bg_nbhd, by = "Id2")

index_agg <- selected_index %>%
    group_by(nbhd_name) %>%
    summarise_at(vars(matches("overall")), funs(mean)) %>%
    left_join(., analysis_subset_aggregated2, by = c("nbhd_name")) %>%
    mutate_at(vars(matches("overall")),funs(weighted = . / analysis_subset_aggregated2$student_age_a)) %>%
    filter(student_age_a != 0) %>%  #drop the two neighborhoods that have 0 values for number of students.
    mutate_at(vars(matches("weighted")), funs(rank = dense_rank(desc(.))))

# define the function that plots Lorenz curve
plotLC <- function(x){
  p <- plot(Lc(x, index_agg$student_age_a),col="darkred",lwd=2)
  return(p)
}

index_agg1 <- index_agg %>% dplyr::select("AI_overall.x_weighted", "AI_overall_free.x_weighted", "AI_overall.y_weighted", "AI_overall_free.y_weighted")
apply(index_agg1, 2, FUN =  plotLC) #plot

# define the scatter plot
  plotScatter <- function(var1, var2, df) {
            print(data.frame(df) %>%
            ggplot( aes(student_age_a, df[,var1])) +
            geom_point(aes(color= df[,var2]>66)) +
            scale_color_manual(values = c("black", "red")) +
            geom_text(aes(label=ifelse(df[,var2]>66,as.character(nbhd_name),'')),hjust=0.7,vjust=1, size=3.5) +
            xlab("Number of students in each neighborhood") +
            ylab("Access Index score") +
            theme_bw() +
            theme(legend.position="none") +
            ggtitle("Neighborhoods with low access and high student age population"))
  }
  
var_list = list(c("AI_overall.x","AI_overall.x_weighted_rank"),c("AI_overall_free.x","AI_overall_free.x_weighted_rank"),
                c("AI_overall.y","AI_overall.y_weighted_rank"),c("AI_overall_free.y","AI_overall_free.y_weighted_rank"))   

for(i in 1:length(var_list)){
  plotScatter(var1=var_list[[i]][1],var2=var_list[[i]][2],df=index_agg)
  }
```

### Leaflet map 
To get the context of things.. 
```{r}
nbhd_merged_AI <- geo_join(nbhd_merged_AI, index_agg, by = "nbhd_name", how = "inner")

leaflet(nbhd_merged_AI) %>%
  addTiles() %>%
  addPolygons(stroke = FALSE, weight = 1, smoothFactor = 0.5,
    opacity = 1.0, fillOpacity = 0.5, 
    color = ~pal(nbhd_merged_AI@data$AI_score), popup = pop) %>% #replace AI_score with each index
  setView(lat=39.7,lng=-104.9,zoom=10)
```

## Tidy the DPS data for later use.
```{r}
head(student_bg) # number of students in each block group 
head(student_new) # each row being a student
head(student_DPS_bg) # aggregate to block group levels

student_DPS_bg$perc_otheraces_students <- 
  100 - student_DPS_bg$perc_hispanic_students - student_DPS_bg$perc_black_students - student_DPS_bg$perc_white_students
# Because the numbers were rounded up, we got 4 values that are negative. Change them to zeros. 
student_DPS_bg$perc_otheraces_students <- ifelse(
  student_DPS_bg$perc_otheraces_students < 0, 0, student_DPS_bg$perc_otheraces_students )
colnames(student_DPS_bg)[1] <- "Id2"
```


## spatial autocorrelation - Local Moran's I 

First we get a new data set with Moran's I statistics. 

```{r}
#merge census shapefile to driving index for overall programs
acs_shape@data$Id2 <- gsub("0(\\d{11})","\\1",acs_shape@data$Id2)
indices <- full_index %>% select(Id2, AI_overall.x, AI_overall.y)
moran.data <- geo_join(acs_shape, indices, by = "Id2")

library(spdep)
w <- poly2nb(moran.data, row.names=moran.data@data$Id2, queen=TRUE) #find adjacent polygons
#We do this using the ‘Queen’s case’ setting, meaning that adjacent areas which share either a border or a corner are counted as neighbours.
summary(w)
 # convert it to a spatial weights matrix
listw <-  nb2listw(w, style="W") 

# calculate the local moran of the distribution of index score
lmoran <- localmoran(moran.data@data$AI_overall.x, listw) #driving index for all programs
summary(lmoran)

# padronize the variable and save it to a new column
moran.data@data$s_index <- scale(moran.data@data$AI_overall.x)  %>% as.vector()

# create a spatially lagged variable and save it to a new column
moran.data@data$lag_s_index <- lag.listw(listw, moran.data@data$s_index)

# summary of variables, to inform the analysis
summary(moran.data@data$s_index)
summary(moran.data@data$lag_s_index)

# create a new variable identifying the moran plot quadrant for each observation, dismissing the non-significant ones

moran.data@data$SPATIAL_LAG_CAT <- 
  ifelse(moran.data@data$s_index>0 & moran.data@data$lag_s_index>0, "High-High",
       ifelse(moran.data@data$s_index>0 & moran.data@data$lag_s_index<0, "High-Low",
              ifelse(moran.data@data$s_index<0 & moran.data@data$lag_s_index<0, "Low-Low",
                     ifelse(moran.data@data$s_index<0 & moran.data@data$lag_s_index>0, "Low-High",
       "Equivalent"))))

# add Moran's I back into the shape file
moran.data@data$lmoran_i <- lmoran[,1]
moran.data@data$lmoran_p <-lmoran[,5]
moran.data@data$lmoran_sig <-lmoran[,5]<0.05

moran.data@data$id <- rownames(moran.data@data)
```

Then we move on to sample comparison. Start by creating the samples. 
```{r}
# filter out the significant areas
df <- fortify(moran.data, region="id")
df <- left_join(moran.data@data, df)
df$SPATIAL_LAG_CAT2 <- 
  factor(ifelse(df$lmoran_sig == TRUE, df$SPATIAL_LAG_CAT, "NotSig"
      ))

df2 <- df %>% filter(lmoran_sig == TRUE)

Id2_moran <- unique(df2$Id2) #159 bgs

# filter out the high areas and the low areas
high_low <- moran.data@data %>%
  filter(Id2 %in% Id2_moran) %>%
  filter(SPATIAL_LAG_CAT == "High-High" |SPATIAL_LAG_CAT == "Low-Low")

# merge it with student info
high_low_students <- left_join(high_low, student_DPS_bg, by = "Id2")

# create new function to aggregate the sample so we can plot them later 
create_samples2 <- function(df) {

sample_comparison = 
  bind_rows(
    df %>%
    filter(SPATIAL_LAG_CAT == "High-High") %>%
    summarise_at(vars(col),funs( mean = mean(.,na.rm=T))) %>%
    mutate(sample = 'High Access Areas'),
    
    df %>%
    filter(SPATIAL_LAG_CAT == "Low-Low") %>%
    summarise_at(vars(col),funs( mean = mean(.,na.rm=T))) %>%
    mutate(sample = 'Low Access Areas')

)

    return(sample_comparison)
}

# prepare the columns we need 
# edu from census data
high_low_students$PCT_College <- high_low_students$BACHELO*100/high_low_students$TTPOP_2
high_low_students$PCT_L_HS <- high_low_students$LESS_TH*100/high_low_students$TTPOP_2
high_low_students$PCT_HS <- high_low_students$HSGREAD*100/high_low_students$TTPOP_2
high_low_students$PCT_SC <- high_low_students$SOMECOL*100/high_low_students$TTPOP_2
# nativity from census
high_low_students$PCT_nativity <- high_low_students$Native*100/high_low_students$TTL_ppl
# English learner from DPS
high_low_students$PCT_ELL <- high_low_students$perc_nonenglish_students

col = c("perc_hispanic_students", "perc_black_students", "perc_white_students", "perc_otheraces_students", "Mdn_HH_", "PCT_College", "PCT_L_HS", "PCT_HS", "PCT_SC", "PCT_nativity", "PCT_ELL")

# prepare the datasets 
moran_sample <- create_samples2(high_low_students)
# sample for high access areas
high <- high_low_students %>% filter(SPATIAL_LAG_CAT == "High-High")
# sample for low access areas
low <- high_low_students %>% filter(SPATIAL_LAG_CAT == "Low-Low")
```

## We are now ready for plotting and Mann-Whitney Test! 

### Race

Plot the comparision of races for high vs. low access areas. 

```{r}
plotdata <- moran_sample %>% 
  select(sample, perc_black_students_mean, perc_otheraces_students_mean, perc_hispanic_students_mean,
         perc_white_students_mean) %>%
  gather(Race, value, -sample) 
plotdata$Race <- factor(plotdata$Race, 
                        levels = c("perc_hispanic_students_mean", "perc_white_students_mean", "perc_black_students_mean", "perc_otheraces_students_mean"))


ggplot(plotdata, aes(sample, value, fill = Race)) +
  geom_bar(stat = "identity", position = "dodge") +
  xlab("") +
  ylab("Percentages") +
  ggtitle("Student Racial Compostion for High/Low Access Areas") +
  scale_fill_discrete(name="Race and Ethnicity") +
  scale_fill_manual(labels = c("Hispanic", "White", "Black", "Others"), values=c("#2b8cbe", "#fb6a4a", "#fecc5c", "#78c679")) +
  theme_bw() +
  theme(axis.text.x = element_text(size = 13),
        legend.text=element_text(size=10),
        plot.title = element_text(size=20)) +
  scale_y_continuous(expand = c(0,5))

ggsave(file="race.png", width=7, height=7, dpi=300)
```

T Tests
```{r}
# white 
t.test(high$perc_white_students, low$perc_white_students) #p< 0.001
# Hispanic
t.test(high$perc_hispanic_students, low$perc_hispanic_students) #p< 0.001

# black
t.test(high$perc_black_students, low$perc_black_students) #p< 0.001

# others
t.test(high$perc_otheraces_students, low$perc_otheraces_students) #p = 0.9
```

Get the table: Index scores by race and program types. 

```{r}
# I need the row number of students races from DPS 
DPS_race <- dbGetQuery(con, "select * from public.dps_student_aggregate_bgs")
colnames(DPS_race)[1] <- "Id2"

DPS_race <- merge(DPS_race, indices, by = "Id2")
DPS_race$Others <- DPS_race$unique_students - DPS_race$hispanic_students - DPS_race$black_students - DPS_race$white_students

indices_selected <- DPS_race[,9:10]
colnames(DPS_race)[4:6] <- c("Hispanic", "Black", "White")

get_race_access_means <- function(access_inds){
  race_col_names <- c("Hispanic","White","Black", "Others")
  race_names <- c("Hispanic", "White", "Black", "Others")
  race_access_list <- list()
  for(i in 1:length(race_col_names)){
    race_pops <- DPS_race[,race_col_names[i]]
    tot_race_pop <- sum(DPS_race[,race_col_names[i]])
    race_access_list[race_names[i]] <- sum(access_inds*(race_pops/tot_race_pop))
  }
  return(as.vector(race_access_list))
}

result <- sapply(indices_selected, get_race_access_means)

test <- as.data.frame(result)
result1 <- test %>% mutate(race = c("Hispanic", "White", "Black", "Others"))

result1
```

### Social-economic status

Income
```{r}
t.test(high$Mdn_HH_ , low$Mdn_HH_)
```


EDU
```{r}
sample_compare_edu <- moran_sample %>% 
  select(sample, PCT_L_HS_mean, PCT_HS_mean, PCT_SC_mean, PCT_College_mean) %>%
  gather(edu, value, -sample) 
sample_compare_edu$edu <-
  factor(sample_compare_edu$edu,
  levels = c("PCT_L_HS_mean", "PCT_HS_mean", "PCT_SC_mean", "PCT_College_mean"))

  ggplot(sample_compare_edu) +
  geom_bar(aes(sample, value, fill = edu), stat = "identity", position = "dodge", width = 0.7) +
  xlab("") +
  ylab("Percentages") +
  ggtitle("Education Attainments for High/Low Access Areas",
          subtitle = "Adults Over 25 Years Old") +
  scale_fill_manual(labels = c("Less Than High School", "High School and Equivalent", "Some College", "Bachelor's Degree and Above"), values=c("#2b8cbe", "#fb6a4a", "#fecc5c", "#78c679")) +
  guides(fill=guide_legend(title=" ")) +
  theme_bw() +
  theme(legend.position = "bottom",
        axis.text.x = element_text(size = 13),
        legend.text=element_text(size=10),
        plot.title = element_text(size=16)) 
  
ggsave(file="edu.png", width=7, height=7, dpi=300)
```

```{r}
t.test(high$PCT_College, low$PCT_College)
t.test(high$PCT_L_HS, low$PCT_L_HS)
t.test(high$PCT_SC, low$PCT_SC)
t.test(high$PCT_HS, low$PCT_HS)
```

English Language Learners
```{r}
# non_english_speakers
ggplot(moran_sample, aes(sample, PCT_ELL_mean)) +
  geom_bar(stat = "identity", fill = "#2166ac") +
  xlab("Access Index Quantiles") +
  ylab("Percentages") +
  ggtitle("Students as English Learners for Areas with Different Levels of Access to OSR") +
  theme_tufte() 

t.test(high$PCT_ELL, low$PCT_ELL)
```

Nativity
```{r}
t.test(high$PCT_nativity, low$PCT_nativity)
```

## Close the link when you are done.
```{r}
dbDisconnect(con) 
dbUnloadDriver(drv)
```
