---
title: "Census descriptives"
author: "Haowen Zheng"
date: "7/25/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

## Import data
```{r}
census_shape <- readOGR(dsn="../data/clean data/shape_census", layer="shape_census")
```

## Census Descriptives
### How are the population distributed? 
```{r}
# set up bins manually
bins_ttlpop <- c(0, 500, 1000, 1500, 2000, max(census_shape@data$TTL_ppl))
pal_ttlpop <- colorBin("Greys", domain = census_shape@data$TTL_ppl, bins = bins_ttlpop)

leaflet(census_shape) %>%
  addTiles() %>%
  addPolygons(stroke = FALSE, smoothFactor = 0.2, fillOpacity = 1,
    color = ~pal_ttlpop(census_shape@data$TTL_ppl)
  ) %>%
  addLegend("bottomright", pal = pal_ttlpop, values = ~census_shape@data$TTL_ppl,
    title = "Total Population",
    opacity = 1
  )

# too many bins use neighborhood data

# aggregate ttp to nbhd level and merge with nbhd shapefile and plot
acs_demographics$ttl_population
subset_acs <- acs_demographics %>% select(Id2, ttl_population,pct_less_18,pct_white,pct_african, pct_otherraces, pct_hispanic, median_hh_income)
merged <- merge(subset_acs, bg_nbhd, by = "Id2")
col <- c("ttl_population","pct_less_18","pct_white","pct_african", "pct_otherraces", "pct_hispanic", "median_hh_income")

new <- merged %>%
    group_by(nbhd_name) %>%
    summarise_at(vars(col), funs(mean))
 

new_shape <- geo_join(nbhd_shape, new, "nbhd_name", "nbhd_name")


summary(new$ttl_population)

bins_ttlpop <- c(0, 500, 1000, 1500, 2000, max(new_shape@data$ttl_population))
pal_ttlpop <- colorBin("Greys", domain = new_shape@data$ttl_population, bins = bins_ttlpop)

leaflet(new_shape) %>%
  addTiles() %>%
  addPolygons(stroke = FALSE, smoothFactor = 0.2, fillOpacity = 1,
    color = ~pal_ttlpop(new_shape@data$ttl_population)
  ) %>%
  addLegend("bottomright", pal = pal_ttlpop, values = ~new_shape@data$ttl_population,
    title = "Total Population",
    opacity = 1
  )

summary(nbhd_shape@data$MED_HH_)
```

### Racial distributions
Spatial distribution of African Americans over Denver (block group). Other maps for other races can be generated using the same sets of codes. 
```{r}
leaflet(census_shape) %>%
  addTiles() %>%
  addPolygons(color = "#444444", weight = 1, smoothFactor = 0.5,
  opacity = 1.0, fillOpacity = 0.5, 
  fillColor = ~colorFactor(colors5, census_shape@data$PCT_Afr)(census_shape@data$PCT_Afr))
```

### Income & poverty distribution in Denver
```{r}
#why are there so many N.A.s in census block group data Median Household Income? 
sum(is.na(shape_census@data$Mdn_HH_)) #15
 # From Factfinder: either no sample observations or too few sample observations were available to compute an estimate, or a ratio of medians cannot be calculated because one or both of the median estimates falls in the lowest interval or upper interval of an open-ended distribution.

#Let's use the nbhd level data from oepn denver 2015
nbhd_demo <- readOGR(dsn = "/Users/luna/dssg2018/osr_dssg2018/data/nbhd_dem_shapes", layer = "nbhd_dem_shapes")

sum(is.na(nbhd_demo@data$MED_HH_)) #check if there are N.A.s. None, good to go! 

# map the census median income 
popup.income <-  paste("Name:", nbhd_demo@data$NBHD_NA,"<br/>",
              "MedianHHIncome:",nbhd_demo@data$MED_HH_)
pal <- colorQuantile(
  palette = "Blues",
  domain = nbhd_demo@data$MED_HH_, n = 5)

# map for costs of programs on income, neighborhood level 
leaflet(nbhd_demo) %>%
  addTiles() %>% 
  addPolygons(stroke = FALSE, smoothFactor = 0.5, popup = popup.income, 
    fillOpacity = 0.8, color = ~pal(nbhd_demo@data$MED_HH_)) %>%
  setView(lat=39.7,lng=-104.9,zoom=10)

unique(nbhd_demo@data$NBHD_NA) # 78 unique nbhds
summary(nbhd_demo@data$MED_HH_)
boxplot(nbhd_demo@data$MED_HH_,data=nbhd_demo@data)
```

The same pattern shows itself on block group level, too.
```{r}
pal2 <- colorQuantile(
  palette = "Blues",
  domain = census_shape@data$Mdn_HH_, n = 5)

m_income <- leaflet(census_shape) %>%
  addTiles() %>%
  addPolygons(stroke = FALSE, smoothFactor = 0.5, popup = popup.income, 
    fillOpacity = 0.8, color = ~pal2(census_shape@data$Mdn_HH_)) %>%
  addCircles(data = reschool_programs,opacity = 0.4, 
             col = ~colorFactor(col3, reschool_programs$factor_cost)(reschool_programs$factor_cost), weight = 4) %>%
  setView(lat=39.7,lng=-104.9,zoom=10) 

# What if we add another layer of poverty? Map the rate of households below poverty line.
leaflet(census_shape) %>%
  addTiles() %>%
  addPolygons(stroke = FALSE, smoothFactor = 0.5, popup = popup.income, 
    fillOpacity = 0.8, color = ~pal2(census_shape@data$Mdn_HH_), group = "Income") %>%
  addPolygons(color = "#444444", weight = 1, smoothFactor = 0.5,
    opacity = 1.0, fillOpacity = 0.5, 
    fillColor = ~colorFactor(colors5, census_shape@data$P_PCT_H)(census_shape@data$P_PCT_H), group = "Poverty") %>%
  addCircles(data = reschool_programs,opacity = 0.4, 
             col = ~colorFactor(col3, reschool_programs$factor_cost)(reschool_programs$factor_cost), 
             weight = 4, group = "Programs") %>%
  setView(lat=39.7,lng=-104.9,zoom=10) %>%
  addLayersControl(
    overlayGroups = c("Programs", "Income","Poverty"),
                        options = layersControlOptions(collapsed = F))
```

### Education
```{r}
head(census_shape@data)

# high school degree
# bechelor's degree

census_shape@data$pct_Bache <- census_shape@data$BACHELO/census_shape@data$TTPOP_2

pal2 <- colorQuantile(
  palette = "Blues",
  domain = census_shape@data$pct_Bache, n = 5)

leaflet(census_shape) %>%
  addTiles() %>%
  addPolygons(stroke = FALSE, smoothFactor = 0.5, popup = popup.income, 
    fillOpacity = 0.8, color = ~pal2(census_shape@data$pct_Bache)) %>%
  setView(lat=39.7,lng=-104.9,zoom=10) 
```

### Crime
```{r}
# crimes 
head(crimes)
  #too many crimes to plot

head(hatecrimes)
 
# hatecrimes
leaflet(census_shape) %>%
  addTiles() %>% 
  addPolygons(color = "#444444", weight = 1, smoothFactor = 0.5,
    opacity = 0.5) %>%
  addCircles(data = reschool_programs,opacity = 0.4, 
             col = ~colorFactor(color3, reschool_programs$factor_cost)(reschool_programs$factor_cost), weight = 2) %>%
  addCircles(data = hatecrimes,opacity = 0.4, 
             col = "purple", weight = 2) %>%
  setView(lat=39.7,lng=-104.9,zoom=10)  
```


## Sample comparison between DPS choice students and census student age population
### Map population distribution
```{r}
# merge student_location into students data set
student_new <- students %>% filter(!is.na(block)) %>% left_join(student_location)

# merge the number of DPS students (per block group) into census data set
student_bg <- student_new %>% group_by(block_group) %>% summarise(student_n = n())
student_bg$block_group <- paste("0", student_bg$block_group, sep = "")
summary(student_bg$student_n)

census_shape2 <- geo_join(census_shape, student_bg, "Id2", "block_group", how = "left")

head(census_shape2@data)

# I used age5-18 as student age population for the below map.
census_shape2@data$census_student_n <- census_shape2@data$Ag_L_18 - census_shape2@data$Ag_Ls_5
census_shape2@data$pct_census_student <- census_shape2@data$census_student_n/census_shape2@data$TTL_ppl

#show the student age population from census and DPS choice 
leaflet(census_shape2) %>%
  addTiles() %>%
  addPolygons(color = "#444444", weight = 1, smoothFactor = 0.5,
   opacity = 1.0, fillOpacity = 0.5, 
   fillColor = ~colorQuantile("YlOrRd", census_shape2@data$census_student_n)(census_shape2@data$census_student_n), 
   group = "Census") %>%
# let's see how it changes if we change the raw numbers from census to the percentage of student age population
#  addPolygons(color = "#444444", weight = 1, smoothFactor = 0.5,
#    opacity = 1.0, fillOpacity = 0.5, 
#    fillColor = ~colorQuantile("YlOrRd", census_shape2@data$pct_census_student)(census_shape2@data$pct_census_student), 
#    group = "Census") %>%
  addPolygons(color = "#444444", weight = 1, smoothFactor = 0.5,
    opacity = 1.0, fillOpacity = 0.5, popup = info,
    fillColor = ~colorQuantile("YlOrRd", census_shape2@data$student_n)(census_shape2@data$student_n), group = "DPS") %>%
  setView(lat=39.7,lng=-104.9,zoom=10) %>%
  addLayersControl(
    overlayGroups = c("Census", "DPS"),
                        options = layersControlOptions(collapsed = F))
```
