---
title: "student_subset_bias"
author: "Joe"
date: "7/11/2018"
output: html_document
---

```{r}
require("RPostgreSQL")
require("tidyverse")
require("leaflet")
require("leaflet.extras")
require("rgdal")
require("RColorBrewer")
```

```{r}
# if knitting, must set "mypath" manually
#mypath <- dirname(rstudioapi::getActiveDocumentContext()$path)
#setwd(mypath)

# loads the PostgreSQL driver
drv <- dbDriver("PostgreSQL")

# make a file called "cred.txt" of the form
#
# user: "YOUR_AWS_USERNAME"
# password: "YOUR_AWS_PASSWORD"
#
# in the directory above the osr_dssg2018 folder

source(file.path(dirname(dirname(mypath)),"cred.txt"))

# creates a connection to the postgres database
# note that "con" will be used later in each connection to the database
con <- dbConnect(drv, dbname = "dssg2018uw",
                 host = "localhost", port =9000,
                 user = user, password = password)
```

```{r}
google_analytics <- dbGetQuery(con, "select * from clean.google_analytics")
programs <- dbGetQuery(con, "select * from clean.reschool_summer_programs")
museums <- dbGetQuery(con, "select * from clean.museums")
libraries <- dbGetQuery(con, "select * from clean.libraries")
fields <- dbGetQuery(con, "select * from clean.fields")
```

Let's look at the percentage of students entering kindegarten, 6th, and 9th grade who are in the choice program. 

```{r}
students <- dbGetQuery(con, "select * from clean.dps_students")
choice <- dbGetQuery(con, "select * from clean.dps_choice")
enrollment <- dbGetQuery(con, "select * from clean.dps_enrollment")

parts <- data.frame()
totals <- data.frame()
for (grade in 0:12) {
  for (given_year in 2014:2018) {
    students_with_choice <- choice %>% filter(spring_year %in% given_year) %>% select(student_number)
    students_with_choice <- students_with_choice$student_number
    
    part <- enrollment %>% filter(student_number %in% students_with_choice) %>% filter(spring_year %in% given_year) %>% filter(grade_level %in% grade) %>% nrow
    
    total <- enrollment %>% filter(spring_year %in% given_year) %>% filter(grade_level %in% grade) %>% nrow
    
    parts[as.character(grade),as.character(given_year)] <- part
    totals[as.character(grade),as.character(given_year)] <- total
  }
}

proportion <- parts / totals
print("Percentage of students by grade and year who did choice enrollment:")
print(round(proportion*100,1))
```

Unfortunately, not all of the choice students have an associated address:

```{r}
students_with_locations <- unique(students$student_number[!is.na(students[,"block"])])
print(paste("Number of students who did choice enrollment but don't have address on record:",length(unique(choice$student_number)) - length(students_with_locations)))
```

So we need to additionally get rid of those when considering the percentage of students for whom we have addresses in the given year: 

```{r}
parts_both <- data.frame()
totals_both <- data.frame()
for (grade in 0:12) {
  for (given_year in 2014:2018) {
    
    students_with_choice <- choice %>% filter(spring_year %in% given_year) %>% select(student_number)
    students_with_choice <- students_with_choice$student_number
    
    part <- enrollment %>% filter(spring_year %in% given_year) %>% filter(grade_level %in% grade) %>% filter(student_number %in% students_with_locations) %>% filter(student_number %in% students_with_choice) %>% nrow
    
    total <- enrollment %>% filter(spring_year %in% given_year) %>% filter(grade_level %in% grade) %>% nrow
    
    parts_both[as.character(grade),as.character(given_year)] <- part
    totals_both[as.character(grade),as.character(given_year)] <- total
  }
}

proportion_both <- parts_both / totals_both
print("Percentage of students by grade and year who have address on record and did choice enrollment:")
print(round(proportion_both*100,1))
```

Finally, we want to know which student addresses we actually have by considering that when they choose in one year, we know their address for all years before and all years after. 

```{r}
parts_both <- data.frame()
totals_both <- data.frame()
for (grade in 0:12) {
  for (given_year in 2014:2018) {
    
    part <- enrollment %>% filter(spring_year %in% given_year) %>% filter(grade_level %in% grade) %>% filter(student_number %in% students_with_locations) %>% nrow
    
    total <- enrollment %>% filter(spring_year %in% given_year) %>% filter(grade_level %in% grade) %>% nrow
    
    parts_both[as.character(grade),as.character(given_year)] <- part
    totals_both[as.character(grade),as.character(given_year)] <- total
  }
}

proportion_both <- parts_both / totals_both
print("Percentage of students by grade and year who have address on record:")
print(round(proportion_both*100,1))
```

As a final number, we can simply look at the percentage of students for whom we have addresses, for all years (2011-2018) and for years during which we had student choice (2014-2018):

```{r}
paste("Percentage of students for whom we have addresses:",round(length(students_with_locations)/length(unique(enrollment$student_number))*100,1),"%")
paste("Percentage of students enrolled somewhere between 2014 and 2018 for whom we have addresses:",round(length(students_with_locations)/length(unique(enrollment[enrollment$spring_year %in% 2014:2018,"student_number"]))*100,1),"%")
```

Now we'd like to see how different our choice students are from all other choices, to see whether our sample is biased. 

```{r}
columns = c("race","gender","el_status","primary_disability","has_transportation")
for (column in columns) {
  tmp_df <- data.frame()
  groups <- unique(students[,column])
  for (group in groups) {
    tmp_df[group,"address"] <- round((sum((students$student_number %in% students_with_locations) & (students[,column]==group)) / length(students_with_locations))*100,1)
    tmp_df[group,"no_address"] <- round((sum((!students$student_number %in% students_with_locations) & (students[,column]==group)) / (nrow(students)-length(students_with_locations)))*100,1)
  }
  print(column)
  print(tmp_df)
  barplot(t(tmp_df),beside=TRUE,main=column,cex.names=.5,las=2,legend.text = TRUE)
}
```

Not too bad. Let's also check test scores.

```{r}
test <- dbGetQuery(con, "select * from clean.dps_cmas")

tmp_df <- data.frame()

math_scores <- test[!is.na(test$math_scores),c("student_number","math_scores")]
ela_scores <- test[!is.na(test$ela_scores),c("student_number","ela_scores")]
math_scores <- aggregate(math_scores ~ student_number, data=math_scores, mean)
ela_scores <- aggregate(ela_scores ~ student_number, data=ela_scores, mean)

#to convert raw test scores to standardized z scores
make_z_score <- function(vec) {
  return((vec-mean(vec))/sd(vec))
}

math_scores$math_scores <- make_z_score(math_scores$math_scores)
ela_scores$ela_scores <- make_z_score(ela_scores$ela_scores)

tmp_df["ela","address"] <- mean(ela_scores[ela_scores$student_number %in% students_with_locations,"ela_scores"])
tmp_df["ela","no_address"] <- mean(ela_scores[!ela_scores$student_number %in% students_with_locations,"ela_scores"])
tmp_df["math","address"] <- mean(math_scores[math_scores$student_number %in% students_with_locations,"math_scores"])
tmp_df["math","no_address"] <- mean(math_scores[!math_scores$student_number %in% students_with_locations,"math_scores"])

barplot(t(tmp_df),beside=TRUE,main="Test Score Comparison",cex.names=1,las=2,legend.text = TRUE)
```

```{r}
star <- dbGetQuery(con, "select * from clean.dps_star_reading")

get_mode <- function(v) {
   uniqv <- unique(v)
   uniqv[which.max(tabulate(match(v, uniqv)))]
}

tmp_df <- data.frame()

star_proficiency <- aggregate(reading_level ~ student_number, data=star, get_mode)

groups <- unique(star_proficiency$reading_level)
for (group in groups) {
    tmp_df[group,"address"] <- round((sum((star_proficiency$student_number %in% students_with_locations) & (star_proficiency$reading_level==group))/sum(star_proficiency$student_number %in% students_with_locations))*100,1)
    tmp_df[group,"no_address"] <- round((sum((!star_proficiency$student_number %in% students_with_locations) & (star_proficiency$reading_level==group))/sum(!star_proficiency$student_number %in% students_with_locations))*100,1)
}

barplot(t(tmp_df),beside=TRUE,main="Reading Score Comparison",cex.names=1,las=2,legend.text=TRUE, args.legend = list(x = "topleft"))
```

Finally, we want to understand how discipline differs between the groups of students. 

```{r}
discipline <- dbGetQuery(con, "select * from clean.dps_discipline")

tmp_df <- data.frame()

iss <- discipline[,c("student_number","number_iss")]
oss <- discipline[,c("student_number","number_oss")]
expulsions <- discipline[,c("student_number","number_expulsions")]

iss <- aggregate(number_iss ~ student_number, data=iss, mean)
oss <- aggregate(number_oss ~ student_number, data=oss, mean)
expulsions <- aggregate(number_expulsions ~ student_number, data=expulsions, mean)

tmp_df["in_school_suspension","address"] <- mean(iss[iss$student_number %in% students_with_locations,"number_iss"])
tmp_df["out_of_school_suspension","address"] <- mean(oss[oss$student_number %in% students_with_locations,"number_oss"])
tmp_df["expulsion","address"] <- mean(expulsions[expulsions$student_number %in% students_with_locations,"number_expulsions"])

tmp_df["in_school_suspension","no_address"] <- mean(iss[!iss$student_number %in% students_with_locations,"number_iss"])
tmp_df["out_of_school_suspension","no_address"] <- mean(!oss[oss$student_number %in% students_with_locations,"number_oss"])
tmp_df["expulsion","no_address"] <- mean(expulsions[!expulsions$student_number %in% students_with_locations,"number_expulsions"])

barplot(t(tmp_df),beside=TRUE,main="Average Discipline Record Comparison",cex.names=.7,las=1,legend.text = TRUE)
```