---
title: "data_exploration"
author: "Joe"
date: "7/11/2018"
output: html_document
---

```{r}
require("RPostgreSQL")
require("tidyverse")
require("leaflet")
require("leaflet.extras")
require("rgdal")
require("RColorBrewer")
```

We will analyze Blueprint4Summer search data, which includes the zip code searched for. 

```{r}
MakeDir("raw_data")  # make a raw data directory, if one doesn't exist yet

# Download .zip to raw data directory, then unzip in temporary directory.
url <- paste("https://www.denvergov.org/media/gis/DataCatalog/", zipname, "/shape/", zipname, ".zip", sep="")
tempDir <- tempdir()
file <- file.path("raw_data", paste(zipname, ".zip", sep=""))
download.file(url, file)
unzip(file, exdir = tempDir)

# Read in shapefile from unzipped data and return result
spdf <- readOGR(dsn = tempDir, zipname)
return(spdf)

# Read shapefile into SpatialPointsDataFrame.
spdf <- readOGR(dsn = "~/Documents/Experiences/Applications/UWashington/Project/osr_dssg2018/data/zip_codes", "cb_2017_us_zcta510_500k")

spdf@data$GEOID10 <- as.character(spdf@data$GEOID10)
```

Consider all zip codes included in the search data (some of which are outside of Denver, and some are missing from the set of Denver zip codes).

```{r}
google_analytics <- dbGetQuery(con, "select * from clean.google_analytics")
programs <- dbGetQuery(con, "select * from clean.reschool_summer_programs")
museums <- dbGetQuery(con, "select * from clean.museums")
libraries <- dbGetQuery(con, "select * from clean.libraries")
fields <- dbGetQuery(con, "select * from clean.fields")

google_analytics$location <- gsub(".*(80\\d{3}).*","\\1",google_analytics$location)
searches_with_locations <- google_analytics[grep("80\\d{3}",google_analytics$location),]
unique_search_locations <- unique(searches_with_locations$location)
```

The zip codes we care to save are those in the search data and those in Denver generally:

```{r}
relevant_zip_codes <- spdf[spdf@data$GEOID10 %in% unique_search_locations,]
# create a shapefile so we don't have to push such a large dataset
writeOGR(relevant_zip_codes,"../data/zip_codes","zip_codes", driver="ESRI Shapefile")
relevant_zip_codes <- readOGR(dsn="../data/zip_codes")
```

Oddly enough, 3 of the searched zip codes don't exist:

```{r}
unique_search_locations[!unique_search_locations %in% relevant_zip_codes@data$GEOID10]
```

Let's check out the frequencies of searches by zip code:

```{r}
table(searches_with_locations$location)
```

We'll be excluding 2+3+7=12 zip codes, but for now let's just forget them and move on (we'll still have 3,000 searches total):

```{r}
searches_with_locations <- searches_with_locations[searches_with_locations$location %in% unique_search_locations[unique_search_locations %in% relevant_zip_codes@data$GEOID10],]

table(searches_with_locations$location)
```

We can look at the included areas with leaflet, weighted by the number of searches in that area:

```{r}
search_frequencies <- as.vector(table(searches_with_locations[searches_with_locations$category=="athletic","location"]))
leaflet() %>% 
  addProviderTiles("CartoDB.Positron") %>%
  addPolygons(
    data = relevant_zip_codes,
    fillColor = ~colorQuantile("YlOrRd",search_frequencies)(search_frequencies),
    popup = as.character(search_frequencies)
    ) %>%
  addCircleMarkers(
    data=programs,
    radius=5,
    popup = programs$camp_name
  ) %>%
  addScaleBar() %>%
  setView(lat=39.7,lng=-104.9,zoom=10)
```

We also can look at the distribution of programs by type.

We'll begin by considering where athletic programs are. We note that there are very few in the north, which also is a lower income community. 

```{r}
colors5 <- c('#ffffb2','#fecc5c','#fd8d3c','#f03b20','#bd0026') 

popup_string="Income:"
col_name="Mdn_HH_"

spdf@data$color <- 1
breaks <- c(30000,50000,70000,90000)
for (i in 1:length(breaks)) {
  spdf@data$color[spdf@data[,col_name] > breaks[i]] <- i+1
}

lab_factor <- c("Less than 30,000", 
                "Between 30,000 and 50,000", 
                "Between 50,000 and 70,000", 
                "Between 70,000 and 90,000",
                "Greater than 90,000")
spdf@data$color <- factor(spdf@data$color, label = lab_factor)

#search_frequencies <- as.vector(table(searches_with_locations$location)
leaflet() %>% 
  addProviderTiles("CartoDB.Positron") %>%
  addPolygons(
    data=spdf,
    fillColor = ~colorFactor(colors5,spdf@data[,col_name])(spdf@data[,col_name]),
    opacity=.1,
    popup = paste(popup_string,round((spdf@data[,col_name]),1)),
    fillOpacity = 1,
    weight=1
  ) %>%
  addLegend(colors=colors5,labels=lab_factor) %>%
  addCircleMarkers(
    data=programs[programs$has_sports,],
    color='green',
    opacity=10,
    radius=2,
    popup=programs[programs$has_athletic,"session_name"]
  ) %>%
  # addCircleMarkers(
  #   data=fields,
  #   color='blue',
  #   radius=2,
  #   popup=fields$sport
  # ) %>%
  addLegend(labels=c('athletic programs'),
            colors=c('green')
  ) %>%
  setView(lat=39.7,lng=-104.9,zoom=10)
```

```{r}
colors5 <- c('#ffffb2','#fecc5c','#fd8d3c','#f03b20','#bd0026') 

popup_string="Income:"
col_name="Mdn_HH_"

spdf@data$color <- 1
breaks <- c(30000,50000,70000,90000)
for (i in 1:length(breaks)) {
  spdf@data$color[spdf@data[,col_name] > breaks[i]] <- i+1
}

lab_factor <- c("Less than 30,000", 
                "Between 30,000 and 50,000", 
                "Between 50,000 and 70,000", 
                "Between 70,000 and 90,000",
                "Greater than 90,000")
spdf@data$color <- factor(spdf@data$color, label = lab_factor)

#search_frequencies <- as.vector(table(searches_with_locations$location)
leaflet() %>% 
  addProviderTiles("CartoDB.Positron") %>%
  addPolygons(
    data=spdf,
    fillColor = ~colorFactor(colors5,spdf@data[,col_name])(spdf@data[,col_name]),
    opacity=.1,
    popup = paste(popup_string,round((spdf@data[,col_name]),1)),
    fillOpacity = 1,
    weight=1
  ) %>%
  addLegend(colors=colors5,labels=lab_factor) %>%
  addCircleMarkers(
    data=programs,#[programs$has_sports,],
    color='green',
    opacity=10,
    radius=2,
    popup=programs$session_name#[programs$has_athletic,"session_name"]
  ) %>%
  # addCircleMarkers(
  #   data=fields,
  #   color='blue',
  #   radius=2,
  #   popup=fields$sport
  # ) %>%
  addLegend(labels=c('athletic programs'),
            colors=c('green')
  ) %>%
  setView(lat=39.7,lng=-104.9,zoom=10)
```

Interestingly, there are resources already there which could be used for sports programs. Mapping also athletic fields:

```{r}
colors5 <- c('#ffffb2','#fecc5c','#fd8d3c','#f03b20','#bd0026') 

popup_string="Income:"
col_name="Mdn_HH_"

spdf@data$color <- 1
breaks <- c(30000,50000,70000,90000)
for (i in 1:length(breaks)) {
  spdf@data$color[spdf@data[,col_name] > breaks[i]] <- i+1
}

lab_factor <- c("Less than 30,000", 
                "Between 30,000 and 50,000", 
                "Between 50,000 and 70,000", 
                "Between 70,000 and 90,000",
                "Greater than 90,000")
spdf@data$color <- factor(spdf@data$color, label = lab_factor)

#search_frequencies <- as.vector(table(searches_with_locations$location)
leaflet() %>% 
  addProviderTiles("CartoDB.Positron") %>%
  addPolygons(
    data=spdf,
    fillColor = ~colorFactor(colors5,spdf@data[,col_name])(spdf@data[,col_name]),
    opacity=.1,
    popup = paste(popup_string,round((spdf@data[,col_name]),1)),
    fillOpacity = 1,
    weight=1
  ) %>%
  addLegend(colors=colors5,labels=lab_factor) %>%
  addCircleMarkers(
    data=programs[programs$has_sports,],
    color='green',
    opacity=10,
    radius=2,
    popup=programs[programs$has_sports,"session_name"]
  ) %>%
  # addCircleMarkers(
  #   data=fields,
  #   color='blue',
  #   radius=2,
  #   popup=fields$sport
  # ) %>%
  addLegend(labels=c('athletic programs',''),
            colors=c('green','blue')
  ) %>%
  setView(lat=39.7,lng=-104.9,zoom=10)
```


```{r}
# Thanks Haowen
colors5 <- c('#ffffb2','#fecc5c','#fd8d3c','#f03b20')#,'#bd0026') 

popup_string="Black percentage:"
col_name="PCT_Afr"

spdf@data$color <- 1
breaks <- c(20,40,60)
for (i in 1:length(breaks)) {
  spdf@data$color[spdf@data[,col_name] > breaks[i]] <- i+1
}

lab_factor <- c("Less than 20", 
                "Between 20 and 40", 
                "Between 40 and 60", 
                "Between 60 and 80")
                #"Greater than 80")
spdf@data$color <- factor(spdf@data$color, label = lab_factor)

#search_frequencies <- as.vector(table(searches_with_locations$location)
leaflet() %>% 
  addProviderTiles("CartoDB.Positron") %>%
  addPolygons(
    data=spdf,
    fillColor = ~colorFactor(colors5,spdf@data[,col_name])(spdf@data[,col_name]),
    opacity=.1,
    popup = paste(popup_string,round((spdf@data[,col_name]),1)),
    fillOpacity = 1,
    weight=1
  ) %>%
  addLegend(colors=colors5,labels=lab_factor) %>%
  addCircleMarkers(
    data=programs[programs$has_sports,],
    color='green',
    opacity=10,
    radius=2,
    popup=programs[programs$has_sports,"session_name"]
  ) %>%
  # addCircleMarkers(
  #   data=fields,
  #   color='blue',
  #   radius=2,
  #   popup=fields$sport
  # ) %>%
  addLegend(labels=c('athletic programs','fields'),
            colors=c('green','blue')
  ) %>%
  setView(lat=39.7,lng=-104.9,zoom=10)
```

We can look at the locations of the searchers (based on IP address), and we find that there are unfortunately only 2 distinct locations in Denver.

```{r}
google_analytics_aggregated <- aggregate(users ~ lat + long, data=google_analytics, sum)

leaflet() %>% 
  addProviderTiles("CartoDB.Positron") %>%
  # addCircleMarkers(
  #   data = google_analytics %>% filter(lat < 39.8 & long > -105.1 & lat > 39.6 & long < -104.8),
  #   stroke = FALSE, fillOpacity = .5, color = 'blue'
  # ) %>%
  addCircleMarkers(
    data = google_analytics_aggregated,
    stroke = FALSE, fillOpacity = .5, color = 'yellow', popup = as.character(google_analytics_aggregated$users)
  ) %>%
  setView(lat=39.7,lng=-104.9,zoom=10)
```

We also would like to understand how long programs generally last: 

```{r}
programs$length <- as.numeric(as.Date(programs[!duplicated(programs$session_id),"session_date_end"]) - as.Date(programs[!duplicated(programs$session_id),"session_date_start"]))
print(table(programs[!duplicated(programs$session_name),]$length))
plot(table(programs[!duplicated(programs$session_name),]$length))
```

One big fear for the access index is that we will need to adjust program cost, or potentially even transportation time, for the number of days the programs last. By manual inspection of the websites for the programs the program with 119 is online, and some of the monthlong ones meet just once a week. So we may have to manually check those that are greater than about 4 days. 

```{r}
sum(programs[!duplicated(programs$session_name),]$length > 4)
```

Manually inputting the true number of days would be somewhat time-consuming, though doable. But the main problem is that it would not be very reproducible.

```{r}
#when you're done, close the connection and unload the driver 
dbDisconnect(con) 
dbUnloadDriver(drv)
```

```{r}
View(programs)
```
